### 2025-02-11 draft_v1/draft_v1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage[]{palatino}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{amssymb}
\usepackage{booktabs} % Add to your preamble for cleaner table lines
\usepackage{makecell} % Add to your preamble for multi-line cells
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes.geometric}

%% Define TikZ arrow styles
\newcommand{\starstar}{% 
\begin{tikzpicture}[baseline=-3pt]
    \draw [{Rays[n=6]}-{Rays[n=6]}] (0,0) -- (0.55,0);
\end{tikzpicture}
}

\newcommand{\circstar}{% 
\begin{tikzpicture}
    \draw [{Circle[open]}-{Rays[n=6]}] (0,0) -- (0.55, 0);
\end{tikzpicture}
}

\newcommand{\starcirc}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-{Circle[open]}] (0,0) -- (0.55, 0);
\end{tikzpicture}
}

\newcommand{\stararrow}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-{Straight Barb[length=2.5pt]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\arrowstar}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-{Rays[n=6]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\circarrow}{% 
\begin{tikzpicture}
    \draw [{Circle[open]}-{Straight Barb[length=2.5pt]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailcirc}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [-{Circle[open]}] (0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\circtail}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [{Circle[open]}-] (0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\circirc}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [{Circle[open]}-{Circle[open]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailstar}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [-{Rays[n=6]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\startail}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailarrow}{% 
\begin{tikzpicture}
    \draw [-{Straight Barb[length=2.5pt]}](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\arrowtail}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\arrowarrow}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-{Straight Barb[length=2.5pt]}](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand\stackedarrows{%
        \mathrel{\vcenter{\mathsurround0pt
                \ialign{##\crcr
                \noalign{\nointerlineskip}$\arrowtail$\crcr
                \noalign{\nointerlineskip}$\tailarrow$\crcr
                }
        }}%
}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Causal Discovery on Precarity and Depression},
  pdfauthor={Kyuri Park; Leonie K. Elsenburg; Mary Nicolao; Karien Stronks; Vítor V. Vasconcelos},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Causal Discovery on Precarity and Depression}

\author[1]{Kyuri Park}
\author[2]{Leonie K. Elsenburg}
\author[2]{Mary Nicolao}
\author[2]{Karien Stronks}
\author[1, 3]{Vítor V. Vasconcelos}

\affil[1]{\textit{Computational Science Lab, Informatics Institute, University of Amsterdam, PO Box 94323, Amsterdam, 1090GH, the Netherlands}}
\affil[2]{\textit{Department of Public and Occupational Health, Amsterdam Public Health Research Institute, Amsterdam UMC, University of Amsterdam, Amsterdam, the Netherland}}
\affil[3]{\textit{Institute for Advanced Study, University of Amsterdam, Oude Turfmarkt 147, Amsterdam, 1012GC, the Netherland}}

\date{2025-02-11}
\begin{document}
\maketitle
\begin{abstract}
\noindent Understanding the causal mechanisms linking precariousness and
depression is critical for developing effective interventions. This
study utilizes data from the HELIUS cohort study to explore these
relationships using advanced causal discovery methods. By applying
algorithms such as FCI, and CCI, and combining traditional Gaussian CI
tests with non-parametric approaches like RCoT, we investigate how
different indicators of precariousness---including factors related to
employment, social relations and relational stress, financial situation,
and housing---affect depressed mood, both as a sum score and at the
individual symptom level. Our findings reveal that relational stress
consistently emerges as a potential causal factor for depressed mood,
while symptoms such as sleep disturbances, guilt, and anhedonia are
particularly sensitive to external stressors, acting as potential early
warning signals or intervention points for prevention. Moreover, the
results highlight complexities in the data, including the influence of
latent confounders and the challenges of capturing cyclic relationships.
Despite some limitations, such as unresolved ambiguities in causal
directions and challenges with mixed data distributions, this study
demonstrates the utility of causal discovery tools in disentangling the
intricate interplay between social and mental health dynamics. By
mapping these causal structures into computational models, future
research can simulate intervention effects, providing actionable
insights to mitigate the impact of precariousness on mental health. This
study serves as a foundational effort, offering both methodological
advancements and practical implications for addressing depression at a
population level.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}

\section{Introduction}\label{introduction}

Mental health problems in urban areas are on the rise, yet the
complexity of mental health systems poses significant challenges in
understanding the underlying mechanisms driving these issues --- let
alone designing effective interventions. Research has long sought to
identify underlying factors contributing to mental health problems.
Recent research looked into the association between mental health and
indicators of precariousness in various dimensions of life, indicating a
high level of uncertainty and instability in people's lives, such as
factors related to employment, social relations, finances, housing, and
culture (ref Leonie's paper). This comprehensive perspective on
precariousness helped to highlight how different aspects of life may be
interconnected. While this research has advanced our understanding of
which indicators of precariousness are related to mental health, a key
question remains unanswered: how do these factors influence mental
health? Specifically, the lack of directional insight --- knowing what
influences what --- limits our ability to identify and prioritize
effective intervention targets.

This study aims to investigate the causal relationships between
different dimensions of precariousness and mental health outcomes, to
with a specific focus on depression, the most prevalent mental health
issue. Using causal discovery methods, we explore how various aspects of
precariousness contribute to depression and whether depression, in turn,
exacerbates precariousness, potentially creating a feedback loop.
Furthermore, we zoom in at the symptom level, examining individual
depressive symptoms to identify those that may act as initiators ---
triggering the activation of the symptom network --- by displaying
heightened sensitivity to specific precariousness factors. We also
investigate symptoms that may function as bridging elements, linking the
dynamics of depression with precariousness by feeding back into
precarious conditions, potentially reinforcing the cycle of mental
health deterioration and precarious conditions. Through this analysis,
we aim to uncover the causal mechanisms driving mental health challenges
related to precariousness and establish a foundation for more targeted
and effective interventions by identifying the key driving forces within
this system.

\section{Methods}\label{methods}

\subsection{Data}\label{data}

We use data from the HELIUS study, a multi-ethnic cohort that includes
participants of Dutch, Turkish, Moroccan, Surinamese and Ghanaian origin
(Snijder et al., 2017). To operationalize precariousness factors, we
draw on the framework outlined in previous research (i.e., Leonie's
paper) and select a set of relevant variables.

To ensure a robust representation of each precariousness factor, we
conducted various exploratory analyses to identify consistent and
meaningful factor structures. Based on these analyses, we identified
five precariousness factors, including two related to recent stressors,
each comprising multiple variables as outlined below. Detailed
information on the exploratory analyses can be found in the
\hyperref[sec-appendix]{Appendix}.

\begin{itemize}
\tightlist
\item
  Employment precariousness: \texttt{emp\_stat}, \texttt{work\_sit}.
\item
  Social precariousness: \texttt{soc\_freq}, \texttt{soc\_adq}.
\item
  Housing precariousness: \texttt{nb\_safe}, \texttt{nb\_res},
  \texttt{nb\_rent}, \texttt{cul\_rec}.
\item
  Recent relational stressors: \texttt{frd\_brk12}, \texttt{conf12}.
\item
  Recent financial stressors: \texttt{fincri12}, \texttt{inc\_diff}.
\end{itemize}

After preprocessing, the HELIUS dataset comprises 21,628 samples. Along
with the five precariousness factors, we also compute an overall
precarity score as the combined value of these factors. Additionally,
PHQ-9 scores are included to represent depression, both as a total sum
score and as individual symptom scores. In the subsequent causal
discovery analysis, we examine the relationship between depression and
precarity using both aggregated sum scores and their individual-level
representations. Refer to Figure~\ref{fig-dist} for the overall
distributions of the variables used in the analysis.

\begin{figure}

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{draft_v1_files/figure-pdf/fig-dist-1.pdf}

}

\caption{\label{fig-dist}Distributions of variables with density
overlay. \emph{P.emp} = employment precariousness; \emph{P.hou} =
housing precariousness; \emph{P.soc} = social precariousness;
\emph{S.fin} = recent financial stressors; \emph{S.rel} = recent
relational stressors; \emph{Precarity} = overall precarity;
\emph{PHQsum} = PHQ-9 sum score; \emph{anh} = anhedonia; \emph{app} =
appetite; \emph{con} = concentration; \emph{dep} = depressed mood;
\emph{ene} = energy; \emph{glt} = guilty; \emph{mot} = motor; \emph{sui}
= suicidal}

\end{figure}%

\subsection{Causal Discovery}\label{causal-discovery}

There are numerous causal discovery algorithms available; however, in
this study, we focus on algorithms suited to the potential cyclic
relationships within our system. Specifically, we use \emph{FCI} (Fast
Causal Inference) and \emph{CCI} (Cyclic Causal Inference), both capable
of accounting for such cycles (Mooij \& Claassen, 2020; Strobl, 2019).
Additionally, we include the \emph{PC} algorithm as a reference, given
its simplicity and prominence as one of the most widely known causal
discovery methods (Spirtes et al., 2001).

\begin{table}[ht]
\centering
\small % Make the font size smaller
\caption{Assumptions of causal discovery algorithms}
\begin{tabular}{lccccc}
\toprule
Algorithm & Acyclicity & \makecell{Causal \\ sufficiency} & 
\makecell{Absence of \\ selection bias} & Linearity & Output \\ 
\midrule
PC  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & CPDAG \\ 
FCI & $-^{a}$ & $\checkmark$ & $-^{a}$ & $-^{a}$ & PAG \\ 
CCI & x & x & x & $\checkmark$ & \makecell{\footnotesize(partially oriented) \\ MAAG} \\ 
\bottomrule
\end{tabular}
\caption*{\footnotesize{\textit{Note}. $^{a}$The FCI algorithm, introduced by Spirtes (1995), is a constraint-based causal discovery method for DAGs that accounts for latent confounding and selection bias. Mooij \& Claassen (2020) later showed its applicability to cyclic causal discovery with latent confounding under general faithfulness and Markov conditions, assuming non-linear causal relationships.}}
\end{table}

As shown in Table 1, the resulting graphs from FCI and CCI differ
slightly (\emph{PAG}: partial ancestral graph; \emph{MAAG}: maximal
almost ancestral graph) due to their reliance on different underlying
assumptions. Despite these differences, both graphs belong to the class
of \emph{ancestral graphs}, which are designed to encode causal
relationships between variables, where the presence of an edge indicates
causal \emph{ancestry.} In these graphs, directed edges,
\(A \stararrow B\), indicate that \(B\) is not an ancestor of \(A\) in
every graph within the Markov equivalence class, \(Equiv(G)\). The
Markov equivalence class represents a set of graphs that encode the same
conditional independence relationships, ensuring that the same
\emph{d-separation} conditions hold across all graphs in the class
(Spirtes et al., 2001). Conversely, an edge marked as \(A \startail B\)
indicates that \(B\) is an ancestor of \(A\) across all graphs in
\(Equiv(G)\). Circle endpoints, \(A \starcirc B\), represent ambiguity
in the ancestral relationship, meaning \(B\)'s ancestral status relative
to \(A\) varies across graphs in \(Equiv(G)\).\footnote{\(*\) serves as
  a \textit{meta-symbol}, representing one of the three possible
  edge-endpoints. For instance, \(A \tailstar B\) can indicate any of
  the following edges: \(A\) \textemdash~\(B\), \(A\tailarrow B\), or
  \(A \tailcirc B\) (Park et al., 2024).} Finally, when an edge is
represented as \(A \arrowarrow B\), it implies that neither \(A\) nor
\(B\) is an ancestor of the other, suggesting the presence of a latent
confounder influencing both variables.

The algorithms also differ in how they detect cycles and represent
cyclic relationships in their resulting graphs. In CCI's MAAG, \(A\)
\textemdash~\(B\) indicates that \(A\) is an ancestor of \(B\), and
simultaneously \(B\) is an ancestor of \(A\), referring to a cyclic
relationship between \(A \stackedarrows B\). FCI, on the other hand,
identifies potential cycles more subtly. In its PAG, fully-connected
nodes with circle endpoints (\(\circirc\)) may suggest the presence of
cyclic structures. FCI, therefore, provides a sufficient condition to
distinguish variables that are not part of a cycle, offering a more
nuanced approach to handling cyclic relationships (Mooij \& Claassen,
2020).

Table 1 highlights further differences in the assumptions underlying
these algorithms. CCI operates under the assumption of a linear system,
whereas FCI, particularly when used to infer cyclic relationships,
assumes a non-linear system without selection bias and adheres to more
general faithfulness and Markov conditions (i.e.,
\emph{\(\sigma\)-separation} and \emph{\(\sigma\)-faithfulness} setting)
(Forré \& Mooij, 2018). When the respective assumptions of each
algorithm are met, their inferred cyclic relationships align with those
illustrated in Figure~\ref{fig-examplegraphs}.

While CCI's MAAG provides an explicit representation of cyclic
relationships, it has certain theoretical limitations, as the MAAG may
not always fully preserve d-separation relations from the original
graph, \(\mathcal{G}\) (Strobl, 2019). Given these considerations, we
examine and compare results from both algorithms, placing more weight on
the FCI results, which are presented in the main results section, while
discussing CCI results where relevant, with detailed findings provided
in the \hyperref[sec-appendix]{Appendix}. A full discussion of causal
discovery concepts and algorithmic details is beyond the scope of this
paper; however, readers seeking a more in-depth understanding can refer
to Park et al. (2024) for a comprehensive exploration of these methods
and their applications.

\begin{figure}

\begin{minipage}{0.33\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/original_DCG.png}

}

\subcaption{\label{fig-examplegraphs-1}Example graph \(\mathcal{G}\)}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/FCI_graph.png}

}

\subcaption{\label{fig-examplegraphs-2}Corresponding MAAG}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/CCI_graph.png}

}

\subcaption{\label{fig-examplegraphs-3}Corrsponding PAG}

\end{minipage}%

\caption{\label{fig-examplegraphs}Example graph \(\mathcal{G}\)
featuring a cycle and the corresponding MAAG from CCI and PAG from FCI.}

\end{figure}%

The graph produced by the PC algorithm is a \emph{CPDAG} (completed
partially directed acyclic graph), where directed edges (\(A\) → \(B\))
indicate that \(A\) is a direct cause (parent) of \(B\). Unlike FCI and
CCI, the CPDAG does not include circle symbols. Instead, when the PC
algorithm cannot determine the direction of causality, it represents
this uncertainty with bidirectional arrows. While the PC algorithm
serves as a useful reference, its strict assumptions --- acyclicity and
the absence of latent confounders---limit its applicability in more
complex scenarios. For this reason, our primary focus remains on the
results obtained from FCI and CCI, with all PC algorithm results
provided in the \hyperref[sec-appendix]{Appendix} for completeness.

One practical challenge in applying these algorithms to the HELIUS
dataset is that most variables do not follow a Gaussian distribution,
and the relationships among variables are unlikely to be strictly
linear. To address this, we supplement the commonly used Gaussian
conditional independence (CI) test, which relies on partial
correlations, with a non-parametric CI test based on reproducing kernel
methods (Zhang et al., 2012). While kernel-based conditional
independence tests (KCIT) are theoretically effective in capturing
non-linear dependencies and subtle interactions in data with
non-Gaussian noise, they are computationally demanding for large
datasets like HELIUS. Their quadratic scaling with sample size is
primarily due to the need to invert large kernel matrices, making them
impractical for large-scale analysis (Rahimi \& Recht, 2007). To
mitigate this issue, we use the Randomized Conditional Correlation Test
(RCoT), which approximates kernel methods using random Fourier features.
This approach reduces computational complexity from quadratic to linear
scaling with sample size, significantly lowering computational costs
while maintaining robustness (Strobl et al., 2019). For a more detailed
explanation of RCoT, refer to Section~\ref{sec-rcot}.

\subsection{Analysis}\label{sec-analysis}

We analyze the causal structure using three complementary approaches:
(1) Examining the relationship between the five individual
precariousness factors and the PHQ sum score, which represents overall
depression severity. (2) Investigating the relationship between
individual depressive symptoms and individual precariousness factors.
(3) Assessing the relationship between individual depressive symptoms
and overall precariousness (the sum of all precariousness factors).
Using the PHQ sum score as a measure of depression simplifies the
analysis by reducing dimensionality, improving computational efficiency,
and providing a broad, interpretable perspective on how precariousness
factors relate to overall depression severity. However, this aggregated
approach may overlook granular symptom-level relationships.

By zooming in on individual symptoms, we capture a more detailed picture
of how specific depressive symptoms interact with different
precariousness factors. This approach allows us to map individual causal
pathways, identifying which symptoms are particularly sensitive to
certain precarious conditions and vice versa. However, this level of
detail introduces methodological challenges, including non-standard
distributions of symptom variables and the high-dimensional nature of
the analysis, which can reduce statistical power and significantly
increase computational demands---especially when using non-parametric
conditional independence (CI) tests.

Finally, analyzing individual symptoms in relation to overall
precariousness provides an integrative perspective, revealing how
depression symptoms collectively relate to precarious conditions. This
approach complements the fully disaggregated analysis by aggregating
precariousness factors, potentially uncovering relationships that may be
too weak to detect when examined separately. This is particularly
relevant when a symptom is associated with multiple precariousness
factors, but these associations are not strong enough to emerge in an
individual-level analysis. By integrating these three approaches ---
ranging from partially aggregated to fully disaggregated --- we aim to
achieve a balance between capturing detailed granularity and gaining a
broader understanding of the relationship between precarity and
depression.

Furthermore, to address sensitivity to parameter values and enhance the
robustness of our results, we implement multiple settings combined with
bootstrapping. For each condition, we generate 100 bootstrap samples,
estimate causal graphs, and retain only the edge-endpoints that exceed
predefined thresholds. If no edge-endpoint type surpasses the threshold,
we assign a circle (\(\circ\)) to indicate uncertainty in the edge
direction.

The analyses are conducted under the following conditions:

\begin{itemize}
\tightlist
\item
  Significance levels (\(\alpha\)): 0.01 and 0.05
\item
  Thresholds: 0.5, 0.6, 0.7, and 0.8
\item
  CI test: Gaussian CI test, RCoT
\item
  Algorithms: FCI, CCI, and PC
\end{itemize}

This setup yields 16 combinations (2 significance levels × 4 thresholds
× 2 CI tests), applied across three algorithms. Each combination is
repeated for 100 bootstrap samples, yielding a total of 1,600 graphs per
algorithm. For analyses involving individual symptom variables, we
simplify the setup by focusing on thresholds of 0.6 and 0.7 and reducing
the number of bootstrap samples to 30, easing computational demands. To
further enhance efficiency, we fix the skeleton of the symptom network
using a common structure derived from PC, FCI, and CCI algorithms
(\(\alpha\) = 0.01, RCoT CI test). This step constrains which edges can
be present among symptoms but does not restrict the estimation of causal
directions within the symptom network. Additionally, the entire causal
structure between symptoms and precariousness factors remains fully
estimable, ensuring that key relationships are still inferred. Beyond
its computational benefits --- particularly in reducing the time
required for skeleton estimation --- this fixed structure also aligns
well with commonly reported symptom network structures in the literature
(Park et al., 2025), ensuring consistency with existing findings.

To obtain a final, stable summary of causal relationships, we identify
the most frequently occurring edge-endpoints across different
experimental setups. First, within each condition, we determine the most
frequent edge-endpoint from all bootstrap samples. Then, across
different conditions, we aggregate these selected symbols and determine
the overall most common edge-endpoint for each graph entry.

Since we have an even number of experimental setups, ties may
occasionally occur. In such cases, we explicitly denote them in the
resulting graph using dashed lines. To further help with the
interpretation and understanding of the resulting graph, we provide a
corresponding table for each graph, showing the proportion of each
edge-endpoint type to clarify how dominant a given symbol is in each
case.

By double summarizing the results in this way, we generate one final
graph per algorithm for each analysis approach: the partially aggregated
analysis, which uses either the sum of precariousness factors or the sum
of depression factors to capture overall trends, and the fully
disaggregated analysis, which examines individual precariousness and
depression factors separately. This approach ensures that only stable
and consistent edges --- those that persist across different parameter
settings --- are retained, leading to a robust and reliable
understanding of the relationships between precariousness factors and
depression. See Figure~\ref{fig-workflow} for an overview of the
analysis workflow.

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{img/simsetup2.png}

}

\caption{\label{fig-workflow}Analysis workflow applied across all three
algorithms.}

\end{figure}%

\section{Results}\label{results}

\subsection{Depression as sum score}\label{depression-as-sum-score}

The sum score graphs provide a high-level summary of how precarity
factors collectively influence overall depression severity, focusing on
aggregated relationships. Figure~\ref{fig-sum} illustrates the causal
relationships between precariousness factors (\emph{P.hou},
\emph{P.emp}, \emph{P.soc}, \emph{S.rel}, \emph{S.fin}) and the
depression sum score (\emph{PHQsum}), as identified by the FCI
algorithm. As described in Section~\ref{sec-analysis}, dashed edges
indicate ties, with the exact edge endpoint where the tie occurs is
represented by a bold horizontal bar.

The causal graph in panel (a) highlights key pathways in the
relationships between precarity factors and depression (PHQsum).
Employment precarity (P.emp) and social precarity (P.soc) are not
identified as causes of depression, whereas financial stress (S.fin)
appears to play a causal role. While P.emp and P.soc are generally not
recognized as causes of other precariousness factors, S.fin emerges as a
potential cause, as indicated by its circle edge endpoint. Supporting
this, the proportion matrix plot in panel (b) shows that S.fin has some
probability of causally influencing either P.emp or P.soc.

The matrix plot can be interpreted such that the symbol in
matrix{[}\(i\), \(j\){]} represents the relationship
\(i -~[symbol] \  j\). For example, if matrix{[}\(i\), \(j\){]} is
\textgreater{} and matrix{[}j, i{]} is \(\circ\), then the inferred
relationship is \(i \circarrow j\). In the table, different edge types
are represented by distinct colors: light gray for the absence of an
edge, blue for circles, green for arrowheads, and coral for arrowtails.
The colors are blended based on the proportion of each edge type, with
higher proportions increasing the opacity of the corresponding color,
making dominant symbols more visually prominent.

Relational stress (S.rel) plays a more nuanced role, interacting with
depression and S.fin through a latent confounder. Its relationship with
P.soc remains less certain, as the proportion table suggests a fair
probability that this connection is absent, uncertain, or bidirectional.
Meanwhile, housing precarity (P.hou) is not directly causally related to
depression but is linked to P.emp and S.fin. While P.emp and S.fin are
identified as non-causes of P.hou, it remains unclear whether P.hou
causally influences P.emp or if their relationship is mediated by an
unobserved confounder.

The findings from the CCI algorithm (see Figure~\ref{fig-sum-cci}) are
largely consistent with those from the FCI algorithm, with one
exception: CCI does not identify S.fin as a cause of depression.
Additionally, CCI introduces greater uncertainty in edge directions,
with more ties mainly between circle and arrowhead endpoints, and tends
to favor arrowheads more frequently than FCI. Despite these differences,
the skeleton structure and overall causal directions derived from CCI
align well with the results of the FCI algorithm, supporting the key
pathways.

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/FCI_depsum.png}

}

\subcaption{\label{fig-sum-1}FCI PAG}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/depsum_mat_fci.pdf}

}

\subcaption{\label{fig-sum-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sum}Resulting graph of precarity factors and
depression sum score using FCI and proportion of edge endpoint types.}

\end{figure}%

\subsection{Individual depression
symptom}\label{individual-depression-symptom}

Moving from the sum score representation to the disaggregated
symptom-level graph provides a more granular perspective on the causal
relationships between precariousness factors and depression symptoms.
Unlike the sum score graph, which aggregates all symptoms into a single
measure---potentially obscuring nuanced relationships---the
symptom-level graph highlights heterogeneity in how precariousness
factors influence individual depressive symptoms: con (concentration),
slp (sleep), ene (energy), app (appetite), mot (motor), sui (suicidal),
anh (anhedonia), glt (guilt), and dep (depressed mood), and vice versa.

The symptom-level graph in Figure~\ref{fig-sym} panel (a) reveals a more
complex and interconnected structure than the sum score graph,
reflecting the strong interdependence among symptoms and suggesting much
presence of latent confounders, as indicated by numerous bidirectional
edges. As before, dashed edges denote ties, with specific ties marked by
bold horizontal bars. In Figure~\ref{fig-sym} panel (b), the proportion
matrix provides more insight into these directionalities. Here, opaque
green dominates most symptom-to-symptom connections, indicating that
arrowheads are the most frequently inferred edge type. However, blue
regions, which correspond to circle endpoints, are particularly common
for anh, slp, ene, and sui, reflecting uncertainty in the causal
direction among these symptoms. Notably, the only arrowtail connection
appears between dep and anh, suggesting that anhedonia (anh) is a
potential cause of depressed mood (dep).

Looking at the symptom-precariousness connections, one of the key
patterns is that ties are most frequently found in the relationships
between individual symptoms and precariousness factors. This suggests
that these causal links may be less stable across different conditions.
Upon closer examination, most ties arise due to discrepancies between
the Gaussian CI test and the RCoT test---with Gaussian CI favoring
arrowheads and RCoT more often predicting the absence of an edge.

This discrepancy likely arises from a combination of factors. First, as
the analysis shifts from aggregated to disaggregated variables, the
number of conditional independence (CI) tests increases, leading to a
loss of statistical power in high-dimensional causal discovery. With
more variables being conditioned on, the data becomes sparser, making it
harder to detect weaker dependencies. Additionally,
over-conditioning---controlling for too many variables---can
artificially remove statistical associations, leading to false
negatives. Beyond this general statistical power issue, the fundamental
differences in how these CI tests operate contribute to the
inconsistency. RCoT, as a nonparametric test, does not assume linearity
and can capture both linear and nonlinear dependencies. However, it is
more conservative and requires larger sample sizes, making it more prone
to false negatives in high-dimensional settings. In contrast, partial
correlation (Gaussian CI test) assumes linear Gaussian relationships,
making it more permissive and sometimes detecting weak dependencies that
may not be statistically meaningful. As a result, when Gaussian CI
detects an edge while RCoT does not, it could indicate that the
relationship is weak, strictly linear, or requires more data for
reliable nonparametric detection. This explains why many ties in the
proportion matrix show a 50-50 split between symbols---where half of the
conditions with RCoT suggest the absence of an edge, while the other
half using Gaussian CI test favor an arrowhead. This systematic
divergence highlights the challenges of causal discovery in
high-dimensional settings, where varying statistical assumptions can
lead to differing conclusions about the presence and direction of causal
relationships.

Despite these differences, some consistent patterns emerge across both
CI tests, particularly in the case of financial stress (S.fin), which
appears to be connected to nearly all other variables---though many of
these connections are marked by ties, reflecting uncertainty in
directionality. One exception is the stronger evidence suggesting that
S.fin may cause changes in app (appetite), while its relationships with
other symptoms, such as ene, mot, and dep, remain ambiguous. Similarly,
social precarity (P.soc) exhibits numerous connections with depressive
symptoms, with most edges pointing toward P.soc rather than outward from
it. This suggests that depressive symptoms, particularly dep and glt,
may contribute to worsening social precarity rather than the other way
around. Additionally, anh also shows some probability of causally
influencing P.soc, reinforcing the idea that social precarity is more
often a consequence rather than a driver of depressive symptoms.

Other precariousness factors exhibit more uncertain but still notable
relationships. Relational stress (S.rel) shows weak but existing
connections with slp, glt, and app, though directionality remains
unclear in many cases. Employment precarity (P.emp) also connects to
symptoms like slp, mot, and con, but, like S.rel, these relationships
exhibit a 50/50 split in directionality, reflecting uncertainty in the
inferred causal paths. In line with the aggregated graph in
Figure~\ref{fig-sum}, housing precarity (P.hou) does not appear to have
any direct relationship with depression symptoms but consistently shows
associations with employment precarity (P.emp) and financial stress
(S.fin). The causal relationships among precariousness factors remain
largely unchanged from the aggregated analysis, with financial stress
(S.fin) exhibiting the strongest tendency to influence other
precariousness factors.

Among depressive symptoms, dep, glt, and slp appear to be the most
connected to precariousness factors, while P.soc has the most
connections with symptoms, predominantly as a recipient rather than a
driver of influence. Within the symptom network, slp, sui, and anh
emerge as causally influential symptoms, as they exhibit more outgoing
arrows compared to other symptoms. On the other hand, con, mot, dep, and
glt, despite having high connectivity, predominantly receive incoming
arrows, indicating they are more likely effects rather than causes.
Considering both symptom-precariousness connections and symptom-level
dynamics, sleep disturbance (slp) emerges as a central symptom, given
its strong ties to precariousness factors and its influential role
within the symptom network. This suggests that sleep issues could be an
initiating symptom, particularly sensitive to relational stress and
employment precarity. On the other hand, social precarity (P.soc) serves
as a key bridge between the depression and precariousness subsystems, as
depressive symptoms appear to feed back into social precarity,
reinforcing a self-sustaining dynamic between depression and precarious
conditions.

Finally, the results from CCI are largely consistent with those from
FCI, but with some key differences. CCI tends to favor arrowheads more
frequently, resulting in a greater number of bidirectional edges, which
suggests a higher involvement of latent confounders. Additionally, CCI
exhibits more tie situations, though, similar to FCI, most ties occur
between absence and arrowhead edges, reflecting the discrepancy between
the Gaussian CI test and RCoT---where the Gaussian test favors
arrowheads, while RCoT more often suggests the absence of an edge. For a
detailed visualization of the CCI-derived graph and proportion matrix,
see Figure~\ref{fig-sym-cci}.

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{img/symptom_graph_FCI.png}

}

\subcaption{\label{fig-sym-1}FCI PAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{img/symptom_mat_fci.pdf}

}

\subcaption{\label{fig-sym-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sym}Resulting graph of precarity factors and
individual depression symptoms using FCI and proportion of edge endpoint
types.}

\end{figure}%

\subsection{Precarity as sum score}\label{precarity-as-sum-score}

Lastly, we examine the relationships between individual symptoms and
overall precarity, an aggregated measure represented by the sum of five
precariousness factors. This analysis provides a complementary
perspective, capturing broader patterns that may not be evident in the
disaggregated symptom-precariousness analysis. By aggregating
precariousness factors into a single score, this approach may capture
distributed relationships across different precarity factors that were
previously overlooked in the disaggregated analysis.

Unlike the disaggregated graph in Figure~\ref{fig-sym}, the aggregated
symptom-precariousness graph (Figure~\ref{fig-presum} panel (a)) shows
more certainty in causal directions, with only a few edges resulting in
ties, all of which involve the overall precarity factor. This is even
more evident in the proportion matrix (Figure~\ref{fig-presum} panel
(b)), where symptom-to-symptom interactions almost fully converge to a
single edge type. Additionally, all symptom interactions are connected
either through bidirectional edges or a combination of an arrowhead and
a circle, suggesting a strong presence of latent confounders.

Examining the symptom-precarity connections, we observe a stronger
overall trend of depressive symptoms influencing precarity, rather than
the other way around. Specifically, anh, dep, and slp show a high
probability of causing precarity, whereas app, glt, and mot exhibit more
uncertainty in directionality, often resulting in circle endpoints. This
pattern suggests that when precariousness factors are aggregated, the
dominant causal flow is from depressive symptoms to precarity, rather
than precarity driving depression. Particularly, dep emerges as the
strongest predictor of precarity, exhibiting the highest proportion of
arrowtails, consistent with findings from the disaggregated analysis.

Comparing this with the CCI-derived graph (Figure~\ref{fig-presum-cci}),
several key differences stand out. The CCI results reveal that most
symptoms are interconnected through bidirectional edges, suggesting a
strong presence of latent confounders. However, dep appears to be a
distinct exception, as it is predominantly caused by nearly all other
symptoms, with one exception sui, which does not contribute to dep.
Similarly, con is influenced by multiple symptoms, albeit with weaker
support compared to dep. Most interestingly, the CCI results highlight a
distinct feedback loop between dep and overall precarity, represented by
a tail-tail edge (\textemdash ). This suggests a possible reinforcing
cycle between depression and overall precariousness, where depression
not only arises from precarious conditions but also contributes to their
persistence, creating a self-sustaining dynamic.

Overall, the aggregated precariousness graph supports the key findings
from both the disaggregated analysis and the depression sum score
analysis, providing greater certainty in several important patterns. The
results highlight the significant role of latent confounders in
symptom-to-symptom interactions. While certain stressor-related
precariousness factors, particularly S.fin and S.rel, appear to causally
influence depressive symptoms, other precariousness factors are more
likely driven by depression rather than acting as primary causes,
suggesting that precariousness may function more as a consequence of
depression than its source. Some symptoms, particularly glt, slp, and
dep, show greater sensitivity to precarious conditions, while also
playing a central role in symptom dynamics. Among them, slp and glt
emerge as influential symptoms within the depression network, whereas
con and dep, despite their high connectivity, behave more as outcomes
rather than primary drivers. Together, these findings suggest that
depression---especially symptoms like sleep disturbances and guilt---may
contribute to sustaining precarious conditions, possibly forming a
reciprocal cycle between depression and precariousness.

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_graph_FCI.png}

}

\subcaption{\label{fig-presum-1}FCI PAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_mat_fci.pdf}

}

\subcaption{\label{fig-presum-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-presum}Resulting graph of precarity sum score and
individual depression symptoms using FCI and proportion of edge endpoint
types.}

\end{figure}%

\section{Discussion}\label{discussion}

This study examined the causal relationships between precariousness
factors and depression using causal discovery algorithms, integrating
both sum score and symptom-level analyses. By analyzing PHQ sum scores
alongside individual symptom scores, and comparing aggregated precarity
scores with individual precariousness factors, we gained a comprehensive
perspective on how different aspects of precariousness influence
depression. This dual approach revealed nuanced pathways that may be
overlooked when relying solely on aggregate measures, while also
capturing broader interactions that might remain undetected in analyses
focused on individual factors.

Our findings highlight the significant role of recent financial stress
(S.fin) as a potential causal factor for depression, consistently
observed across both sum score and symptom-level analyses. This
consistency underscores the profound impact of financial stressors on
mental health. The symptom-level analysis further identified specific
symptoms---particularly sleep disturbances (slp), depressed mood (dep),
guilt (glt), and anhedonia (anh)---as being especially sensitive to
external precarious conditions.

Among these symptoms, dep stands out as a distinct causal factor for
precarity in both aggregated and disaggregated analyses, with CCI even
suggesting a possible cyclic relationship between them. This implies
that depression and precariousness may reinforce one another,
potentially leading to a worsening cycle over time. Within the symptom
network, dep also exhibits a high degree of connectivity, but with
mostly incoming arrows, suggesting that as symptom dynamics intensify,
dep is quickly impacted by multiple factors---particularly P.soc (social
precarity), further fueling the depression-precariousness cycle.

Beyond dep, slp emerges as a critical symptom in this dynamic. It
connects with both social and employment precariousness, while also
playing a central role in symptom interactions, affecting other symptoms
more than most others. This suggests that slp may act as an initiator
within the depressive symptom system, potentially triggering or
amplifying symptom cascades.

Among the precariousness factors, P.soc appears to be one of the most
pivotal---not as a driver of symptoms, but rather as a consequence of
depression. It maintains the strongest connectivity with symptoms, yet
its directionality indicates that it is largely shaped by depressive
symptoms rather than acting as a primary cause. Together, these patterns
suggest a broader directional dynamic in which financial stress (S.fin)
exacerbates depressive symptoms, which in turn feed into social
precarity, perpetuating the cycle.

This directional insight suggests that interventions targeting specific
depressive symptoms, particularly dep, may help prevent downstream
effects on worsening precarious conditions. Alternatively, early
intervention focusing on slp may be a promising strategy, as slp appears
to act as a potential initiator or activator within the depressive
symptom network. This positioning suggests that slp could serve as an
early warning sign, offering a valuable target for preventive
interventions aimed at breaking the depression-precariousness cycle
before it intensifies.

Several limitations should be acknowledged when interpreting these
findings. The prevalence of bidirectional arrows and circle-marked
endpoints in the graphs reflects unresolved ambiguities in the causal
relationships suggested by the data. These uncertainties highlight the
need for further research, particularly with higher-resolution datasets,
such as time-series data. Incorporating time-series data could capture
temporal dynamics between precarity factors and depressive symptoms,
providing time-specific insights and revealing how these relationships
evolve over time. Methods like \emph{PCMCI} (Runge et al., 2019),
\emph{tsFCI} (Entner \& Hoyer, 2010), and other time-series adaptations
of causal discovery algorithms could offer a more robust framework for
addressing temporal dependencies and refining the analysis.

Another notable limitation is the absence of clear evidence for cycles
within the symptom network, despite employing algorithms designed to
account for cyclic relationships. The CCI algorithm primarily produced
bidirectional arrows, while FCI predominantly generated directional
arrows, yet neither demonstrated patterns indicative of definitive
cyclic structures. This suggests that addressing cyclic relationships is
particularly challenging when relying solely on observational datasets.
Future research could benefit from refined datasets that incorporate
intervention data. Methods such as \emph{LLC} (Hyttinen et al., 2012),
\emph{NODGAS-Flow }(Sethuraman et al., 2023), and the recently developed
\emph{Bicycle} algorithm (Rohbeck et al., 2024), which utilize both
observational and intervention data, may uncover potential cycles more
effectively.

Lastly, regarding conditional independence (CI) testing, the differences
between the graphs generated by Gaussian CI and RCoT underscore the
methodological sensitivities in detecting causal relationships. Gaussian
CI produced denser graphs, which is somewhat counterintuitive, as the
Gaussian CI's strict linearity assumption would typically result in
fewer detected relationships, not more. A possible explanation for this
discrepancy is that Gaussian CI's reliance on partial correlations may
overestimate relationships when specific non-linear dependencies exist
in the data. In contrast, RCoT, free from such assumptions, might better
capture independence patterns under such conditions. However, RCoT is
not without its limitations. While technically non-parametric, its
performance can be influenced by the distributional characteristics of
the data. The Gaussian RBF (radial basis function) kernel, optimized for
continuous data with smooth transitions, may struggle with discrete or
mixed datasets, where distances between discrete points may fail to
convey meaningful similarity (Howlett, 2001). This limitation could
explain why RCoT might miss certain dependencies, particularly in the
HELIUS dataset, where some variables lack smooth continuity. Future
research could tackle these challenges by developing hybrid kernels
designed for mixed datasets, integrating both continuous and discrete
variables into RCoT-like methods. This would have the potential to
improve the reliability of findings, especially when working with
datasets characterized by complex and heterogeneous structures.

Despite its limitations, this study marks a meaningful step toward
understanding the mechanisms linking precarity factors and depression.
By applying causal discovery methods, it moves beyond traditional
association-based analyses, providing insights that can inform more
precise and targeted interventions. While the resulting graphs are
preliminary and contain unresolved ambiguities, they offer a valuable
starting point for leveraging causal discovery tools to investigate the
causal interplay between depression and precarity factors. A promising
next step would involve integrating these causal structures into
computational models, such as the symptom dynamic model proposed in Park
et al. (2025). By simulating intervention effects, such models could
provide more realistic insights into how targeted actions might
influence symptom networks and precarity factors over time. For example,
interventions focused on improving sleep hygiene or alleviating guilt
could be evaluated for their cascading effects on employment and social
relationships, offering actionable guidance for designing
population-level mental health strategies. As one of the early
applications of causal discovery tools to the complex dynamics of
depression and precarity factors, this study lays a foundation for
future research. We hope it inspires further refinement of these methods
and ultimately contribute to more effective solutions for alleviating
depression and improving societal well-being.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-dojer2016learning}
Dojer, N. (2016). Learning bayesian networks from datasets joining
continuous and discrete variables. \emph{International Journal of
Approximate Reasoning}, \emph{78}, 116--124.

\bibitem[\citeproctext]{ref-entner2010causal}
Entner, D., \& Hoyer, P. O. (2010). On causal discovery from time series
data using FCI. \emph{Probabilistic Graphical Models}, \emph{16}.

\bibitem[\citeproctext]{ref-forre2018constraint}
Forré, P., \& Mooij, J. M. (2018). Constraint-based causal discovery for
non-linear structural causal models with cycles and latent confounders.
\emph{arXiv Preprint arXiv:1807.03024}.

\bibitem[\citeproctext]{ref-howlett2001radial}
Howlett, R. J. (2001). \emph{Radial basis function networks 1: Recent
developments in theory and applications}.

\bibitem[\citeproctext]{ref-hyttinen2012learning}
Hyttinen, A., Eberhardt, F., \& Hoyer, P. O. (2012). Learning linear
cyclic causal models with latent variables. \emph{The Journal of Machine
Learning Research}, \emph{13}(1), 3387--3439.

\bibitem[\citeproctext]{ref-lindsay2000moment}
Lindsay, B. G., Pilla, R. S., \& Basak, P. (2000). Moment-based
approximations of distributions using mixtures: Theory and applications.
\emph{Annals of the Institute of Statistical Mathematics}, \emph{52},
215--230.

\bibitem[\citeproctext]{ref-mooij2020constraint}
Mooij, J. M., \& Claassen, T. (2020). Constraint-based causal discovery
using partial ancestral graphs in the presence of cycles.
\emph{Conference on Uncertainty in Artificial Intelligence}, 1159--1168.

\bibitem[\citeproctext]{ref-neapolitan2004learning}
Neapolitan, R. E. et al. (2004). \emph{Learning bayesian networks} (Vol.
38). Pearson Prentice Hall Upper Saddle River.

\bibitem[\citeproctext]{ref-park2024discovering}
Park, K., Waldorp, L. J., \& Ryan, O. (2024). Discovering cyclic causal
models in psychological research. \emph{Advances. In/Psychology},
\emph{2}, e72425.

\bibitem[\citeproctext]{ref-park2025individual}
Park, K., Waldorp, L., \& Vasconcelos, V. V. (2025). \emph{The
individual-and population-level mechanistic implications of statistical
networks of symptoms}.

\bibitem[\citeproctext]{ref-rahimi2007random}
Rahimi, A., \& Recht, B. (2007). Random features for large-scale kernel
machines. \emph{Advances in Neural Information Processing Systems},
\emph{20}.

\bibitem[\citeproctext]{ref-rohbeck2024}
Rohbeck, M., Clarke, B., Mikulik, K., Pettet, A., Stegle, O., \&
Ueltzhöffer, K. (2024). Bicycle: Intervention-based causal discovery
with cycles. In F. Locatello \& V. Didelez (Eds.), \emph{Proceedings of
the third conference on causal learning and reasoning} (Vol. 236, pp.
209--242). PMLR.

\bibitem[\citeproctext]{ref-runge2019detecting}
Runge, J., Nowack, P., Kretschmer, M., Flaxman, S., \& Sejdinovic, D.
(2019). Detecting and quantifying causal associations in large nonlinear
time series datasets. \emph{Science Advances}, \emph{5}(11), eaau4996.

\bibitem[\citeproctext]{ref-sethuraman2023nodags}
Sethuraman, M. G., Lopez, R., Mohan, R., Fekri, F., Biancalani, T., \&
Hütter, J.-C. (2023). NODAGS-flow: Nonlinear cyclic causal structure
learning. \emph{International Conference on Artificial Intelligence and
Statistics}, 6371--6387.

\bibitem[\citeproctext]{ref-snijder2017cohort}
Snijder, M. B., Galenkamp, H., Prins, M., Derks, E. M., Peters, R. J.,
Zwinderman, A. H., \& Stronks, K. (2017). Cohort profile: The healthy
life in an urban setting (HELIUS) study in amsterdam, the netherlands.
\emph{BMJ Open}, \emph{7}(12), e017873.

\bibitem[\citeproctext]{ref-spirtes2001causation}
Spirtes, P., Glymour, C., \& Scheines, R. (2001). \emph{Causation,
prediction, and search}. MIT press.

\bibitem[\citeproctext]{ref-spirtes_causal_1995}
Spirtes, P., Meek, C., \& Richardson, T. (1995). Causal inference in the
presence of latent variables and selection bias. \emph{Proceedings of
the {Eleventh} Conference on {Uncertainty} in Artificial Intelligence},
499--506.

\bibitem[\citeproctext]{ref-strobl2019}
Strobl, E. V. (2019). A constraint-based algorithm for causal discovery
with cycles, latent variables and selection bias. \emph{International
Journal of Data Science and Analytics}, \emph{8}(1), 33--56.
\url{https://doi.org/10.1007/s41060-018-0158-2}

\bibitem[\citeproctext]{ref-strobl2019approximate}
Strobl, E. V., Zhang, K., \& Visweswaran, S. (2019). Approximate
kernel-based conditional independence tests for fast non-parametric
causal discovery. \emph{Journal of Causal Inference}, \emph{7}(1),
20180017.

\bibitem[\citeproctext]{ref-zhang2012kernel}
Zhang, K., Peters, J., Janzing, D., \& Schölkopf, B. (2012).
Kernel-based conditional independence test and application in causal
discovery. \emph{arXiv Preprint arXiv:1202.3775}.

\end{CSLReferences}

\section{Appendix}\label{sec-appendix}

\subsection{Precariousness factors by
Leonie}\label{precariousness-factors-by-leonie}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  EMPLOYMENT PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_Arbeidsparticipatie}: Working status
\item
  \texttt{H1\_WerkSit}: Which work situation most applies to you?
\item
  \texttt{H1\_RecentErv8}: Experiences past 12 months: h. You were
  sacked from your job or became unemployed (\emph{reverse})
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  FINANCIAL PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_InkHhMoeite}: During the past year, did you have problems
  managing your household income?
\item
  \texttt{H1\_RecentErv9}: Experiences past 12 months: i. You had a
  major financial crisis (\emph{reverse})
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  HOUSING PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{veilig\_2012}: Score safety (veiligheid) in 2012
  (\emph{reverse})
\item
  \texttt{vrz\_2012}: Score level of resources (niveau voorzieningen) in
  2012 (\emph{reverse})
\item
  \texttt{P\_HUURWON}: Percentage Huurwoningen
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  CULTURAL PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_Discr\_sumscore}: Perceived discrimination: sum score of 9
  items (range 9-45)
\item
  \texttt{H1\_SBSQ\_meanscore}: Health literacy: SBSQ meanscore (range
  1-5) (\emph{reverse})
\item
  \texttt{A\_BED\_RU}: Aantal bedrijfsvestigingen; cultuur, recreatie,
  overige diensten (\emph{reverse})
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  SOCIAL PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_RecentErv5}: Experiences past 12 months: e. Your steady
  relationship ended (\emph{reverse})
\item
  \texttt{H1\_RecentErv6}: Experiences past 12 months: f.~A long-term
  friendship with a good friend or family member was broken off
  (\emph{reverse})
\item
  \texttt{H1\_RecentErv7}: Experiences past 12 months: g. You had a
  serious problem with a good friend or family member, or neighbour
  (\emph{reverse})
\item
  \texttt{H1\_SSQT}: SSQT (frequency of social contact): sum score of 5
  items (range 5-20) (\emph{reverse})
\item
  \texttt{H1\_SSQSa}: SSQS (adequacy of social contact): sum score of 5
  items, category 3 and 4 not combined (range 5-20) (\emph{reverse})
\end{itemize}

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-9-1.pdf}
\end{center}

\begin{itemize}
\item
  \textbf{High} Correlations: \texttt{emp\_stat} (employment status) and
  \texttt{work\_sit} (work situation) have a strong positive correlation
  of 0.82. This suggests that individuals with higher employment status
  tend to have more secure or favorable work situations.
  \texttt{soc\_freq} (social contact frequency) shows a strong positive
  correlation with \texttt{soc\_adq} (social adequacy) at 0.61. This
  indicates that individuals with more frequent social contact also tend
  to have higher perceived adequacy of social interactions.
\item
  \textbf{Moderate} Correlations: \texttt{nb\_safe} (neighborhood
  safety) and \texttt{nb\_res} (resources) have a moderate positive
  correlation of 0.39, suggesting that areas with higher safety also
  have better resources. \texttt{hea\_lit} (health literacy) has
  moderate correlations with \texttt{emp\_stat} (0.26) and
  \texttt{work\_sit} (0.25), which could mean that higher health
  literacy is associated with better employment situations.
  \texttt{frd\_brk12} (friendship breakups) and \texttt{conf12}
  (conflicts) have a notable correlation of 0.43, indicating a
  relationship between having conflicts and friendship losses.
\item
  \textbf{Low to Moderate} Correlations in Financial Precariousness:
  \texttt{inc\_dif} (income difficulties) has a moderate correlation
  with \texttt{fincri12} (financial crisis) at 0.49. This aligns with
  the expected relationship, where individuals who experience general
  income difficulties are more likely to report financial crises.
\item
  \textbf{Low} Correlations (0.1 - 0.2): Many variables, such as
  \texttt{discrim} (discrimination), \texttt{unemp12} (unemployment
  experience), and \texttt{rel\_end12} (relationship end), have low
  correlations with other variables, suggesting relatively independent
  relationships in the context of this dataset.
\end{itemize}

\subsection{Exploratory Factor Analysis
(EFA)}\label{exploratory-factor-analysis-efa}

\begin{center}
\includegraphics[width=0.55\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-10-1.pdf}
\end{center}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-11-1.pdf}
\end{center}

\subsubsection{Factor Loadings (Pattern
Matrix)}\label{factor-loadings-pattern-matrix}

\begin{itemize}
\tightlist
\item
  \textbf{MR1}: High loadings on \texttt{emp\_stat} and
  \texttt{work\_sit} suggest this factor captures \emph{employment}
  precariousness.
\item
  \textbf{MR2}: Strong loadings on \texttt{soc\_freq} and
  \texttt{soc\_adq} indicate \emph{social} precariousness.
\item
  \textbf{MR3}: Key items like \texttt{frd\_brk12}, \texttt{conf12}, and
  \texttt{fincri12}, suggest recent \emph{stressful events}.
\item
  \textbf{MR4}: High loadings on \texttt{nb\_res} and \texttt{cul\_rec}
  may reflect \emph{community resources} precariousness.
\item
  \textbf{MR5}: Variables \texttt{nb\_safe} and \texttt{nb\_rent} with
  high loadings indicate \emph{housing} precariousness.
\end{itemize}

\subsubsection{Variance Explained}\label{variance-explained}

The factors cumulatively explain \textbf{38\%} of the variance, with MR1
being the most influential factor. Each factor contributes a smaller
proportion to the total variance (MR1 at 12\%, MR2 at 9\%, etc.).

\subsubsection{Factor Intercorrelations}\label{factor-intercorrelations}

Factors are moderately correlated, especially between \emph{MR1 and
MR5}, and \emph{MR2 and MR3}. This indicates that while distinct, these
factors are related---reasonable in a complex socio-economic context.

\subsubsection{Model Fit Statistics}\label{model-fit-statistics}

RMSEA (0.071) suggest an acceptable fit. Tucker Lewis Index (0.802)
suggests moderate reliability for the model.

\subsubsection{Summary}\label{summary}

The 5-factor model appears interpretable and captures distinct
dimensions of precariousness: \emph{employment, social, stressors,
community resources, and housing precariousness}. Although the overall
fit and explained variance could be stronger, these factors offer
insights into the underlying structure of the data, highlighting key
areas of precariousness.

\subsection{PCA}\label{pca}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-12-1.pdf}
\end{center}

\begin{itemize}
\tightlist
\item
  Component Retention: The scree plot shows a clear ``elbow'' after the
  first component. This steep drop suggests that most variance is
  explained by the first component. After Dimension 5, the percentage of
  explained variance decreases slightly more gradually, indicating
  diminishing returns for adding more components. If we need to choose
  multiple components, retaining the first 5 components seems
  reasonable, as they capture most of the variance (cumulatively
  explaining about 54.7\% of the total variance).
\end{itemize}

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-13-1.pdf}
\end{center}

\subsubsection{Explained variance (contributions) of
variables}\label{explained-variance-contributions-of-variables}

It shows the importance of variables within each component.

\begin{itemize}
\item
  \textbf{Dim1}: High contributions are observed from
  \texttt{emp\_stat}, \texttt{work\_sit}, \texttt{inc\_dif}, and
  \texttt{fincri12}, suggesting that this dimension captures aspects of
  \emph{employment and financial} security.
\item
  \textbf{Dim2}: While \texttt{emp\_stat} and \texttt{work\_sit} overlap
  with Dim1, the strong contributions from \texttt{frd\_brk12} and
  \texttt{rel\_end12} indicate that this dimension captures a focus on
  \emph{recent relationship stressors}.
\item
  \textbf{Dim3}: \texttt{cul\_rec}, \texttt{nb\_res} have the highest
  contributions, indicating this dimension likely represents
  \emph{community and cultural} factors.
\item
  \textbf{Dim4}: \texttt{soc\_freq} and \texttt{soc\_adq} stand out in
  this dimension, suggesting an emphasis on \emph{social}
  precariousness.
\item
  \textbf{Dim5}: \texttt{nb\_safe} and \texttt{nb\_rent} are the top
  contributors, pointing to \emph{housing} security as key themes in
  this component.
\end{itemize}

\subsubsection{Cos² Values}\label{cosuxb2-values}

Cos² (squared cosine) values, or the quality of representation, show how
well each variable is represented by each dimension. where higher cos²
values (closer to 1) indicate better representation of a variable by a
component.

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-14-1.pdf}
\end{center}

\begin{itemize}
\item
  \textbf{Dim.1}: Variables \texttt{emp\_stat}, \texttt{work\_sit},
  \texttt{inc\_dif}, and \texttt{fincri12} show high cos² values,
  meaning that PC1 primarily captures variations in employment and
  financial difficulties. This component could represent
  \emph{employment \& finance} precariousness.
\item
  \textbf{Dim.2}: Variables \texttt{work\_sit}, \texttt{emp\_stat},
  \texttt{frd\_brk12}, and \texttt{conf12} are well-represented in this
  component, suggesting PC2 captures aspects of \emph{recent
  relationship stressors}.
\item
  \textbf{Dim.3}: Variables \texttt{nb\_res} and \texttt{cul\_rec} load
  strongly on PC3. This may represent community or cultural resources,
  indicating that this component is associated with \emph{neighborhood
  resources}.
\item
  \textbf{Dim.4}: This component has high cos² values for
  \texttt{nb\_res}, \texttt{cul\_rec}, \texttt{soc\_freq}, and
  \texttt{soc\_adq.} While \texttt{nb\_res} and \texttt{cul\_rec} are
  also prominent in PC3, PC4 uniquely captures nuanced differentiation
  in \emph{social} precariousness.
\item
  \textbf{Dim.5}: \texttt{nb\_safe} and \texttt{nb\_rent} are
  well-represented by PC5. This component might capture \emph{housing}
  precariousness.
\end{itemize}

\subsection{ICA}\label{ica}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-15-1.pdf}
\end{center}

\subsubsection{Dominant Variables per
Component:}\label{dominant-variables-per-component}

For each Independent Component (IC), we can identify variables with
\emph{high absolute} values in each column. These values indicate that
the IC captures a strong, independent signal associated with these
variables.

\begin{itemize}
\item
  \textbf{IC1}: \texttt{soc\_freq} and \texttt{soc\_adq} have strong
  negative loadings on this component, indicating that this component
  might represent \emph{social} precariousness.
\item
  \textbf{IC2}: \texttt{frd\_brk12}, \texttt{conf12},
  \texttt{rel\_end12}, \texttt{fincri12} and \texttt{unemp12} have the
  most substantial loadings on this component, all with negative signs.
  This might point to a \emph{recent relational or social stressor}
  component.
\item
  \textbf{IC3}: \texttt{nb\_res} and \texttt{cul\_rec} show notable
  negative loadings, pointing to a focus on \emph{community resource}
  precariousness.
\item
  \textbf{IC4}: High loadings for \texttt{nb\_safe}, \texttt{nb\_rent},
  \texttt{nb\_res}, \texttt{cul\_rec}, and \texttt{discrim} suggest a
  theme of \emph{housing and community-based} precariousness, reflecting
  both safety and social challenges within the neighborhood context.
\item
  \textbf{IC5}: \texttt{emp\_stat} and \texttt{work\_sit} both have
  strong negative loadings on this component, suggesting it captures
  \emph{employment} precariousness.
\end{itemize}

\subsection{Hierarchical clustering}\label{hierarchical-clustering}

\subsubsection{Using Euclidean distance}\label{using-euclidean-distance}

\begin{itemize}
\tightlist
\item
  Ward.D's method: Minimizes the variance within clusters, producing
  more compact and spherical clusters.
\item
  Single linkage: Groups clusters based on the minimum distance between
  points.
\item
  Complete linkage: Groups clusters based on the maximum distance
  between points.
\item
  Average linkage: Uses the average distance between all pairs of points
  in the two clusters.
\end{itemize}

\begin{figure}

\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-1.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-2.pdf}
\end{center}
\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-3.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-4.pdf}
\end{center}
\end{minipage}%

\end{figure}%

\paragraph{Consistent Groupings (Across All or Most
Methods)}\label{consistent-groupings-across-all-or-most-methods}

\begin{itemize}
\item
  \texttt{emp\_stat} and \texttt{work\_sit}: This pair consistently
  clusters together across all linkage methods, suggesting that they are
  closely related variables, likely capturing a similar aspect of the
  data (possibly employment status or employment-related information).
\item
  \texttt{nb\_safe}, \texttt{nb\_res}, and \texttt{nb\_rent}: These
  variables are often grouped closely in several methods (especially
  Ward.D, average, and complete linkage). This suggests a similarity or
  common theme among them, potentially related to neighborhood or
  housing precariousness.
\item
  \texttt{soc\_freq} and \texttt{soc\_adq}: These two variables
  frequently cluster together, indicating they likely measure aspects of
  social frequency and adequacy in similar ways. They appear together in
  Ward.D, average, and complete linkage.
\item
  \texttt{frd\_brk12} and \texttt{conf12}: These variables are often
  clustered closely (though they sometimes join with other variables
  like rel\_end12), suggesting they may capture aspects of relationship
  or social conflict. This pair appears in close proximity, especially
  in average and Ward.D.
\end{itemize}

\paragraph{Inconsistent Groupings (Variability Across
Methods)}\label{inconsistent-groupings-variability-across-methods}

\begin{itemize}
\item
  \texttt{hea\_lit}: This variable shows inconsistent clustering across
  methods. In Ward.D, it joins with \texttt{fincri12}, while in other
  methods, it's often more isolated or grouped with variables that do
  not appear similar. This may suggest that \texttt{hea\_lit} does not
  strongly correlate with other variables, or it has multidimensional
  aspects affecting its grouping across methods.
\item
  \texttt{discrim}: This variable also shows variable groupings. In
  Ward.D, it is grouped with \texttt{hea\_lit}, while in other methods
  (e.g., complete and single linkage), it clusters differently,
  sometimes on its own. This variability may indicate that
  \texttt{discrim} has weaker associations with the main clusters in the
  data or overlaps partially with multiple clusters.
\item
  Social and Financial Variables (\texttt{inc\_dif}, \texttt{fincri12},
  \texttt{unemp12}): These variables appear together in some methods
  (e.g., Ward.D clusters \texttt{fincri12} and \texttt{inc\_dif}), but
  in others, they are spread out. This inconsistency suggests that
  social and financial variables may not have strong or consistent ties
  across different methods, perhaps due to capturing different aspects
  of precariousness.
\end{itemize}

\paragraph{Summary}\label{summary-1}

The consistent clusters are likely capturing distinct thematic
dimensions of the data (e.g., employment, housing, social contact),
while the inconsistent variables may reflect multifaceted or weakly
correlated attributes that do not fit neatly into one cluster.

\subsubsection{Using Mutual Information}\label{using-mutual-information}

Using mutual information (MI) as a basis for hierarchical clustering
differs from using traditional distance measures (like Euclidean
distance) in a few key ways.

\begin{center}
\includegraphics[width=0.6\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-17-1.pdf}
\end{center}

\paragraph{Comparison to Euclidean Distance
Clustering}\label{comparison-to-euclidean-distance-clustering}

\begin{itemize}
\item
  \textbf{Housing and Community} Cluster: The variables
  \texttt{nb\_safe}, \texttt{nb\_res}, \texttt{nb\_rent}, and
  \texttt{cul\_rec} cluster together, indicating a strong association
  among housing-related and community-based factors. This suggests a
  shared theme of housing or community precariousness. This grouping is
  also observed in the Euclidean-based clustering, but it appears more
  tightly connected here, potentially due to the non-linear
  relationships highlighted by mutual information.
\item
  \textbf{Employment and Social Support} Cluster: \texttt{emp\_stat} and
  \texttt{work\_sit} form a cluster, linking employment status and work
  situation together as they did in Euclidean-based clustering. These
  remain closely associated regardless of the distance metric used.
  \texttt{soc\_freq} and \texttt{soc\_adq}, related to social contact
  frequency and adequacy, cluster nearby, indicating they have a
  stronger non-linear relationship with employment variables. This is a
  subtle difference as Euclidean distance might not capture this
  association as effectively.
\item
  \textbf{Financial Stressor} Cluster: \texttt{inc\_dif} and
  \texttt{fincri12}, representing income difficulties and recent
  financial crises, consistently cluster together in both approaches,
  showing a strong association, likely linear. However, mutual
  information-based clustering links these financial stressors with
  social support variables, suggesting that financial challenges may
  have complex dependencies with social support in this dataset.
\item
  \textbf{Relational Stressor} Cluster: \texttt{frd\_brk12},
  \texttt{conf12}, \texttt{discrim}, \texttt{hea\_lit},
  \texttt{unemp12}, and \texttt{rel\_end12} form a \emph{looser} cluster
  focused on social and relational stressors (e.g., friendship breakup,
  conflicts, and discrimination). Compared to Euclidean clustering,
  \texttt{discrim} and \texttt{hea\_lit} (health literacy) appear closer
  to relational stressors here, indicating that non-linear relationships
  might play a larger role in linking these variables.
\end{itemize}

\paragraph{Summary}\label{summary-2}

In conclusion, mutual information-based clustering provides an
alternative perspective that can reveal more intricate associations
between variables, especially for those with non-linear relationships.
Compared to Euclidean clustering, it shows a similar high-level
structure but emphasizes nuanced connections between variables,
particularly around social support, employment, and financial stress.

\subsection{Conclusions on Precariousness
factors}\label{conclusions-on-precariousness-factors}

Based on the consistent findings across multiple analyses, we decided to
exclude the variables \texttt{discrim}, \texttt{hea\_lit},
\texttt{umemp12}, and \texttt{rel\_end12}, as they do not clearly belong
to any specific precariousness factor nor exhibit strong associations
with depression (see the correlation table above). Therefore, we propose
retaining the following key precariousness factors:

\begin{itemize}
\tightlist
\item
  Employment Precariousness: \texttt{emp\_stat}, \texttt{work\_sit}
\item
  Social Precariousness: \texttt{soc\_freq}, \texttt{soc\_adq}
\item
  Housing Precariousness: \texttt{nb\_safe}, \texttt{nb\_res},
  \texttt{nb\_rent}, \texttt{cul\_rec}
\item
  Recent Relational Stressors: \texttt{frd\_brk12}, \texttt{conf12}
\item
  Recent Financial Stressors: \texttt{fincri12}, \texttt{inc\_diff}
\end{itemize}

We construct each precariousness factor by calculating the mean value of
the combined variables. Below, we present the updated correlation table
for the newly composed factors, along with the corresponding
distributions of all variables to be used in the causal discovery
analysis.

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-18-1.pdf}
\end{center}

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-19-1.pdf}
\end{center}

\subsection{Results from CCI algorithm}\label{sec-cci}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/CCI_depsum.png}

}

\subcaption{\label{fig-sum-cci-1}CCI MAAG}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/depsum_mat_cci.pdf}

}

\subcaption{\label{fig-sum-cci-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sum-cci}Resulting graph of precarity factors and
depression sum score using CCI and proportion of edge endpoint types.}

\end{figure}%

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/symptom_graph_CCI.png}

}

\subcaption{\label{fig-sym-cci-1}CCI MAAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/symptom_mat_cci.pdf}

}

\subcaption{\label{fig-sym-cci-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sym-cci}Resulting graph of precarity factors and
individual depression symptoms using CCI and proportion of edge endpoint
types.}

\end{figure}%

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_graph_CCI.png}

}

\subcaption{\label{fig-presum-cci-1}CCI MAAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_mat_cci.pdf}

}

\subcaption{\label{fig-presum-cci-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-presum-cci}Resulting graph of precarity sum score
and individual depression symptoms using CCI and proportion of edge
endpoint types.}

\end{figure}%

\clearpage

\subsection{Results from PC algorithm}\label{results-from-pc-algorithm}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/sum_PC_all.png}

}

\subcaption{\label{fig-pc_sum-1}Using both GaussianCI and RCoT}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/sum_PC_RCoTonly.png}

}

\subcaption{\label{fig-pc_sum-2}Using only RCoT}

\end{minipage}%

\caption{\label{fig-pc_sum}Resulting graphs of precarity factors and
depression sum score using PC.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/PC_symptom.png}

}

\caption{\label{fig-pc_sym}Resulting graphs of precarity factors and
individual depression symptoms using PC.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/PC_presum.png}

}

\caption{\label{fig-pc_presum}Resulting graphs of precarity sum score
and individual depression symptoms using PC.}

\end{figure}%

\clearpage

\subsection{Randomized Conditional Independence / Correlation Test (RCIT
\& RCoT)}\label{sec-rcot}

RCIT (Randomized Conditional Independence Test) and RCoT (Randomized
conditional Correlation Test) are advanced methods for scalable
conditional independence (CI) testing, offering computational efficiency
while maintaining the accuracy of kernel-based approaches. These methods
evaluate conditional independence between two variables \(X\) and \(Y\)
given a third variable \(Z\) while addressing computational challenges
inherent in kernel-based CI tests. In this section, we provide a
high-level overview of RCIT and RCoT based on (Strobl et al., 2019).

\subsubsection{Kernel-Based Conditional Independence
Testing}\label{kernel-based-conditional-independence-testing}

Traditional kernel-based CI tests, such as the Kernel Conditional
Independence Test (KCIT), compute dependencies using the Hilbert-Schmidt
Independence Criterion (HSIC) in reproducing kernel Hilbert spaces
(RKHS) (Zhang et al., 2012). KCIT uses the following hypothesis
framework: \[
H_0: X \perp\!\!\!\perp Y \,|\, Z, \quad H_1: X \not\!\perp\!\!\!\perp Y \,|\, Z.
\] The core quantity in KCIT is the partial cross-covariance operator:
\[
\Sigma_{XY \cdot Z} = \Sigma_{XY} - \Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY},
\] where \(\Sigma_{XY}\) represents the cross-covariance operator
between \(X\) and \(Y\), and
\(\Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY}\) removes the dependence
mediated by \(Z\).

The squared Hilbert-Schmidt (HS) norm of \(\Sigma_{XY \cdot Z}\) serves
as the test statistic: \[
\|\Sigma_{XY \cdot Z}\|^2_{HS} = 0 \quad \text{if and only if} \quad X \perp\!\!\!\perp Y \,|\, Z.
\]

KCIT estimates residual dependencies using kernel ridge regression: \[
f^*(z) = K_Z (K_Z + \lambda I)^{-1} f(x),
\] where \(K_Z\) is the kernel matrix for \(Z\), \(f(x)\) is the kernel
feature map for \(X\), and \(\lambda\) is the ridge regularization
parameter. The residual function for \(X\) is: \[
f_\text{res}(x) = f(x) - f^*(z) = R_Z f(x),
\] with: \[
R_Z = I - K_Z (K_Z + \lambda I)^{-1}.
\] The kernel matrix for residualized \(X\) is: \[
K_{X \cdot Z} = R_Z K_X R_Z,
\] and similarly for \(Y\), \(K_{Y \cdot Z} = R_Z K_Y R_Z\).

The test statistic is computed as: \[
T_{XY \cdot Z} = \frac{1}{n^2} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
\] which estimates the Hilbert-Schmidt (HS) norm of the partial
cross-covariance operator. To ensure convergence, KCIT scales the
statistic by \(n\): \[
S_K = n T_{XY \cdot Z}.
\] The null hypothesis \(H_0\) is rejected if \(S_K\) exceeds a
threshold determined by permutation or moment-matching-based null
distribution (Lindsay et al., 2000).

\subsubsection{Random Fourier Features
(RFFs)}\label{random-fourier-features-rffs}

Kernel-based methods like KCIT face scalability issues, as they involve
operations on \(n \times n\) kernel matrices, which scale quadratically
with the sample size \(n\). RCIT and RCoT overcome this bottleneck using
\emph{Random Fourier Features (RFFs)} to approximate kernel operations
efficiently.

\paragraph{Bochner's Theorem}\label{bochners-theorem}

Bochner's theorem provides the foundation for RFFs, stating that any
continuous shift-invariant kernel \(k(x, y)\) can be expressed as: \[
k(x, y) = \int_{\mathbb{R}^p} e^{i \omega^\top (x - y)} \, dP_\omega,
\] where \(P_\omega\) is the spectral distribution of the kernel. For
the widely used RBF kernel: \[
k(x, y) = \exp\left(-\frac{\|x - y\|^2}{2\sigma^2}\right),
\] \(P_\omega\) follows a Gaussian distribution:
\(\omega \sim \mathcal{N}(0, \sigma^2 I)\).

\paragraph{RFF Approximation}\label{rff-approximation}

Using Monte Carlo sampling, the kernel function is approximated as: \[
k(x, y) \approx \phi(x)^\top \phi(y),
\] where \(\phi(x)\) is the random Fourier feature mapping: \[
\phi(x) = \sqrt{\frac{2}{D}} \cos(W^\top x + b),
\] with \(W \sim \mathcal{N}(0, \sigma^2 I)\) and
\(b \sim \text{Uniform}(0, 2\pi)\). Here, \(D\) is the number of Fourier
features, which balances computational efficiency and approximation
accuracy.

\subsubsection{Differences Between RCIT and
RCoT}\label{differences-between-rcit-and-rcot}

RCIT and RCoT differ in their test statistics, computational efficiency,
and practical performance, which makes them suited for different
scenarios in causal discovery. RCIT evaluates the Hilbert-Schmidt norm
of the full partial cross-covariance operator, providing a general test
for conditional independence but at a higher computational cost. RCoT
simplifies the process by using the Frobenius norm of a
finite-dimensional residualized cross-covariance matrix, significantly
reducing complexity and improving scalability.

These distinctions are particularly important for large-scale datasets,
where RCoT's computational efficiency makes it a practical choice for
high-dimensional causal discovery tasks.

\paragraph{RCIT: Randomized Conditional Independence
Test}\label{rcit-randomized-conditional-independence-test}

RCIT tests full conditional independence by examining the squared
Hilbert-Schmidt (HS) norm of the partial cross-covariance operator
\(\Sigma_{XY \cdot Z}\): \[
S_K = n T_{XY \cdot Z} = \frac{1}{n} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
\] where \(T_{XY \cdot Z}\) is an empirical estimate of
\(\|\Sigma_{XY \cdot Z}\|^2_{HS}\). The null and alternative hypotheses
are: \[
H_0: \|\Sigma_{XY \cdot Z}\|^2_{HS} = 0, \quad H_1: \|\Sigma_{XY \cdot Z}\|^2_{HS} > 0.
\]

RCIT is a general test for conditional independence but becomes
computationally demanding as the size of \(Z\) increases, due to the
high-dimensional kernel operations required.

\paragraph{RCoT: Randomized Conditional Correlation
Test}\label{rcot-randomized-conditional-correlation-test}

RCoT simplifies the testing process by using a finite-dimensional
partial cross-covariance matrix, avoiding full HS norm calculations.
Instead, it uses the Frobenius norm of the residualized cross-covariance
matrix: \[
S' = n \|C_{AB \cdot C}\|_F^2,
\] where \(C_{AB \cdot C}\) represents the residualized cross-covariance
matrix. The hypotheses are: \[
H_0: \|C_{AB \cdot C}\|_F^2 = 0, \quad H_1: \|C_{AB \cdot C}\|_F^2 > 0.
\]

RCoT is computationally efficient and well-suited for large conditioning
sets (\(|Z| \geq 4\)). Its simplicity enables robust calibration of the
null distribution and improved scalability for high-dimensional data.

\end{document}