### 2025-05-16 draft_v1/draft_v1.qmd
draft_update

---
title: "From Causal Discovery to Intervention Simulation: Modeling Precarity and Depression in the HELIUS Study"
#subtitle: Using the HELIUS dataset
author: 
  - name: Kyuri Park
    affil-id: 1
  - name: Leonie K. Elsenburg
    affil-id: 2
  - name: Mary Nicolao
    affil-id: 2
  - name: Karien Stronks
    affil-id: 2
  - name: Vítor V. Vasconcelos
    affil-id: 1, 3

affiliations:
  - id: 1
    name: \textit{Computational Science Lab, Informatics Institute, University of Amsterdam, PO Box 94323, Amsterdam, 1090GH, the Netherlands}
  - id: 2
    name: \textit{Department of Public and Occupational Health, Amsterdam Public Health Research Institute, Amsterdam UMC, University of Amsterdam, Amsterdam, the Netherland}
  - id: 3
    name: \textit{Institute for Advanced Study, University of Amsterdam, Oude Turfmarkt 147, Amsterdam, 1012GC, the Netherland}   
format: 
  # docx:
  #   number-sections: true
  #   appendix: true
  pdf:
    pdf-engine: pdflatex
    fontfamily: palatino
    toc: true
    number-sections: true
    appendix: true
    colorlinks: true
    keep-tex: true
    template-partials:
      - title.tex
    include-in-header: header.tex
date: today
execute:
  echo: false
  warning: false
bibliography: references_draft.bib
nocite: |
  @*
csl: apa.csl
abstract: |
  \noindent This study investigates the dynamic relationship between precarious life conditions and depression by combining constraint-based causal discovery with computational modeling. Using cross-sectional data from the HELIUS study, we applied cycle-capable causal discovery algorithms to uncover directional relationships between financial stress, domain-specific precariousness indicators, and depressive symptoms. Our multi-resolution analysis—spanning both aggregate and symptom-level constructs—identified financial stress as a robust driver of depression, and highlighted key symptoms, such as sleep disturbance and depressed mood, as likely initiators or reinforcers of downstream precariousness.
  To explore how these structural insights shape intervention effectiveness, we developed a simplified nonlinear dynamical model simulating the co-evolution of depression and social precarity under varying levels of financial stress reduction. Simulation results revealed that the system's responsiveness depends not only on the intensity of external intervention but also on internal structural features such as feedback strength and noise. In particular, some configurations exhibited bistability, requiring interventions to surpass a critical threshold to trigger lasting improvement, while others responded more proportionally.
  Together, these findings emphasize that internal feedback dynamics—not just external adversity—can shape the impact of mental health interventions. More broadly, our study offers a scalable framework for integrating causal discovery and dynamical simulation to better understand and inform policy responses to complex psychosocial phenomena.

  <!-- Understanding the causal mechanisms linking precariousness and depression is critical for developing effective interventions. This study utilizes data from the HELIUS cohort study to explore these relationships using advanced causal discovery methods. By applying algorithms such as FCI, and CCI, and combining traditional Gaussian CI tests with non-parametric approaches like RCoT, we investigate how different indicators of precariousness—including factors related to employment, social relations and relational stress, financial situation, and housing—affect depressed mood, both as a sum score and at the individual symptom level. Our findings reveal that relational stress consistently emerges as a potential causal factor for depressed mood, while symptoms such as sleep disturbances, guilt, and anhedonia are particularly sensitive to external stressors, acting as potential early warning signals or intervention points for prevention. Moreover, the results highlight complexities in the data, including the influence of latent confounders and the challenges of capturing cyclic relationships. Despite some limitations, such as unresolved ambiguities in causal directions and challenges with mixed data distributions, this study demonstrates the utility of causal discovery tools in disentangling the intricate interplay between social and mental health dynamics. By mapping these causal structures into computational models, future research can simulate intervention effects, providing actionable insights to mitigate the impact of precariousness on mental health. This study serves as a foundational effort, offering both methodological advancements and practical implications for addressing depression at a population level. -->
---

```{r, include=FALSE}
library(haven)
library(dplyr)
library(skimr)
#library(bestNormalize)
library(furrr)
library(future)
library(tidyr)
library(GGally)
library(purrr)
library(gt)

library(corrplot)
library(psych)
library(FactoMineR)
library(factoextra)
library(fastICA)
library(umap)

```

```{r, include=FALSE, cache=TRUE}
# Load the dataset 
dat <- read_sav("../../data/HELIUS_LEONIE.sav")

# Preprocess and scale the precariousness data
scaled_data <- dat %>%
  # Select precariousness-related variables
                # Employment Precariousness
  dplyr::select(H1_Arbeidsparticipatie, H1_WerkSit, H1_RecentErv8, 
                # Financial Precariousness
                H1_InkHhMoeite, H1_RecentErv9,                      
                # Housing Precariousness
                veilig_2012, vrz_2012, P_HUURWON,                   
                # Cultural Precariousness
                H1_Discr_sumscore, H1_SBSQ_meanscore, A_BED_RU,         
                # Social Precariousness
                H1_RecentErv5, H1_RecentErv6, H1_RecentErv7,       
                H1_SSQT, H1_SSQSa,                                  
                # Depression symptoms / sum scores
                H1_WlbvRecent1, H1_WlbvRecent2, H1_WlbvRecent3, H1_WlbvRecent4, H1_WlbvRecent5, H1_WlbvRecent6, H1_WlbvRecent7, H1_WlbvRecent_89, H1_WlbvRecent10, H1_PHQ9_sumscore,   
                # Ethnicity
                H1_etniciteit,
                # Ags
                H1_lft) %>%                                
  # Replace missing codes with NA 
  mutate(across(everything(), ~ na_if(.x, -9)), 
         across(everything(), ~ na_if(.x, -1))) %>%  
  # Remove NAs 
  na.omit() %>%
  # Convert all columns to numeric
  mutate(across(everything(), as.numeric),
        # Combine categories 5 to 10 into a single category (5-10 as one unemployed)
         H1_WerkSit = case_when(
    H1_WerkSit %in% c(5, 6, 7, 8, 9, 10) ~ 5,  
    .default = H1_WerkSit  # Keep other values as they are 
  ))  %>%
  # Reverse the values of specific variables (reverse transformation)
  mutate(across(c(H1_RecentErv8, H1_RecentErv9, veilig_2012, vrz_2012, 
                  H1_SBSQ_meanscore, A_BED_RU, H1_RecentErv5, H1_RecentErv6, 
                  H1_RecentErv7, H1_SSQT, H1_SSQSa), 
                ~ max(.x) - .x)) %>%
  # scale the data
  mutate(across(!c(H1_etniciteit, H1_lft), scale))
  # Apply Min-Max scaling to all columns
  # mutate(across(everything(), min_max_scaling))

# redefine colnames
colnames(scaled_data) <- c(
  "emp_stat",       # H1_Arbeidsparticipatie
  "work_sit",   # H1_WerkSit
  "unemp12",     # H1_RecentErv8
  "inc_dif",      # H1_InkHhMoeite
  "fincri12",# H1_RecentErv9
  "nb_safe",     # veilig_2012
  "nb_res",   # vrz_2012
  "nb_rent",         # P_HUURWON
  "discrim",      # H1_Discr_sumscore
  "hea_lit",       # H1_SBSQ_meanscore
  "cul_rec",      # A_BED_RU
  "rel_end12",   # H1_RecentErv5
  "frd_brk12", # H1_RecentErv6
  "conf12",         # H1_RecentErv7
  "soc_freq",      # H1_SSQT
  "soc_adq",      # H1_SSQSa
  "anh",       # H1_WlbvRecent1
  "dep",       # H1_WlbvRecent2
  "slp",       # H1_WlbvRecent3
  "ene",       # H1_WlbvRecent4
  "app",       # H1_WlbvRecent5
  "glt",       # H1_WlbvRecent6
  "con",       # H1_WlbvRecent7
  "mot",     # H1_WlbvRecent_89
  "sui",      # H1_WlbvRecent10
  "PHQsum",         # H1_PHQ9_sumscore
  "ethn",         # H1_etniciteit
  "age"          #"H1_lft" 
)
```

# Introduction

Mental health disorders represent a growing global health concern, especially in urbanized regions [@WHO_2022; @gruebner2017cities]. In 2019, one in eight people worldwide were living with a mental health condition, with the highest disability-adjusted life years (DALYs) due to mental and addictive disorders concentrated in high-income countries such as those in Northern Europe, North America, and Australia [@rehm2019global; @WHO_2022]. Despite growing awareness, governments have struggled to design effective responses. Mental health outcomes arise from complex, multi-level dynamics—intertwined with the social, economic, and spatial structures of urban environments—making both diagnosis and intervention design deeply challenging [@van2021advancing].

Recent work has shifted focus from individual-level vulnerability to broader social and structural determinants. While high-income nations show elevated overall DALYs, within-country disparities reveal that income inequality is a significant predictor of mental health burden [@rehm2019global]. Moreover, precarious working conditions, housing instability, and neighborhood disadvantage have all been linked to increased risk of depression and anxiety [@pevalin2017impact; @fone2014effect; @rugulies2023work; @ronnblad2019precarious]. This growing literature highlights that poor mental health is not merely a personal issue, but one rooted in sustained socioeconomic stress and structural uncertainty.

Building on this, recent studies have introduced the concept of precariousness as a multidimensional condition characterized by instability and lack of control across several life domains—including employment, finances, housing, social relations, and cultural belonging [@elsenburg2025clustering]. This broader view reveals how different forms of insecurity can co-occur and compound, forming an ecosystem of risk. However, despite growing evidence of association between precariousness and mental health, a fundamental question remains unresolved: *How do these factors causally influence one another?* Most existing studies rely on cross-sectional correlations, leaving open questions about directionality and feedback. Without stronger causal insight, it remains difficult to identify priority targets for intervention or to anticipate unintended effects.

This study aims to investigate the causal mechanisms linking precariousness and depression—and how those mechanisms shape system responses to external intervention. Specifically, we ask: *How does the internal configuration of the precarity–depression relationship affect the outcomes of stress-reducing interventions?*
To answer this, we combine two approaches. First, we use cycle-capable causal discovery algorithms to infer directional relationships among financial stress, domain-specific precariousness factors, and depressive symptoms. By analyzing both aggregate scores (e.g., PHQ-9 total, composite precarity index) and disaggregated variables (individual symptoms and life-domain indicators), we identify both broad associations and fine-grained causal pathways.
Second, we translate these insights into a computational dynamical model that simulates how depression and precarity co-evolve under varying levels of external stress. This model enables us to probe how system properties—such as feedback strength and stochastic noise—govern responsiveness to intervention. In some configurations, the system exhibits bistability, remaining stuck in a high-risk state unless external support exceeds a critical threshold. In others, it shows unistable behavior, responding more gradually to incremental changes.
By integrating causal discovery with dynamic simulation, we aim to demonstrate how the internal structure of the depression–precarity system shapes its susceptibility to change—and, by extension, how intervention strategies might be tailored to the architecture of underlying vulnerability.

<!-- Mental health problems are rising globally, and populations facing chronic stress and socioeconomic instability bear a disproportionate share of this burden. Despite growing recognition of these challenges, designing effective policies remains difficult—largely because mental health outcomes emerge from complex, interacting factors that do not operate in isolation. To inform better interventions, we must move beyond identifying which factors are associated with mental health problems and instead uncover how those factors influence one another, the underlying causal mechanisms. -->

<!-- Precarious life conditions—such as unstable housing, insecure employment, and financial hardship—are now well established as key correlates of depression and other mental health outcomes. However, the causal dynamics linking these domains remain poorly understood. In particular, it is unclear how different forms of precariousness and depressive symptoms influence one another, and whether their interplay creates self-reinforcing loops that stabilize poor mental health and social adversity over time.  -->

<!-- This study aims to investigate the causal mechanisms linking precariousness and depression, and to examine how those mechanisms shape the system’s response to external intervention. Specifically, we ask: How do different configurations of the precarity–depression relationship affect the outcomes of interventions aimed at reducing external stress? -->

<!-- To address this, we combine causal discovery and mechanistic modeling. We apply cycle-capable causal discovery algorithms to infer directional relationships among financial stress, precariousness indicators, and depressive symptoms. Our analysis includes both aggregate scores (e.g., PHQ-9 total, precarity composite index) and disaggregated variables (individual symptoms and domain-specific precariousness), allowing us to capture both broad associations and fine-grained pathways. -->

<!-- Building on the discovered causal patterns, we simulate a set of scenarios using a computational model that captures the co-evolution of depression and precarity under varying levels of financial stress. This framework allows us to examine how internal system properties—particularly feedback strength and stochastic noise—influence responsiveness to intervention. In some regimes, the system exhibits bistability, resisting change unless pushed past a critical threshold. In others, it transitions gradually, reflecting unistable dynamics. Through this integrative approach, we aim to show how the internal architecture of the depression–precarity system fundamentally shapes its susceptibility to change—highlighting the importance of structural understanding in designing effective, equitable mental health interventions. -->

# Methods

## Data

We use data from the HELIUS (HEalthy LIfe in an Urban Setting) study, which has been described in detail elsewhere [@stronks2013unravelling; @snijder2017cohort]. It  captures the diverse population of the city of Amsterdam by including the six main ethnic groups and which provides comprehensive health and lifestyle data, including depressive symptoms as measured by the PHQ-9 [@galenkamp2017measurement]. To operationalize indicators of precariousness, we draw on the framework outlined in previous research [@elsenburg2025clustering] and select a set of relevant variables.
To ensure a robust representation of precariousness in our causal discovery models, we conducted various exploratory analyses to identify consistent and meaningful data-driven factor structures. Based on these analyses, we identified five precariousness factors, including two related to recent stressors, each comprising multiple variables as outlined below. DFurther details on the HELIUS study and factor extraction are provided in the [Appendix](#sec-appendix).

-   Employment precariousness: `emp_stat`, `work_sit`.
-   Social precariousness: `soc_freq`, `soc_adq`.
-   Housing precariousness: `nb_safe`, `nb_res`, `nb_rent`, `cul_rec`.
-   Recent relational stressors: `frd_brk12`, `conf12`.
-   Recent financial stressors: `fincri12`, `inc_diff`.

After preprocessing, the HELIUS dataset comprises 21,628 samples. Along with the five precariousness factors, we also compute an overall precarity score as the combined value of these factors. Additionally, PHQ-9 scores are included to represent depression, both as a total sum score and as individual symptom scores. In the subsequent causal discovery analysis, we examine the relationship between depression and precarity using both aggregated sum scores and their individual-level representations. Refer to @fig-dist for the overall distributions of the variables used in the analysis.

```{r echo=FALSE, fig.align='center', fig.height=5, fig.width=8, out.width="90%", cache=TRUE}
#| fig-cap: >
#|   Distributions of variables with density overlay.  
#|   
#|   *P.emp* = employment precariousness; *P.hou* = housing precariousness; *P.soc* = social precariousness; *S.fin* = recent financial stressors; *S.rel* = recent relational stressors; *Precarity* = overall precarity;
#|   *PHQsum* = PHQ-9 sum score; *anh* = anhedonia; *app* = appetite; *con* = concentration; *dep* = depressed mood; *ene* = energy; *glt* = guilty; *mot* = motor; *sui* = suicidal
#| label: fig-dist

clust_data <- scaled_data |> 
  mutate(
    P.emp = rowMeans(cbind(emp_stat, work_sit)),
    P.soc = rowMeans(cbind(soc_freq, soc_adq)),
    P.hou = rowMeans(cbind(nb_safe, nb_res, nb_rent, cul_rec)),
    S.rel = rowMeans(cbind(frd_brk12, conf12)),
    S.fin = rowMeans(cbind(fincri12, inc_dif)),
    Precarity = P.emp + P.soc + P.hou + S.rel + S.fin
  ) |>
  dplyr::select(17:34) |>
  as.data.frame()

# Reshape the data to long format
clust_data_long <- clust_data |>
  dplyr::select(-age, -ethn) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot distributions with density lines
ggplot(clust_data_long, aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)), fill = "lightblue", color = "white", bins = 15) +  # Use density for histogram
  geom_density(adjust=2, linewidth=0.3, color = "gray", alpha = 0.5) +  # Add density line in red
  facet_wrap(~ factor(variable, levels =c("P.emp", "P.hou", "P.soc", "S.fin", "S.rel", "Precarity", "PHQsum", "anh", "app", "con", "dep", "ene", "glt", "mot", "slp", "sui")), scales = "free") +
  labs(x = "", y = "") +
  theme_minimal()
```

## Causal Discovery{#sec-analysis}

To uncover directional relationships between precariousness and depression, we used constraint-based causal discovery algorithms that allow for feedback loops and latent confounding. Specifically, we applied the *Fast Causal Inference* (FCI) and *Cyclic Causal Inference* (CCI) algorithms, both of which can detect cycles and non-acyclic structures [@mooij2020constraint; @strobl2019]. For reference, we also included the *PC* algorithm, one of the most widely used methods for acyclic causal discovery [@spirtes2001causation].

A key challenge in applying these algorithms to the HELIUS dataset is that many variables are non-Gaussian and their relationships are likely nonlinear. To address this, we supplemented the commonly used Gaussian conditional independence (CI) test—based on partial correlations—with a non-parametric alternative: the *Randomized Conditional Correlation Test* (RCoT). RCoT approximates kernel-based CI testing using random Fourier features, significantly reducing computational demands while maintaining sensitivity to nonlinear dependencies [@strobl2019approximate; @zhang2012kernel].

We examined the causal structure using three complementary approaches:

1. Aggregate-level analysis: relationships between five domain-specific precariousness factors and the PHQ-9 sum score.
2. Fully disaggregated analysis: relationships between individual precariousness items and individual depressive symptoms.
3. Mixed analysis: relationships between individual symptoms and a composite index of overall precariousness.

The aggregate analysis reduces dimensionality and yields interpretable summaries of how domains of precarity relate to overall depression severity. However, it may obscure finer-grained effects. The disaggregated analysis allows for precise mapping of which symptoms are influenced by (or influence) specific precariousness conditions, but it introduces complexity due to the high dimensionality and distributional properties of the data. The mixed approach, in turn, integrates these perspectives by examining how individual symptoms respond to cumulative precarity across domains.

To ensure robust estimation and guard against sensitivity to parameter settings, we implemented a systematic grid of analysis conditions:

* Significance levels: $\alpha$ = 0.01 and 0.05
* Bootstrap thresholds: 0.5, 0.6, 0.7, 0.8
* CI tests: Gaussian and RCoT
* Algorithms: FCI, CCI, and PC

Each combination was bootstrapped 100 times for sum-score level analyses and 30 times for symptom-level analyses, yielding a total of 1,600 graphs per algorithm for the full analysis. For symptom networks, we fixed the skeleton structure using a consensus graph estimated across PC, FCI, and CCI ($\alpha$ = 0.01, RCoT). This constraint improves computational efficiency without limiting the estimation of directional edges between symptoms and precariousness variables.

To summarize the results, we adopted a two-level aggregation strategy. First, within each analysis condition, we recorded the most frequently observed edge-endpoint type (e.g., directed, undirected, uncertain) from the bootstraps. Then, across all conditions, we tallied the most frequent endpoint type per edge. In cases of ties, we assigned dashed edges to reflect uncertainty. For transparency, we accompany each graph with a table showing the relative frequencies of each edge symbol.

This procedure yields a final graph for each algorithm and analysis type—offering both robustness and interpretability. See @fig-workflow for an overview of the causal discovery pipeline.

For detailed explanations of the algorithms, edge interpretations, and graph types (e.g., PAG, MAAG, CPDAG), as well as algorithmic assumptions and limitations, see the [Appendix](#sec-appendix).

<!-- To uncover directional relationships between precariousness and depression, we used constraint-based causal discovery algorithms capable of detecting feedback loops and latent confounding. Specifically, we applied the *Fast Causal Inference* (FCI) and *Cyclic Causal Inference* (CCI) algorithms, both of which support cyclic and non-acyclic structures [@mooij2020constraint; @strobl2019]. Additionally, we include the *PC* algorithm as a reference, given its simplicity and prominence as one of the most widely known causal discovery methods [@spirtes2001causation]. -->

<!-- One practical challenge in applying these algorithms to the HELIUS dataset is that most variables do not follow a Gaussian distribution, and the relationships among variables are unlikely to be strictly linear. To address this, we supplement the commonly used Gaussian conditional independence (CI) test, which relies on partial correlations, with a non-parametric CI test based on reproducing kernel methods [@zhang2012kernel]. -->
<!-- While kernel-based conditional independence tests (KCIT) are theoretically effective in capturing non-linear dependencies and subtle interactions in data with non-Gaussian noise, they are computationally demanding for large datasets like HELIUS. Their quadratic scaling with sample size is primarily due to the need to invert large kernel matrices, making them impractical for large-scale analysis [@rahimi2007random]. -->
<!-- To mitigate this issue, we use the Randomized Conditional Correlation Test (RCoT), which approximates kernel methods using random Fourier features. This approach reduces computational complexity from quadratic to linear scaling with sample size, significantly lowering computational costs while maintaining robustness [@strobl2019approximate]. For a more detailed explanation of RCoT, refer to @sec-rcot. -->

<!-- We analyze the causal structure using three complementary approaches: (1) Examining the relationship between the five individual precariousness factors and the PHQ sum score, which represents overall depression severity. -->
<!-- (2) Investigating the relationship between individual depressive symptoms and individual precariousness factors. (3) Assessing the relationship between individual depressive symptoms and overall precariousness (the sum of all precariousness factors). -->
<!-- Using the PHQ sum score as a measure of depression simplifies the analysis by reducing dimensionality, improving computational efficiency, and providing a broad, interpretable perspective on how precariousness factors relate to overall depression severity. However, this aggregated approach may overlook granular symptom-level relationships. -->

<!-- By zooming in on individual symptoms, we capture a more detailed picture of how specific depressive symptoms interact with different precariousness factors. This approach allows us to map individual causal pathways, identifying which symptoms are particularly sensitive to certain precarious conditions and vice versa. However, this level of detail introduces methodological challenges, including non-standard distributions of symptom variables and the high-dimensional nature of the analysis, which can reduce statistical power and significantly increase computational demands—especially when using non-parametric conditional independence (CI) tests. -->

<!-- Finally, analyzing individual symptoms in relation to overall precariousness provides an integrative perspective, revealing how depression symptoms collectively relate to precarious conditions. This approach complements the fully disaggregated analysis by aggregating precariousness factors, potentially uncovering relationships that may be too weak to detect when examined separately. This is particularly relevant when a symptom is associated with multiple precariousness factors, but these associations are not strong enough to emerge in an individual-level analysis. -->
<!-- By integrating these three approaches — ranging from partially aggregated to fully disaggregated — we aim to achieve a balance between capturing detailed granularity and gaining a broader understanding of the relationship between precarity and depression.  -->

<!-- Furthermore, to address sensitivity to parameter values and enhance the robustness of our results, we implement multiple settings combined with bootstrapping. For each condition, we generate 100 bootstrap samples, estimate causal graphs, and retain only the edge-endpoints that exceed predefined thresholds. If no edge-endpoint type surpasses the threshold, we assign a circle ($\circ$) to indicate uncertainty in the edge direction. -->

<!-- The analyses are conducted under the following conditions: -->

<!-- -   Significance levels ($\alpha$): 0.01 and 0.05 -->
<!-- -   Thresholds: 0.5, 0.6, 0.7, and 0.8 -->
<!-- -   CI test: Gaussian CI test, RCoT -->
<!-- -   Algorithms: FCI, CCI, and PC -->

<!-- This setup yields 16 combinations (2 significance levels × 4 thresholds × 2 CI tests), applied across three algorithms. Each combination is repeated for 100 bootstrap samples, yielding a total of 1,600 graphs per algorithm.  -->
<!-- For analyses involving individual symptom variables, we simplify the setup by focusing on thresholds of 0.6 and 0.7 and reducing the number of bootstrap samples to 30, easing computational demands. -->
<!-- To further enhance efficiency, we fix the skeleton of the symptom network using a common structure derived from PC, FCI, and CCI algorithms ($\alpha$ = 0.01, RCoT CI test). This step constrains which edges can be present among symptoms but does not restrict the estimation of causal directions within the symptom network. Additionally, the entire causal structure between symptoms and precariousness factors remains fully estimable, ensuring that key relationships are still inferred. -->
<!-- Beyond its computational benefits — particularly in reducing the time required for skeleton estimation — this fixed structure also aligns well with commonly reported symptom network structures in the literature [@park2025individual], ensuring consistency with existing findings. -->

<!-- To obtain a final, stable summary of causal relationships, we identify the most frequently occurring edge-endpoints across different experimental setups. First, within each condition, we determine the most frequent edge-endpoint from all bootstrap samples. Then, across different conditions, we aggregate these selected symbols and determine the overall most common edge-endpoint for each graph entry. -->

<!-- Since we have an even number of experimental setups, ties may occasionally occur. In such cases, we explicitly denote them in the resulting graph using dashed lines. To further help with the interpretation and understanding of the resulting graph, we provide a corresponding table for each graph, showing the proportion of each edge-endpoint type to clarify how dominant a given symbol is in each case. -->

<!-- By double summarizing the results in this way, we generate one final graph per algorithm for each analysis approach: the partially aggregated analysis, which uses either the sum of precariousness factors or the sum of depression factors to capture overall trends, and the fully disaggregated analysis, which examines individual precariousness and depression factors separately. -->
<!-- This approach ensures that only stable and consistent edges — those that persist across different parameter settings — are retained, leading to a robust and reliable understanding of the relationships between precariousness factors and depression. See @fig-workflow for an overview of the analysis workflow. -->

```{r}
#| label: fig-workflow
#| fig-cap: "Analysis workflow applied across all three algorithms."
#| fig-align: center
#| out-width: 80%

knitr::include_graphics("img/simsetup2.png")

```

## Modeling Dynamics between Precarity and Depression

To explore how internal system structure shapes responses to external support, we implement a high-level computational model that formalizes key mechanisms suggested by the causal discovery analysis. Rather than attempting to replicate the full estimated causal graph, this model abstracts the essential structure into a compact, tractable system focused on dynamic interactions between depression and social precariousness, both influenced by financial stress.

As illustrated in @fig-concept, the model includes three components: depression ($D$), represented by the PHQ-9 sum score, social precariousness ($P$), and external financial stress ($S$). Depression and precariousness are modeled as coupled state variables that influence each other over time, with $S$ acting as a shared input. These dynamics are formalized as a pair of nonlinear stochastic differential equations (SDEs):

$$
\begin{aligned}
dP &= \lambda_P \left( \tanh(\alpha_1 S + \alpha_2 D) - P \right) dt + \sigma_1 dW_1 \\
dD &= \lambda_D \left( \tanh(\beta_1 S + \beta_2 P) - D \right) dt + \sigma_2 dW_2 
\end{aligned}
$$

Here, $D$ denotes depression, $P$ denotes social precariousness, and $S$ is the external stressor associated with financial stress. The parameters $\lambda_D$ and $\lambda_P$ control the timescales over which variable adjusts toward its input-driven target state. The coefficients $\alpha_1, \alpha_2$ and $\beta_1, \beta_2$ represent the strengths of the causal influences between variables. 
Stochastic fluctuations are introduced through the terms $\sigma_1 dW_1$ and $\sigma_2 dW_2$, where $dW_1$ and $dW_2$ are independent Wiener processes. The system uses the hyperbolic tangent function ($tanh$) to bound the influence of inputs, ensuring smooth, saturating dynamics and preventing runaway feedback loops. Importantly, the nonlinearity introduced by the $tanh$ function also allows the system to exhibit bistability, supporting the existence of two stable states: one corresponding to low depression and precariousness, and the other to high depression and precariousness.

<!-- Building on the causal structure derived from the result of causal discovery, we implement a high-level simulation model to explore the potential impact of targeted interventions on precariousness. This serves as a proof-of-concept for translating causal relationships into dynamic systems that can generate actionable insights. -->

<!-- The structure of the simulation model, illustrated in @fig-concept, captures the dynamic interplay between depression ($PHQsum$) and social precariousness ($P.soc$), both of which are influenced by an external financial stressor ($S.fin$). Importantly, the model incorporates cyclic dependencies between depression and precariousness, reflecting their reciprocal influence over time. -->
<!-- This system is formalized as a set of non-linear stochastic differential equations (SDEs): -->

<!-- $$ -->
<!-- \begin{aligned} -->
<!-- dP &= \lambda_P \left( \tanh(\alpha_1 S + \alpha_2 D) - P \right) dt + \sigma_1 dW_1 \\ -->
<!-- dD &= \lambda_D \left( \tanh(\beta_1 S + \beta_2 P) - D \right) dt + \sigma_2 dW_2  -->
<!-- \end{aligned} -->
<!-- $$ -->

<!-- Here, $D$ denotes depression, $P$ denotes social precariousness, and $S$ is the external stressor associated with financial stress. The parameters $\lambda_D$ and $\lambda_P$ control the timescales over which variable adjusts toward its input-driven target state. The coefficients $\alpha_1, \alpha_2$ and $\beta_1, \beta_2$ represent the strengths of the causal influences between variables.  -->
<!-- Stochastic fluctuations are introduced through the terms $\sigma_1 dW_1$ and $\sigma_2 dW_2$, where $dW_1$ and $dW_2$ are independent Wiener processes. The system uses the hyperbolic tangent function ($tanh$) to bound the influence of inputs, ensuring smooth, saturating dynamics and preventing runaway feedback loops. Importantly, the nonlinearity introduced by the $tanh$ function also allows the system to exhibit bistability, supporting the existence of two stable states: one corresponding to low depression and precariousness, and the other to high depression and precariousness. -->

```{r}
#| label: fig-concept
#| fig-cap: "Conceptual high-level causal structure of precariousness and depression system."
#| fig-align: center
#| out-width: 35%

knitr::include_graphics("img/conceptmodel.png")
```

### Model Calibration

To identify realistic parameter settings for the simulation model, we calibrated six free parameters—$\alpha_1$, $\alpha_2$, $\beta_1$, $\beta_2$, and noise amplitudes $\sigma_1$, $\sigma_2$—using the Non-dominated Sorting Genetic Algorithm II (NSGA-II) [@deb2002fast], a multi-objective evolutionary optimization method well-suited for noisy and high-dimensional search spaces. The goal was to match simulated system behavior to empirical summary statistics derived from the HELIUS dataset, including the means and variances of depression ($D$) and social precariousness ($P$), as well as their partial correlations after adjusting for financial stress ($S$), specifically $\rho_{DP.S}$, $\rho_{DS.P}$, and $\rho_{PS.D}$. 
Given the stochastic nature of the model, each candidate parameter set was evaluated by simulating 1,000 independent trajectories, each representing an individual with a fixed $S$ value sampled from the observed distribution. Final-state values were extracted and aggregated to compute population-level summary statistics. Fitness was assessed as the absolute deviation from the empirical targets, aggregated across all objectives. 
NSGA-II was implemented via the nsga2R package [@nsga2R], with a population of 300 evolved over 300 generations using standard hyperparameters (crossover probability = 0.9, mutation probability = 0.25, and tournament size = 2). This process yielded a diverse ensemble of Pareto-efficient solutions balancing trade-offs between objectives. From the resulting Pareto front (@fig-pareto), we selected the top 30 parameter sets with the lowest combined error to use in the subsequent simulations assessing intervention effects.

```{r}
#| label: fig-pareto
#| fig-cap: >
#|   Pareto front of simulated parameter sets.
#|   *Each point represents a solution evaluated by its total variance error (x-axis) and total partial correlation error (y-axis), with color indicating total mean error.
#|   Red crosses highlight the top 30 solutions used for simulation.*
#| fig-align: center
#| out-width: 80%
knitr::include_graphics("img/paretoplot.png")
```

### Intervention Simulation

To evaluate the effects of hypothetical support, we simulate scenarios in which each individual’s financial stress level ($S$) is proportionally reduced toward the minimum observed value. This models a population-wide intervention applied at varying strengths --- from no reduction in stress (0%) to complete alignment with the minimum observed stress level (100%).

For each of the top 30 calibrated parameter sets, we simulate the system’s evolution across 300 individuals under a continuum of intervention levels. Each simulation run generates final values for $D$ and $P$, which are then used to assess how internal model parameters influence the system’s behavior under changing external conditions.

# Results

## Causal Structure Linking Precarity and Depression

### Depression as sum score
The sum score graphs provide a high-level summary of how precarity factors collectively influence overall depression severity, focusing on aggregated relationships.
@fig-sum illustrates the causal relationships between precariousness factors (*P.hou*, *P.emp*, *P.soc*, *S.rel*, *S.fin*) and the depression sum score (*PHQsum*), as identified by the FCI algorithm.
As described in [@sec-analysis], dashed edges indicate ties, with the exact edge endpoint where the tie occurs is represented by a bold horizontal bar.

The causal graph in panel (a) highlights key pathways in the relationships between precarity factors and depression (PHQsum). Employment precarity (P.emp) and social precarity (P.soc) are not identified as causes of depression, whereas financial stress (S.fin) appears to play a causal role. While P.emp and P.soc are generally not recognized as causes of other precariousness factors, S.fin emerges as a potential cause, as indicated by its circle edge endpoint. Supporting this, the proportion matrix plot in panel (b) shows that S.fin has some probability of causally influencing either P.emp or P.soc.

The matrix plot can be interpreted such that the symbol in matrix[$i$, $j$] represents the relationship $i -~[symbol] \  j$. For example, if matrix[$i$, $j$] is > and matrix[j, i] is $\circ$, then the inferred relationship is $i \circarrow j$.  In the table, different edge types are represented by distinct colors: light gray for the absence of an edge, blue for circles, green for arrowheads, and coral for arrowtails. The colors are blended based on the proportion of each edge type, with higher proportions increasing the opacity of the corresponding color, making dominant symbols more visually prominent.

Relational stress (S.rel) plays a more nuanced role, interacting with depression and S.fin  through a latent confounder. Its relationship with P.soc remains less certain, as the proportion table suggests a fair probability that this connection is absent, uncertain, or bidirectional. Meanwhile, housing precarity (P.hou) is not directly causally related to depression but is linked to P.emp and S.fin. While P.emp and S.fin are identified as non-causes of P.hou, it remains unclear whether P.hou causally influences P.emp or if their relationship is mediated by an unobserved confounder.

The findings from the CCI algorithm (see [@fig-sum-cci]) are largely consistent with those from the FCI algorithm, with one exception: CCI does not identify S.fin as a cause of depression. Additionally, CCI introduces greater uncertainty in edge directions, with more ties mainly between circle and arrowhead endpoints, and tends to favor arrowheads more frequently than FCI. Despite these differences, the skeleton structure and overall causal directions derived from CCI align well with the results of the FCI algorithm, supporting the key pathways.

<!-- In  graph  (a),  the  key  pathways  suggest  that  employment  precarity  (P.emp) and social precarity (P.soc) do not cause depression (PHQsum).P.soc does not cause recent relational stress (S.rel) or financial stress (S.fin), and these stressors are likely related by a latent confounder.   Additionally, P.emp and S.fin are identified as non-causes of housing precarity (P.hou). Both FCI and CCI detect a dependency between P.emp and S.fin, but they disagree on the direction of the relationship, leaving it unclear whether S.fin causes P.emp or whether a latent confounder mediates their relationship. Similar ambiguities are found in the connections between S.rel and PHQsum and between S.fin and PHQsum -->

<!-- Both FCI and CCI detect a dependency between *P.emp* and *S.fin*, but they disagree on the direction of the relationship, leaving it unclear whether *S.fin* causes *P.emp* or whether a latent confounder mediates their relationship. Similar ambiguities are found in the connections between *S.rel* and *PHQsum* and between *S.fin* and *PHQsum*. -->

<!-- Graph (b), derived solely from the nonparametric RCoT method. Both graphs consistently identify that *P.emp* is not a cause of *P.hou* or *PHQsum*, and that *P.soc* does not cause *PHQsum* or *S.fin*.  -->
<!-- The relationship between stressors and depression remains unresolved between the two algorithms. However, the graph derived from RCoT provides greater confidence that depression is not the cause of *S.fin* and suggests a stronger likelihood that the financial stress may contribute to depression or that their relationship is mediated by a latent confounder. -->
<!-- The role of relational stress has become less pronounced, as the edge between *P.soc* and *S.rel* is omitted, and the relationship between *S.rel* and *S.fin* is now inconsistent between the algorithms. -->

<!-- Overall, both graphs together highlight the potential roles of stressors like *S.fin* and *S.rel*, which are closely tied to depression, either causally or via latent confounders. Employment and social precarity may also be influenced by depression, either directly or through latent variables. Lastly, housing precarity does not directly impact depression but remains connected to employment precarity and financial stress, likely mediated by direct or latent pathways. -->

```{r}
#| layout-ncol: 2
#| label: fig-sum
#| fig-cap: "Resulting graph of precarity factors and depression sum score using FCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "FCI PAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/FCI_depsum.png")
knitr::include_graphics("img/depsum_mat_fci.pdf")

```

<!-- \begin{figure}[htbp] -->
<!--     \centering -->
<!--         \caption{Summary of all sum score graphs} -->
<!--         \includegraphics[width=1\textwidth]{img/sum_all.png} -->
<!-- \end{figure} -->
<!-- \begin{figure}[htbp] -->
<!--     \centering -->
<!--         \caption{Summary of sum score graphs given RCoT} -->
<!--         \includegraphics[width=1\textwidth]{img/sum_rcotonly.png} -->
<!-- \end{figure} -->

### Individual depression symptom

Moving from the sum score representation to the disaggregated symptom-level graph provides a more granular perspective on the causal relationships between precariousness factors and depression symptoms. Unlike the sum score graph, which aggregates all symptoms into a single measure—potentially obscuring nuanced relationships—the symptom-level graph highlights heterogeneity in how precariousness factors influence individual depressive symptoms: con (concentration), slp (sleep), ene (energy), app (appetite), mot (motor), sui (suicidal), anh (anhedonia), glt (guilt), and dep (depressed mood), and vice versa.

The symptom-level graph in @fig-sym panel (a) reveals a more complex and interconnected structure than the sum score graph, reflecting the strong interdependence among symptoms and suggesting much presence of latent confounders, as indicated by numerous bidirectional edges. As before, dashed edges denote ties, with specific ties marked by bold horizontal bars. 
In @fig-sym panel (b), the proportion matrix provides more insight into these directionalities. Here, opaque green dominates most symptom-to-symptom connections, indicating that arrowheads are the most frequently inferred edge type. However, blue regions, which correspond to circle endpoints, are particularly common for anh, slp, ene, and sui, reflecting uncertainty in the causal direction among these symptoms. Notably, the only arrowtail connection appears between dep and anh, suggesting that anhedonia (anh) is a potential cause of depressed mood (dep).

Looking at the symptom-precariousness connections, one of the key patterns is that ties are most frequently found in the relationships between individual symptoms and precariousness factors. This suggests that these causal links may be less stable across different conditions. Upon closer examination, most ties arise due to discrepancies between the Gaussian CI test and the RCoT test—with Gaussian CI favoring arrowheads and RCoT more often predicting the absence of an edge.

This discrepancy likely arises from a combination of factors. 
First, as the analysis shifts from aggregated to disaggregated variables, the number of conditional independence (CI) tests increases, leading to a loss of statistical power in high-dimensional causal discovery. 
With more variables being conditioned on, the data becomes sparser, making it harder to detect weaker dependencies. Additionally, over-conditioning—controlling for too many variables—can artificially remove statistical associations, leading to false negatives.
Beyond this general statistical power issue, the fundamental differences in how these CI tests operate contribute to the inconsistency. RCoT, as a nonparametric test, does not assume linearity and can capture both linear and nonlinear dependencies. However, it is more conservative and requires larger sample sizes, making it more prone to false negatives in high-dimensional settings. 
In contrast, partial correlation (Gaussian CI test) assumes linear Gaussian relationships, making it more permissive and sometimes detecting weak dependencies that may not be statistically meaningful. 
As a result, when Gaussian CI detects an edge while RCoT does not, it could indicate that the relationship is weak, strictly linear, or requires more data for reliable nonparametric detection.
This explains why many ties in the proportion matrix show a 50-50 split between symbols—where half of the conditions with RCoT suggest the absence of an edge, while the other half using Gaussian CI test favor an arrowhead. This systematic divergence highlights the challenges of causal discovery in high-dimensional settings, where varying statistical assumptions can lead to differing conclusions about the presence and direction of causal relationships.

Despite these differences, some consistent patterns emerge across both CI tests, particularly in the case of financial stress (S.fin), which appears to be connected to nearly all other variables—though many of these connections are marked by ties, reflecting uncertainty in directionality. One exception is the stronger evidence suggesting that S.fin may cause changes in app (appetite), while its relationships with other symptoms, such as ene, mot, and dep, remain ambiguous.
Similarly, social precarity (P.soc) exhibits numerous connections with depressive symptoms, with most edges pointing toward P.soc rather than outward from it. This suggests that depressive symptoms, particularly dep and glt, may contribute to worsening social precarity rather than the other way around. Additionally, anh also shows some probability of causally influencing P.soc, reinforcing the idea that social precarity is more often a consequence rather than a driver of depressive symptoms.

Other precariousness factors exhibit more uncertain but still notable relationships. Relational stress (S.rel) shows weak but existing connections with slp, glt, and app, though directionality remains unclear in many cases. Employment precarity (P.emp) also connects to symptoms like slp, mot, and con, but, like S.rel, these relationships exhibit a 50/50 split in directionality, reflecting uncertainty in the inferred causal paths.
In line with the aggregated graph in @fig-sum, housing precarity (P.hou) does not appear to have any direct relationship with depression symptoms but consistently shows associations with employment precarity (P.emp) and financial stress (S.fin). The causal relationships among precariousness factors remain largely unchanged from the aggregated analysis, with financial stress (S.fin) exhibiting the strongest tendency to influence other precariousness factors.

Among depressive symptoms, dep, glt, and slp appear to be the most connected to precariousness factors, while P.soc has the most connections with symptoms, predominantly as a recipient rather than a driver of influence. Within the symptom network, slp, sui, and anh emerge as causally influential symptoms, as they exhibit more outgoing arrows compared to other symptoms. On the other hand, con, mot, dep, and glt, despite having high connectivity, predominantly receive incoming arrows, indicating they are more likely effects rather than causes.
Considering both symptom-precariousness connections and symptom-level dynamics, sleep disturbance (slp) emerges as a central symptom, given its strong ties to precariousness factors and its influential role within the symptom network. This suggests that sleep issues could be an initiating symptom, particularly sensitive to relational stress and employment precarity. On the other hand, social precarity (P.soc) serves as a key bridge between the depression and precariousness subsystems, as depressive symptoms appear to feed back into social precarity, reinforcing a self-sustaining dynamic between depression and precarious conditions.

Finally, the results from CCI are largely consistent with those from FCI, but with some key differences. CCI tends to favor arrowheads more frequently, resulting in a greater number of bidirectional edges, which suggests a higher involvement of latent confounders. Additionally, CCI exhibits more tie situations, though, similar to FCI, most ties occur between absence and arrowhead edges, reflecting the discrepancy between the Gaussian CI test and RCoT—where the Gaussian test favors arrowheads, while RCoT more often suggests the absence of an edge. For a detailed visualization of the CCI-derived graph and proportion matrix, see @fig-sym-cci.

<!-- <!-- Centrality and Connectivity --> 
<!-- Certain nodes in the network emerge as more *causally* central, highlighting their potential importance as intervention points. Among the depressive symptoms, *anh*, *dep*, *slp*, and *glt* emerge as particularly influential, with multiple outgoing edges (o->) to other symptoms,  suggesting their roles as potential key drivers within the symptom network. Especially, symptoms such as *glt*, *slp*, and *anh* are particularly critical, potentially acting as initiator or activator nodes within the network due to their apparent connections with precarity factors. These symptoms appear to be especially sensitive to external stressors, potentially manifesting early in response to such conditions and subsequently activating other interconnected symptom nodes. -->
<!-- <!-- This sensitivity indicates their potential as early warning signs or key targets for interventions aimed at preventing the progression to severe depression. --> 
<!-- The causal structure involving precarity factors in the symptom-level graph is largely consistent with the patterns observed in the sum score graphs. As in the sum graphs, relational stress (*S.rel*) emerges as a potential causal factor for depression, while employment (*P.emp*) and social precarity (*P.soc*) are more likely to be influenced by depressive symptoms.  -->

<!-- The symptom-level analysis, however, provides more specificity by pinpointing the symptoms involved in these relationships. For instance, *slp* is identified as influencing *P.emp* or potentially through a latent confounder (*slp* o-> *P.emp*) , while *glt* appears to affect *P.soc* or may be linked via a confounder (*glt* o-> *P.soc*). -->
<!-- Additional edges are observed in the graph generated using Gaussian CI tests, such as *slp* <-> *S.rel* and *slp* o-> *P.emp*. As seen in the sum graph, Gaussian CI testing tends to produce a denser graph. In this case, causal relationships involving *slp* are particularly prominent. -->

<!-- Some discrepancies, however, exist between the symptom-level and sum graphs. For example, financial stress (*S.fin*) does not have any edges with depressive symptoms in the symptom-level graph, although it maintains associations with other precarity factors. Additionally, housing precarity (*P.hou*) becomes entirely disconnected from the rest of the network, appearing as an isolated node.  -->

<!-- However, due to computational constraints that required limiting the sample size, as well as the challenging distributions of certain symptoms, the statistical power is lower compared to the depression sum score graph. -->
```{r}
#| label: fig-sym
#| layout-nrow: 2
#| fig-cap: "Resulting graph of precarity factors and individual depression symptoms using FCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "FCI PAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 70%

knitr::include_graphics("img/symptom_graph_FCI.png")
knitr::include_graphics("img/symptom_mat_fci.pdf")

```

### Precarity as sum score

Lastly, we examine the relationships between individual symptoms and overall precarity, an aggregated measure represented by the sum of five precariousness factors. This analysis provides a complementary perspective, capturing broader patterns that may not be evident in the disaggregated symptom-precariousness analysis. By aggregating precariousness factors into a single score, this approach may capture distributed relationships across different precarity factors that were previously overlooked in the disaggregated analysis.

Unlike the disaggregated graph in @fig-sym, the aggregated symptom-precariousness graph (@fig-presum panel (a)) shows more certainty in causal directions, with only a few edges resulting in ties, all of which involve the overall precarity factor. This is even more evident in the proportion matrix (@fig-presum panel (b)), where symptom-to-symptom interactions almost fully converge to a single edge type. Additionally, all symptom interactions are connected either through bidirectional edges or a combination of an arrowhead and a circle, suggesting a strong presence of latent confounders.

Examining the symptom-precarity connections, we observe a stronger overall trend of depressive symptoms influencing precarity, rather than the other way around. Specifically, anh, dep, and slp show a high probability of causing precarity, whereas app, glt, and mot exhibit more uncertainty in directionality, often resulting in circle endpoints. This pattern suggests that when precariousness factors are aggregated, the dominant causal flow is from depressive symptoms to precarity, rather than precarity driving depression. Particularly, dep emerges as the strongest predictor of precarity, exhibiting the highest proportion of arrowtails, consistent with findings from the disaggregated analysis.

Comparing this with the CCI-derived graph (@fig-presum-cci), several key differences stand out. The CCI results reveal that most symptoms are interconnected through bidirectional edges, suggesting a strong presence of latent confounders. However, dep appears to be a distinct exception, as it is predominantly caused by nearly all other symptoms, with one exception sui, which does not contribute to dep. Similarly, con is influenced by multiple symptoms, albeit with weaker support compared to dep.
Most interestingly, the CCI results highlight a distinct feedback loop between dep and overall precarity, represented by a tail-tail edge (\textemdash \). This suggests a possible reinforcing cycle between depression and overall precariousness, where depression not only arises from precarious conditions but also contributes to their persistence, creating a self-sustaining dynamic.

Overall, the aggregated precariousness graph supports the key findings from both the disaggregated analysis and the depression sum score analysis, providing greater certainty in several important patterns. The results highlight the significant role of latent confounders in symptom-to-symptom interactions. While certain stressor-related precariousness factors, particularly S.fin and S.rel, appear to causally influence depressive symptoms, other precariousness factors are more likely driven by depression rather than acting as primary causes, suggesting that precariousness may function more as a consequence of depression than its source. Some symptoms, particularly glt, slp, and dep, show greater sensitivity to precarious conditions, while also playing a central role in symptom dynamics. Among them, slp and glt emerge as influential symptoms within the depression network, whereas con and dep, despite their high connectivity, behave more as outcomes rather than primary drivers. Together, these findings suggest that depression—especially symptoms like sleep disturbances and guilt—may contribute to sustaining precarious conditions, possibly forming a reciprocal cycle between depression and precariousness.

<!-- As before, black solid lines indicate edges agreed upon by both FCI and CCI, while gray dashed lines represent edges where the two algorithms disagree. Circles at the endpoints of gray dashed edges denote unresolved directionality. Additionally, navy dashed edges highlight relationships that appear exclusively in graphs generated using Gaussian CI tests. -->
<!-- Compared to @fig-sym, this analysis reveals greater disagreement between FCI and CCI regarding edge directionality, as evidenced by more dashed edges. FCI tends to estimate more undirected edges (circles), while CCI more frequently assigns directed arrows between symptom-symptom relationships. For a detailed breakdown, see [@sec-propmatrix], which shows the proportion of symbols produced by each algorithm. -->
<!-- When examining overall precarity, we observe that more symptoms are now connected to the precarity score (*Precarity*) compared to the individual precarity factor analysis in @fig-sym. In addition to *glt*, *slp*, *anh*, and *sui*, which are previously shown to be connected to individual precarity factors, *mot*, *app*, and *dep* are now identified as having causal relationships with overall precarity. Furthermore, when using the Gaussian CI test, con is also shown to be causally related to overall precarity. Notably, all these connections are directed towards precarity (o→), indicating that overall precarity is not the cause of symptoms. The exception is the bidirectional edge between *sui* <-> *Precarity*, which suggests a connection via latent confounders. -->
<!-- FCI predominantly estimates directed arrows (→) for symptom-precarity relationships, suggesting that symptoms are driving precarity. With fewer variables (reduced from 14 to 9), this approach appears to have slightly more statistical power, enabling algorithms to estimate relationships with greater confidence compared to the previous analysis involving individual symptoms and individual precarity factors. -->

```{r}
#| layout-nrow: 2
#| label: fig-presum
#| fig-cap: "Resulting graph of precarity sum score and individual depression symptoms using FCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "FCI PAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/presum_graph_FCI.png")
knitr::include_graphics("img/presum_mat_fci.pdf")
```

## Simulating the Dynamics Between Precarity and Depression Under Intervention

To evaluate how internal system structure influences the effect of external support, we simulated outcomes under a range of intervention strengths across the top 30 parameter sets from the calibrated model ensemble. The top panel of @fig-intervention shows mean levels of depression ($D$) and precarity ($P$) after simulation, with each line representing one parameter set. Lines are color-coded by combined feedback and input strength, calculated as the product $\alpha_1 \times \alpha_2 \times \beta_1 \times \beta_2$.

Parameter sets with stronger interactions exhibit more pronounced declines in both depression and precarity in response to intervention. At full intervention—modeled as a complete shift of external stress ($S$) toward its observed minimum—the mean depression level decreases from -0.113 to -0.963, and precarity from -0.100 to -0.958. These changes correspond to reductions of approximately 90.8\% relative to their initial magnitudes, indicating that targeted intervention can shift the system toward a substantially lower-risk equilibrium.

To better understand these dynamics, we focused on two illustrative parameter sets: one with strong feedback/input coupling and low noise (a "bistable" candidate), and one with weaker coupling and higher noise (a "unistable" candidate). In nonlinear systems, bistability refers to the existence of two distinct stable equilibria—meaning the system’s long-term outcome depends on its starting point or perturbations.

In our model, bistability arises from the interaction between strong internal coupling—particularly the product $\alpha_2 \times \beta_2$—and the system’s nonlinearity. The $\tanh(\cdot)$ activation function saturates at high input values, meaning sufficient gain is required to push the system into nonlinear regimes where multiple stable points emerge. If the feedback gain is too low, or the system is overwhelmed by noise, it behaves more linearly and settles into a single equilibrium—yielding a unistable configuration.[^1]

[^1]: This resembles a classic *pitchfork bifurcation* in discrete dynamical systems: when the gain $k$ in a scalar map like $x_{t+1} = \tanh(k x_t)$ exceeds a critical value, the system shifts from a single fixed point to two stable attractors and one unstable one. In our model, high values of $\alpha_2 \times \beta_2$ similarly enable the emergence of multiple attractor states.

The bottom panel of @fig-intervention illustrates these contrasting behaviors. In the bistable model, a clear bifurcation emerges: small interventions have little effect, but once a threshold is crossed, the system abruptly transitions to a low-risk state. In the unistable model, the transition is gradual and continuous, with outcomes improving incrementally as stress is reduced.

These findings underscore that the impact of intervention depends not only on its intensity, but on the system's internal configuration. In bistable regimes, small or moderate interventions may be insufficient unless they push the system past a tipping point. Unistable systems, by contrast, respond more linearly, offering greater predictability and flexibility in designing incremental interventions.

This distinction is especially relevant for mental health interventions in populations exposed to persistent stressors like housing instability, financial strain, or job insecurity. In such contexts, external adversity may be reinforced through internal feedback loops between psychological distress and precarious conditions. Recognizing when a system is bistable may help identify when strong, early, or multifaceted interventions are needed to produce lasting change—while unistable dynamics may justify more gradual or targeted efforts.

Tailoring intervention strategies to the system’s underlying structure may enhance their effectiveness and sustainability, particularly in addressing the compounding effects of socioeconomic precarity on mental health.

<!-- The final outcomes, specifically the mean levels of depression ($D$) and precarity ($P$), are presented in the top panel of @fig-intervention. Each trajectory corresponds to a parameter set and is color-coded by its combined feedback and input strength, defined as the product $\alpha_1 \times \alpha_2 \times \beta_1 \times \beta_2$. -->

<!-- Systems with stronger parameter interactions exhibit more pronounced shifts in behavior, particularly under intensive interventions. At full intervention (i.e., maximum stress reduction), the mean depression level decreased from -0.113 to -0.963, and precarity from -0.100 to -0.958. These changes represent reductions of approximately 90.8\% relative to their respective baseline magnitudes, indicating that effective intervention can push the system close to its lowest-risk equilibrium. -->

<!-- To explore these dynamics in more detail, we focused on two representative parameter sets: one with high feedback/input strength and low noise (a bistable candidate), and another with weak parameter interactions and high noise (a unistable candidate). In nonlinear dynamical systems, bistability refers to the existence of two distinct stable equilibria—meaning the long-term outcome depends on initial conditions or perturbations. -->

<!-- In our model, bistability emerges from an interplay between the strength of parameter-driven feedback/input loops and the magnitude of stochastic noise. The internal loop between $D$ and $P$ — governed by the product $\alpha_2 \times \beta_2$ — allows changes in one variable to reinforce the other. At the same time, the $\tanh(\cdot)$ nonlinearity in the system saturates for large inputs, flattening the system’s responsiveness. For multiple stable equilibria to emerge, the internal gain must be large enough to push the system into these nonlinear (saturated) regions. If this gain is too small, or if noise is too high, the system behaves more linearly and converges to a single stable point — a unistable regime.[^1]  -->

<!-- [^1]: This condition resembles the classical *pitchfork* bifurcation in discrete dynamical systems: when the gain $k$ in a scalar system such as $x_{t+1} = \tanh(k x_t)$ exceeds a critical threshold (e.g., $k>1$), the system admits three fixed points—two stable and one unstable—thus producing bistability. In our two-variable model, sufficiently high values of $\alpha_2 \times \beta_2$ similarly give rise to a self-reinforcing loop that sustains two attractor states. -->

<!-- This behavior is illustrated in the bottom panel of @fig-intervention. In the bistable model, we observe a bifurcation in final outcomes: as stress is reduced, the system undergoes a sharp transition from a high-$D$/$P$ regime to a low one—indicating a tipping point. In contrast, the unistable model shows a smooth, continuous response to increasing intervention, consistent with a single stable equilibrium. -->

<!-- These findings highlight how a system’s internal structure — including both nonlinear feedback gain and noise levels — fundamentally shapes its response to intervention. In bistable regimes, small or moderate interventions may have little effect unless they exceed a critical threshold sufficient to tip the system out of a high-risk state. In unistable systems, by contrast, outcomes improve more gradually and predictably with increasing support. -->

<!-- This distinction is particularly relevant when addressing persistent stressors associated with precariousness, such as unstable housing, employment insecurity, or chronic financial strain. In these contexts, the effects of external adversity may be reinforced internally through feedback loops between stress exposure and mental health. Such dynamics can stabilize harmful states and limit responsiveness to minor interventions—emphasizing the need for targeted, high-leverage support. -->

<!-- Recognizing whether a population’s mental health dynamics follow a bistable or unistable structure can guide policy design. For bistable systems, early or forceful interventions may be required to trigger lasting improvement. For unistable systems, more moderate or incremental measures may suffice. By tailoring strategies to the system’s underlying structure, interventions can be made more effective and context-sensitive—ultimately enhancing support for those facing precarity-driven mental health risks. -->

```{r}
#| label: fig-intervention
#| fig-cap: "**Top**: Final Mean D and P Across Intervention Sizes. *Mean trajectories of depression ($D$) and precarity ($P$) across varying intervention strengths, simulated over the top 30 parameter sets. Each line represents one parameter set and is colored by its combined feedback and input strength.* **Bottom**: Smoothed Density Heatmap of Final $D$ and $P$ for Two Parameter Sets. *One exhibits unistability (top two panels) and the other exhibits bistability (bottom two panels). Bistable dynamics reveal a sharp transition between distinct outcome regimes, while unistable dynamics show gradual, continuous change across the intervention gradient.*"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/intervention_plot2.pdf")
```

# Discussion

<!-- Summmary & Interpretation of Findings -->
This study combined constraint-based causal discovery and dynamical modeling to investigate how depression and socioeconomic precarity interact—and how their internal structure modulates responsiveness to intervention. Using cross-sectional data from the HELIUS study, we applied cycle-capable discovery algorithms to estimate plausible directional relationships among financial stress, multiple domains of precariousness, and depressive symptoms. By analyzing both aggregate constructs and individual variables, we aimed to capture patterns that may be masked in either extreme of abstraction (i.e., purely aggregate or disaggregated analyses).

The causal analysis consistently identified financial stress (*S.fin*) as a key plausible driver of depression, appearing across both aggregated and disaggregated models. At the symptom level, we found that sleep disturbance (*slp*), depressed mood (*dep*), guilt (*glt*), and anhedonia (*anh*) were especially sensitive to external stressors. Among these, *dep* emerged as a central hub: not only was it heavily influenced by other symptoms, it also exerted directional influence on social precarity (*P.soc*), suggesting a feedback loop in which worsening mood contributes to social disconnection, which in turn may exacerbate depressive symptoms. The CCI algorithm even suggested a possible cyclic relationship between dep and overall precarity, hinting at a structurally self-reinforcing mechanism.
*Slp* also played a key initiating role. It showed directed connections to both social and employment precarity and exhibited many outgoing edges within the symptom network. This suggests *slp* may function as an early warning signal or entry point into the depression–precarity cycle, making it a promising target for preventive intervention. In contrast, *P.soc* more often appeared as an effect rather than a cause—receiving arrows from symptoms like *dep* and *glt*, but rarely sending them. This asymmetry implies that social precarity may often reflect downstream consequences of worsening mental health rather than being its root cause.
These patterns offer actionable insights for intervention. Addressing symptoms like *slp* early may prevent escalation into broader depressive states or worsening precarity. Intervening on *dep*, as a central node with potential feedback effects, could also disrupt the self-reinforcing dynamics between mental health and social isolation.

To examine how such structures shape responsiveness to support, we developed a simplified two-variable nonlinear dynamical model capturing the feedback between depression and precarity, modulated by financial stress. Calibrated against empirical summary statistics from the HELIUS dataset, the model allowed us to simulate how these dynamics evolve under varying levels of stress reduction.
The simulations revealed that intervention effectiveness depends not just on the magnitude of external stressors, but on the internal configuration of the system itself. In bistable regimes—characterized by strong feedback and low noise—the system could remain stuck in a high-risk state unless an intervention crosses a critical threshold, tipping it into a healthier equilibrium. By contrast, unistable systems—driven by weaker coupling or greater noise—responded more gradually and proportionally to increasing support.
These dynamics carry practical policy implications. In bistable systems, even moderate interventions may have little effect unless they are forceful enough to tip the system toward a lower-risk state. This underscores the importance of early or intensive support when targeting self-reinforcing cycles of depression and precarity. Unistable systems, by contrast, are more responsive to incremental support, suggesting that sustained, moderate interventions may be sufficient in such contexts.

<!-- <!-- Interpretation of Findings --> 
<!-- Our findings highlight the significant role of recent financial stress (S.fin) as a potential causal factor for depression, consistently observed across both sum score and symptom-level analyses. This consistency underscores the profound impact of financial stressors on mental health. The symptom-level analysis further identified specific symptoms—particularly sleep disturbances (slp), depressed mood (dep), guilt (glt), and anhedonia (anh)—as being especially sensitive to external precarious conditions. -->

<!-- Among these symptoms, dep stands out as a distinct causal factor for precarity in both aggregated and disaggregated analyses, with CCI even suggesting a possible cyclic relationship between them. This implies that depression and precariousness may reinforce one another, potentially leading to a worsening cycle over time. Within the symptom network, dep also exhibits a high degree of connectivity, but with mostly incoming arrows, suggesting that as symptom dynamics intensify, dep is quickly impacted by multiple factors—particularly P.soc (social precarity), further fueling the depression-precariousness cycle. -->

<!-- Beyond dep, slp emerges as a critical symptom in this dynamic. It connects with both social and employment precariousness, while also playing a central role in symptom interactions, affecting other symptoms more than most others. This suggests that slp may act as an initiator within the depressive symptom system, potentially triggering or amplifying symptom cascades. -->

<!-- Among the precariousness factors, P.soc appears to be one of the most pivotal—not as a driver of symptoms, but rather as a consequence of depression. It maintains the strongest connectivity with symptoms, yet its directionality indicates that it is largely shaped by depressive symptoms rather than acting as a primary cause. Together, these patterns suggest a broader directional dynamic in which financial stress (S.fin) exacerbates depressive symptoms, which in turn feed into social precarity, perpetuating the cycle. -->

<!-- This directional insight suggests that interventions targeting specific depressive symptoms, particularly dep, may help prevent downstream effects on worsening precarious conditions. Alternatively, early intervention focusing on slp may be a promising strategy, as slp appears to act as a potential initiator or activator within the depressive symptom network. This positioning suggests that slp could serve as an early warning sign, offering a valuable target for preventive interventions aimed at breaking the depression-precariousness cycle before it intensifies. -->

<!-- Building on these insights, we developed a simplified computational model capturing the core mechanisms identified in the data: bidirectional feedback between depression and precarity, driven by external stress. Calibrated against empirical summary statistics, the model allowed us to explore how different parameter configurations—particularly feedback strength and stochastic noise—modulate the system’s response to simulated interventions. Some systems exhibited bistability, resisting change until pushed past a critical threshold, while others responded more gradually and predictably. These patterns have direct policy implications: in bistable regimes, intensive or early interventions may be required to shift individuals out of high-risk states, whereas in unistable systems, more moderate, sustained support may suffice. -->

<!-- A key contribution of this work is to show that the structure of the depression–precarity system matters: not all forms of adversity respond equally to intervention, and this heterogeneity may stem from internal feedback dynamics as much as from external conditions. Recognizing whether a given population operates within a bistable or unistable regime could inform the design and timing of mental health interventions—particularly in contexts of chronic socioeconomic stress. -->

<!-- Building on these findings, we developed a simplified two-variable nonlinear dynamical system to simulate the co-evolution of depression and social precarity under varying levels of financial stress. This model served as a conceptual abstraction of the inferred causal structure, enabling us to assess how the system’s internal configuration—particularly feedback strength, input coupling, and stochastic noise—modulates its behavior in response to intervention. -->

<!-- Simulation results revealed that system structure critically influences responsiveness to intervention. In regimes with strong internal feedback loop and low noise, the model exhibits bistability: it can remain in a high-risk state unless a sufficiently large intervention tips it into a lower-risk configuration. By contrast, systems with weaker interactions or higher noise display unistable dynamics, where outcomes improve more gradually and proportionally with increasing intervention strength. -->

<!-- Limitations -->
That said, our study has several limitations. While the FCI and CCI algorithms are designed to recover causal structure from cross-sectional data—including feedback and latent confounding—many edge directions in our results remain ambiguous, as indicated by the prevalence of circle endpoints and bidirectional arrows, especially in weakly connected regions of the graph. These uncertainties were most evident in the links between individual symptoms and precariousness factors, where edge estimates varied across algorithms and conditional independence (CI) tests. This likely reflects limited statistical signal in high-dimensional settings, where large conditioning sets reduce power and small effect sizes are harder to detect. Although the CCI algorithm is explicitly designed to detect cycles, we found little clear evidence of feedback: most candidate loops appeared only as bidirectional edges, which may reflect latent confounding or statistical imprecision rather than true cyclic structure. This could point to either an absence of strong feedback at the symptom level or limitations in the sensitivity of full-graph discovery approaches applied to densely interconnected variables. More focused strategies, such as local structure tests or theory-guided subgraph modeling, may help resolve these ambiguities. In addition, while our analyses relied solely on observational data, some structural uncertainties—particularly those involving latent confounding or bidirectional influence—may ultimately require interventional or semi-interventional designs. Extensions such as LLC [@hyttinen2012learning], NODAGS-Flow [@sethuraman2023nodags], or Bicycle [@rohbeck2024], which integrate observational and interventional data, may prove valuable where such data are available.

Second, our findings showed sensitivity to the choice of conditional independence (CI) test. The Gaussian CI test, which assumes linear relationships and Gaussian noise, often produced denser graphs—likely due to its tendency to overestimate weak or noisy linear associations. In contrast, RCoT, a nonparametric alternative, avoids these assumptions but relies on a Gaussian radial basis function (RBF) kernel to assess dependence. While flexible in capturing nonlinear relationships, this kernel is optimized for smooth, continuous variables and may underperform with discrete or ordinal data. Its similarity estimates can be unreliable when applied to variables with discontinuities, potentially leading to missed dependencies—particularly relevant in the HELIUS dataset, where some variables are ordinal or mixed-type [@howlett2001radial]. To address this, future work should consider adapting kernel-based CI tests to better handle mixed data structures—for example, by employing hybrid kernels that explicitly combine continuous and discrete similarity measures. This could improve the reliability of nonparametric causal discovery in settings with heterogeneous variable types.

Finally, the structure of the dynamical model introduces its own limitations. While the model was designed to reflect key features of the discovered causal structure, it represents a highly simplified system: only two variables—depression and social precarity—modulated by external financial stress. This abstraction allowed for tractable analysis and visual intuition, particularly around nonlinear behaviors such as bistability, but omits other potentially influential constructs highlighted in the causal graphs, including relational stress, employment conditions and housing conditions. Moreover, the model’s nonlinearity—introduced via the hyperbolic tangent function—was chosen for convenience rather than empirical justification. While this function supports saturation and the emergence of tipping points, the presence of bistability is a product of simulation, not empirical verification. It remains possible that simpler, linear dynamics could account for the observed data just as well. Future work should further investigate whether bistability is a necessary feature of depression–precarity dynamics, and explore alternative functional forms—both linear and nonlinear—using richer data and stronger empirical constraints.

<!-- Several limitations should be acknowledged when interpreting these findings. The prevalence of bidirectional arrows and circle-marked endpoints in the graphs reflects unresolved ambiguities in the causal relationships suggested by the data. These uncertainties highlight the need for further research, particularly with higher-resolution datasets, such as time-series data.  Incorporating time-series data could capture temporal dynamics between precarity factors and depressive symptoms, providing time-specific insights and revealing how these relationships evolve over time. Methods like *PCMCI* [@runge2019detecting], *tsFCI* [@entner2010causal], and other time-series adaptations of causal discovery algorithms could offer a more robust framework for addressing temporal dependencies and refining the analysis. -->
<!-- Another notable limitation is the absence of clear evidence for cycles within the symptom network, despite employing algorithms designed to account for cyclic relationships. The CCI algorithm primarily produced bidirectional arrows, while FCI predominantly generated directional arrows, yet neither demonstrated patterns indicative of definitive cyclic structures. This suggests that addressing cyclic relationships is particularly challenging when relying solely on observational datasets. Future research could benefit from refined datasets that incorporate intervention data. Methods such as *LLC* [@hyttinen2012learning], *NODGAS-Flow *[@sethuraman2023nodags], and the recently developed *Bicycle* algorithm [@rohbeck2024], which utilize both observational and intervention data, may uncover potential cycles more effectively.  -->
<!-- Lastly, regarding conditional independence (CI) testing, the differences between the graphs generated by Gaussian CI and RCoT underscore the methodological sensitivities in detecting causal relationships. Gaussian CI produced denser graphs, which is somewhat counterintuitive, as the Gaussian CI's strict linearity assumption would typically result in fewer detected relationships, not more. -->
<!-- A possible explanation for this discrepancy is that Gaussian CI's reliance on partial correlations may overestimate relationships when specific non-linear dependencies exist in the data. In contrast, RCoT, free from such assumptions, might better capture independence patterns under such conditions. -->
<!-- However, RCoT is not without its limitations. While technically non-parametric, its performance can be influenced by the distributional characteristics of the data. The Gaussian RBF (radial basis function) kernel, optimized for continuous data with smooth transitions, may struggle with discrete or mixed datasets, where distances between discrete points may fail to convey meaningful similarity [@howlett2001radial]. This limitation could explain why RCoT might miss certain dependencies, particularly in the HELIUS dataset, where some variables lack smooth continuity. -->
<!-- Future research could tackle these challenges by developing hybrid kernels designed for mixed datasets, integrating both continuous and discrete variables into RCoT-like methods. This would have the potential to improve the reliability of findings, especially when working with datasets characterized by complex and heterogeneous structures. -->

<!-- A key limitation of our study lies in the structure and assumptions of the dynamical model. While the model is grounded in causal patterns identified from the data, it is intentionally simplified, focusing solely on two interacting variables—depression and social precarity—modulated by financial stress. This abstraction facilitates tractable simulation and exploration of systemic behaviors such as bistability. However, it excludes several factors identified in the causal graph as potentially influential, including relational stress, housing conditions, and other contextual or protective variables like coping strategies or social support. These omissions constrain the model’s ability to represent the multifaceted nature of real-world dynamics. -->
<!-- Additionally, the model’s nonlinearity—implemented via a hyperbolic tangent function—was chosen for its analytical properties, such as bounded outputs and saturating responses, rather than empirical grounding. While such nonlinearities are necessary to produce bistable dynamics, the presence of bistability in our model is a product of simulation rather than empirical verification. We do not currently have time-series data to confirm the existence of tipping points or regime shifts in real-world trajectories of depression and precarity. It is also possible that a linear model would suffice to capture much of the observed behavior, and that the current structure overcomplicates model formulation. Future work should seek for securing stronger evidence for bistability in the system and empirically test for the nonlinearity and compare model variants—linear and nonlinear—against richer longitudinal data to validate which formulations best capture the underlying dynamics. -->

<!-- The numerous bidirectional arrows in the CCI outputs also imply the presence of latent confounders. Expanding the analysis to include additional variables, such as other mental health conditions and coping mechanisms, could provide a more comprehensive understanding of the factors influencing depressive symptoms. Additionally, integrating qualitative approaches could enrich these findings, offering contextual insights into the quantitative relationships identified. Understanding individuals' lived experiences of precarity and depression could inform the development of more tailored and effective interventions. -->

<!-- However, it could be also that RCoT performed somewhat suboptimal. while RCoT is technically non-parametric, its performance can still be influenced by the distributional characteristics of the data. Specifically, the Gaussian RBF (radial basis function) kernel, optimized for continuous data with smooth transitions, may struggle to capture relationships in datasets with discrete or mixed distributions. In such cases, the distances between discrete points may fail to convey meaningful similarity information. As a result, RCoT might miss certain dependencies, particularly when variables in the dataset lack smooth continuity like in our HELIUS dataset. -->

<!-- Future research could address these issues by exploring a broader range of CI testing approaches. One traditional approach to handle this is discretizing variables and use $G^2$ test, which may better capture dependencies in non-continuous datasets [@dojer2016learning; @neapolitan2004learning]. A more promising direction, however, lies in the development of hybrid kernels tailored for mixed datasets, effectively integrating both continuous and discrete variables into RCoT-like methods. -->
<!-- By systematically employing and comparing a wider variety of CI testing techniques, researchers could gain more robust insights and mitigate the limitations inherent in specific approaches. This broader exploration holds the potential to enhance the reliability of findings, particularly in datasets with complex and heterogeneous structures. -->

<!-- Implications -->
<!-- 1. still impressive -- symptom directions aligns much well with the other empirical studies. -->
<!-- 2. understood better which aspect of precarity is more important for depression and we can target depends on what level of intervention we are planning. -->

<!-- Implications -->
Our results reveal that the effectiveness of interventions targeting depression and precarity is shaped not only by the strength of external stressors but also by the internal structure of the system. The interplay between feedback strength and noise determines whether a system responds gradually or only after crossing a tipping point—highlighting the need for intervention strategies attuned to these underlying dynamics.

These findings underscore the importance of considering internal feedback—not just external adversity—when designing interventions. Mental health systems are not always passively responsive to support; their structure can constrain how much impact any given policy has. By diagnosing these structural features—through tools like causal discovery and computational modeling—researchers and policymakers can better tailor interventions to the underlying mechanics of vulnerability.

Beyond these specific findings, this study offers a generalizable framework for studying complex mental health phenomena. Causal discovery enables principled hypothesis generation from observational data, while dynamical modeling helps test how those structures behave under intervention. Together, these methods move us from descriptive patterns to mechanistic insight, with practical relevance for real-world mental health policy.

Future work could build richer dynamical models grounded in higher-resolution empirical data—potentially incorporating time, interventions, or more granular psychosocial constructs. Such extensions would help validate whether mechanisms proposed here, such as bistability, manifest in real-world trajectories. Ultimately, we hope this study encourages further integration of causal structure learning and system-level modeling to improve mental health outcomes under conditions of socioeconomic precarity.

# References

::: {#refs}
:::

# Appendix {#sec-appendix}

## HELIUS study

The HELIUS (HEalthy LIfe in an Urban Setting) study is a large-scale, multiethnic cohort study conducted in Amsterdam. Participants were randomly selected from the municipality register and stratified by ethnic origin to ensure balanced representation across six major groups. Invitation letters were sent by mail, followed by a reminder after two weeks. Non-respondents were contacted via home visits where applicable.

Of those invited, approximately 55% responded (Dutch: 55%, Surinamese: 62%, Ghanaian: 57%, Turkish: 46%, Moroccan: 48%). Among those contacted, 50% agreed to participate (Dutch: 60%, Surinamese: 51%, Ghanaian: 61%, Turkish: 41%, Moroccan: 43%), resulting in an overall participation rate of 28%.

Participants completed either a digital or paper version of the questionnaire, with assistance offered when needed, and received a confirmation letter for a physical examination appointment.

At baseline (2011–2015), 24,780 individuals were enrolled. Of these, 23,936 participants completed the questionnaire that included items relevant to precariousness. After excluding individuals with missing data on key variables, the final analytical sample consisted of 21,628 participants.

The HELIUS study received approval from the Medical Ethics Committee of the Academic Medical Center (AMC), and written informed consent was obtained from all participants prior to enrollment.

## Causal Discovery Primer

As shown in Table 1, the resulting graphs from FCI and CCI differ slightly (*PAG*: partial ancestral graph; *MAAG*: maximal almost ancestral graph) due to their reliance on different underlying assumptions. 
Despite these differences, both graphs belong to the class of *ancestral graphs*, which are designed to encode causal relationships between variables, where the presence of an edge indicates causal *ancestry.*
In these graphs, directed edges, $A \stararrow B$, indicate that $B$ is not an ancestor of $A$ in every graph within the Markov equivalence class, $Equiv(G)$. 
The Markov equivalence class represents a set of graphs that encode the same conditional independence relationships, ensuring that the same *d-separation* conditions hold across all graphs in the class [@spirtes2001causation].
Conversely, an edge marked as $A \startail B$ indicates that $B$ is an ancestor of $A$ across all graphs in $Equiv(G)$. 
Circle endpoints, $A \starcirc B$, represent ambiguity in the ancestral relationship, meaning $B$'s ancestral status relative to $A$ varies across graphs in $Equiv(G)$.[^4] 
Finally, when an edge is represented as $A \arrowarrow B$, it implies that neither $A$ nor $B$ is an ancestor of the other, suggesting the presence of a latent confounder influencing both variables.

\begin{table}[ht]
\centering
\small % Make the font size smaller
\caption{Assumptions of causal discovery algorithms}
\begin{tabular}{lccccc}
\toprule
Algorithm & Acyclicity & \makecell{Causal \\ sufficiency} & 
\makecell{Absence of \\ selection bias} & Linearity & Output \\ 
\midrule
PC  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & CPDAG \\ 
FCI & $-^{a}$ & $\checkmark$ & $-^{a}$ & $-^{a}$ & PAG \\ 
CCI & x & x & x & $\checkmark$ & \makecell{\footnotesize(partially oriented) \\ MAAG} \\ 
\bottomrule
\end{tabular}
\caption*{\footnotesize{\textit{Note}. $^{a}$The FCI algorithm, introduced by Spirtes (1995), is a constraint-based causal discovery method for DAGs that accounts for latent confounding and selection bias. Mooij \& Claassen (2020) later showed its applicability to cyclic causal discovery with latent confounding under general faithfulness and Markov conditions, assuming non-linear causal relationships.}}
\end{table}

<!-- For word docx compiling -->
<!-- **Table 1**: Assumptions of causal discovery algorithms. -->

<!-- | Algorithm | Acyclicity | Causal sufficiency | Absence of selection bias | Linearity | Output | -->
<!-- |-----------|------------|--------------------|---------------------------|-----------|--------| -->
<!-- | PC        | ✓          | ✓                  | ✓                         | ✓         |  CPDAG | -->
<!-- | FCI       | -¹         | ✓                  | -¹                        | -¹        |  PAG   | -->
<!-- | CCI       | x          | x                  | x                         | ✓         |  MAAG  | -->
<!-- ------------------------------------------------------------------------------------------------ -->
<!-- ¹FCI can detect cycles in systems that lack selection bias and exhibit non-linear relationships (Mooij & Claassen, 2020). -->

The algorithms also differ in how they detect cycles and represent cyclic relationships in their resulting graphs.
In CCI's MAAG, $A$ \textemdash\ $B$ indicates that $A$ is an ancestor of $B$, and simultaneously $B$ is an ancestor of $A$, referring to a cyclic relationship between $A \stackedarrows B$. 
FCI, on the other hand, identifies potential cycles more subtly. In its PAG, fully-connected nodes with circle endpoints ($\circirc$) may suggest the presence of cyclic structures.
FCI, therefore, provides a sufficient condition to distinguish variables that are not part of a cycle, offering a more nuanced approach to handling cyclic relationships [@mooij2020constraint].

Table 1 highlights further differences in the assumptions underlying these algorithms. CCI operates under the assumption of a linear system, whereas FCI, particularly when used to infer cyclic relationships, assumes a non-linear system without selection bias and adheres to more general faithfulness and Markov conditions (i.e., *$\sigma$-separation* and *$\sigma$-faithfulness* setting) [@forre2018constraint].
When the respective assumptions of each algorithm are met, their inferred cyclic relationships align with those illustrated in @fig-examplegraphs.

While CCI's MAAG provides an explicit representation of cyclic relationships, it has certain theoretical limitations, as the MAAG may not always fully preserve d-separation relations from the original graph, $\mathcal{G}$ [@strobl2019]. Given these considerations, we examine and compare results from both algorithms, placing more weight on the FCI results, which are presented in the main results section, while discussing CCI results where relevant, with detailed findings provided in the [Appendix](#sec-appendix).
A full discussion of causal discovery concepts and algorithmic details is beyond the scope of this paper; however, readers seeking a more in-depth understanding can refer to @park2024discovering for a comprehensive exploration of these methods and their applications.

[^4]: $*$ serves as a \textit{meta-symbol}, representing one of the three possible edge-endpoints. For instance, $A \tailstar B$ can indicate any of the following edges: $A$ \textemdash\ $B$, $A\tailarrow B$, or $A \tailcirc B$ [@park2024discovering].

```{r}
#| layout-ncol: 3
#| label: fig-examplegraphs
#| fig-cap: Example graph $\mathcal{G}$ featuring a cycle and the corresponding MAAG from CCI and PAG from FCI.
#| fig-subcap: 
#|   - Example graph $\mathcal{G}$
#|   - Corresponding MAAG 
#|   - Corrsponding PAG
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/original_DCG.png")
knitr::include_graphics("img/FCI_graph.png")
knitr::include_graphics("img/CCI_graph.png")
```

The graph produced by the PC algorithm is a *CPDAG* (completed partially directed acyclic graph), where directed edges ($A$ → $B$) indicate that $A$ is a direct cause (parent) of $B$. Unlike FCI and CCI, the CPDAG does not include circle symbols. Instead, when the PC algorithm cannot determine the direction of causality, it represents this uncertainty with bidirectional arrows. 
While the PC algorithm serves as a useful reference, its strict assumptions — acyclicity and the absence of latent confounders—limit its applicability in more complex scenarios. For this reason, our primary focus remains on the results obtained from FCI and CCI, with all PC algorithm results provided in the [Appendix](#sec-appendix) for completeness.

## Precariousness factors by Leonie

1.  EMPLOYMENT PRECARIOUSNESS

-   `H1_Arbeidsparticipatie`: Working status
-   `H1_WerkSit`: Which work situation most applies to you?
-   `H1_RecentErv8`: Experiences past 12 months: h. You were sacked from your job or became unemployed (*reverse*)

2.  FINANCIAL PRECARIOUSNESS

-   `H1_InkHhMoeite`: During the past year, did you have problems managing your household income?
-   `H1_RecentErv9`: Experiences past 12 months: i. You had a major financial crisis (*reverse*)

3.  HOUSING PRECARIOUSNESS

-   `veilig_2012`: Score safety (veiligheid) in 2012 (*reverse*)
-   `vrz_2012`: Score level of resources (niveau voorzieningen) in 2012 (*reverse*)
-   `P_HUURWON`: Percentage Huurwoningen

4.  CULTURAL PRECARIOUSNESS

-   `H1_Discr_sumscore`: Perceived discrimination: sum score of 9 items (range 9-45)
-   `H1_SBSQ_meanscore`: Health literacy: SBSQ meanscore (range 1-5) (*reverse*)
-   `A_BED_RU`: Aantal bedrijfsvestigingen; cultuur, recreatie, overige diensten (*reverse*)

5.  SOCIAL PRECARIOUSNESS

-   `H1_RecentErv5`: Experiences past 12 months: e. Your steady relationship ended (*reverse*)
-   `H1_RecentErv6`: Experiences past 12 months: f. A long-term friendship with a good friend or family member was broken off (*reverse*)
-   `H1_RecentErv7`: Experiences past 12 months: g. You had a serious problem with a good friend or family member, or neighbour (*reverse*)
-   `H1_SSQT`: SSQT (frequency of social contact): sum score of 5 items (range 5-20) (*reverse*)
-   `H1_SSQSa`: SSQS (adequacy of social contact): sum score of 5 items, category 3 and 4 not combined (range 5-20) (*reverse*)

```{r, echo=FALSE, fig.align='center', fig.width=15}
# check correlation
#cor(df) |> round(2)

# visualize corr
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(cor(scaled_data[,c(1:16, 26, 27)]), method="color", col=col(200),
         type="upper", order="hclust",
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", number.cex=0.50, tl.srt=45,tl.cex=0.7, #Text label color and rotation
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
         )
```

-   **High** Correlations: `emp_stat` (employment status) and `work_sit` (work situation) have a strong positive correlation of 0.82. This suggests that individuals with higher employment status tend to have more secure or favorable work situations. `soc_freq` (social contact frequency) shows a strong positive correlation with `soc_adq` (social adequacy) at 0.61. This indicates that individuals with more frequent social contact also tend to have higher perceived adequacy of social interactions.

-   **Moderate** Correlations: `nb_safe` (neighborhood safety) and `nb_res` (resources) have a moderate positive correlation of 0.39, suggesting that areas with higher safety also have better resources. `hea_lit` (health literacy) has moderate correlations with `emp_stat` (0.26) and `work_sit` (0.25), which could mean that higher health literacy is associated with better employment situations. `frd_brk12` (friendship breakups) and `conf12` (conflicts) have a notable correlation of 0.43, indicating a relationship between having conflicts and friendship losses.

-   **Low to Moderate** Correlations in Financial Precariousness: `inc_dif` (income difficulties) has a moderate correlation with `fincri12` (financial crisis) at 0.49. This aligns with the expected relationship, where individuals who experience general income difficulties are more likely to report financial crises.

-   **Low** Correlations (0.1 - 0.2): Many variables, such as `discrim` (discrimination), `unemp12` (unemployment experience), and `rel_end12` (relationship end), have low correlations with other variables, suggesting relatively independent relationships in the context of this dataset.

## Exploratory Factor Analysis (EFA)

```{r, echo=FALSE, fig.align='center', fig.height=5, out.width="55%", message=FALSE, results='hide'}
# precarious factors
df <- scaled_data[,1:16]

# check suggested number of factors
pl <- fa.parallel(df, fa = "fa", n.iter = 100, show.legend = FALSE)
# run EFA
efa_result <- fa(df, nfactors = 5, rotate = "promax") # oblique rotation
#print(efa_result)
```

```{r, echo=FALSE, fig.align='center', out.width="70%"}
# Create a data frame from the pattern matrix
loadings <- data.frame(
  Variable = c("emp_stat", "work_sit", "unemp12", "inc_dif", "fincri12", "nb_safe", "nb_res", "nb_rent",
               "discrim", "hea_lit", "cul_rec", "rel_end12", "frd_brk12", "conf12", "soc_freq", "soc_adq"),
  MR1 = c(0.98, 0.97, 0.17, 0.21, 0.11, -0.08, 0.02, -0.03, -0.07, 0.27, -0.03, -0.04, -0.09, -0.08, -0.02, -0.03),
  MR2 = c(-0.06, -0.08, -0.03, 0.09, 0.05, -0.08, -0.05, -0.06, 0.17, 0.08, -0.08, -0.04, -0.04, 0.00, 1.00, 0.77),
  MR3 = c(-0.01, -0.06, 0.26, 0.28, 0.47, -0.05, -0.04, -0.02, 0.20, -0.10, -0.05, 0.36, 0.63, 0.56, -0.14, -0.01),
  MR4 = c(-0.02, -0.03, -0.04, 0.06, 0.03, -0.02, 0.71, -0.09, 0.13, 0.04, 0.61, -0.03, -0.01, -0.02, -0.04, -0.12),
  MR5 = c(-0.14, -0.11, -0.03, 0.16, 0.09, 0.61, -0.21, 0.54, 0.10, 0.15, 0.04, -0.01, -0.06, -0.06, -0.10, -0.13)
)

# Convert to long format
loadings_long <- loadings %>%
  pivot_longer(cols = starts_with("MR"), names_to = "Factor", values_to = "Loading")

# Plot heatmap
ggplot(loadings_long, aes(x = Factor, y = Variable, fill = Loading)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, limits = c(-1, 1)) +
  theme_minimal() +
  labs(title = "Factor Loadings Heatmap", x = "", y = "")
```

### Factor Loadings (Pattern Matrix)

-   **MR1**: High loadings on `emp_stat` and `work_sit` suggest this factor captures *employment* precariousness.
-   **MR2**: Strong loadings on `soc_freq` and `soc_adq` indicate *social* precariousness.
-   **MR3**: Key items like `frd_brk12`, `conf12`, and `fincri12`, suggest recent *stressful events*.
-   **MR4**: High loadings on `nb_res` and `cul_rec` may reflect *community resources* precariousness.
-   **MR5**: Variables `nb_safe` and `nb_rent` with high loadings indicate *housing* precariousness.

### Variance Explained

The factors cumulatively explain **38%** of the variance, with MR1 being the most influential factor. Each factor contributes a smaller proportion to the total variance (MR1 at 12%, MR2 at 9%, etc.).

### Factor Intercorrelations

Factors are moderately correlated, especially between *MR1 and MR5*, and *MR2 and MR3*. This indicates that while distinct, these factors are related—reasonable in a complex socio-economic context.

### Model Fit Statistics

RMSEA (0.071) suggest an acceptable fit. Tucker Lewis Index (0.802) suggests moderate reliability for the model.

### Summary

The 5-factor model appears interpretable and captures distinct dimensions of precariousness: *employment, social, stressors, community resources, and housing precariousness*. Although the overall fit and explained variance could be stronger, these factors offer insights into the underlying structure of the data, highlighting key areas of precariousness.

## PCA

```{r, echo=FALSE, fig.align='center', out.width="70%"}
# run PCA
res.pca <- prcomp(df, scale = TRUE)
# print(res.pca)
#summary(res.pca)

# Visualize eigenvalues (scree plot)
fviz_eig(res.pca, addlabels = TRUE)
#
# fviz_pca_var(res.pca,
#              col.var = "contrib", # Color by contributions to the PC
#              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#              repel = TRUE     # Avoid text overlapping
#              )
```

-   Component Retention: The scree plot shows a clear "elbow" after the first component. This steep drop suggests that most variance is explained by the first component. After Dimension 5, the percentage of explained variance decreases slightly more gradually, indicating diminishing returns for adding more components. If we need to choose multiple components, retaining the first 5 components seems reasonable, as they capture most of the variance (cumulatively explaining about 54.7% of the total variance).

```{r, echo=FALSE, fig.align='center', fig.height=4.5, fig.width = 8}
# Results for Variables
res.var <- get_pca_var(res.pca)
# res.var$coord |> round(2)         # Coordinates
# res.var$contrib |> round(2)        # Contributions to the PCs
# res.var$cos2           # Quality of representation

# corrplot(res.var$cos2, is.corr=FALSE)

res.var$contrib |> as.data.frame() |> dplyr::select(1:5) |>
  tibble::rownames_to_column(var = "var") |> # adds a column for rownames
    pivot_longer(cols = starts_with("Dim"),
               names_to = "Dimension",
               values_to = "Contribution") |>
# Plot with facet_wrap to show each dimension in a separate facet
ggplot(aes(x = reorder(var, -Contribution), y = Contribution, fill = Contribution)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ Dimension, scales="free_y") +
  coord_flip() +
  labs(title = "Contribution of Variables to 5 Principal Components",
       x = "", y = "Contribution (%)") +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```

### Explained variance (contributions) of variables

It shows the importance of variables within each component.

-   **Dim1**: High contributions are observed from `emp_stat`, `work_sit`, `inc_dif`, and `fincri12`, suggesting that this dimension captures aspects of *employment and financial* security.

-   **Dim2**: While `emp_stat` and `work_sit` overlap with Dim1, the strong contributions from `frd_brk12` and `rel_end12` indicate that this dimension captures a focus on *recent relationship stressors*.

-   **Dim3**: `cul_rec`, `nb_res` have the highest contributions, indicating this dimension likely represents *community and cultural* factors.

-   **Dim4**: `soc_freq` and `soc_adq` stand out in this dimension, suggesting an emphasis on *social* precariousness.

-   **Dim5**: `nb_safe` and `nb_rent` are the top contributors, pointing to *housing* security as key themes in this component.

### Cos² Values

Cos² (squared cosine) values, or the quality of representation, show how well each variable is represented by each dimension. where higher cos² values (closer to 1) indicate better representation of a variable by a component.

```{r, echo=FALSE, fig.align='center', out.width="70%"}
res.var$cos2 |> as.data.frame() |> dplyr::select(1:5) %>%
  mutate(var = rownames(.)) |>
  pivot_longer(!var, names_to = "dim", values_to = "cos2") |>
  ggplot(aes(x = dim, y = var, fill = cos2)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue") +
  labs(title = "Cos2 of Variables on Each Component (1-5)", fill = "Cos2", x = "", y = "") +
  theme_minimal()
```

-   **Dim.1**: Variables `emp_stat`, `work_sit`, `inc_dif`, and `fincri12` show high cos² values, meaning that PC1 primarily captures variations in employment and financial difficulties. This component could represent *employment & finance* precariousness.

-   **Dim.2**: Variables `work_sit`, `emp_stat`, `frd_brk12`, and `conf12` are well-represented in this component, suggesting PC2 captures aspects of *recent relationship stressors*.

-   **Dim.3**: Variables `nb_res` and `cul_rec` load strongly on PC3. This may represent community or cultural resources, indicating that this component is associated with *neighborhood resources*.

-   **Dim.4**: This component has high cos² values for `nb_res`, `cul_rec`, `soc_freq`, and `soc_adq.` While `nb_res` and `cul_rec` are also prominent in PC3, PC4 uniquely captures nuanced differentiation in *social* precariousness.

-   **Dim.5**: `nb_safe` and `nb_rent` are well-represented by PC5. This component might capture *housing* precariousness.

## ICA

```{r, echo=FALSE,  fig.align='center', out.width="70%"}
# Load the library
library(fastICA)
# set the seed
set.seed(1)

# Set the number of components you want to extract
n_components <- 5

# Run ICA
ica_model <- fastICA(df,
  n.comp = n_components, alg.typ = "parallel",
  fun = "exp", method = "C", maxit = 5000
)
# glimpse(ica_model)

weight_matrix <- data.frame(t(ica_model$A))
names(weight_matrix) <- stringr::str_glue("IC{seq(weight_matrix)}")

# dist_matrix <- dist(as.matrix(weight_matrix))
# clust <- hclust(dist_matrix)
# plot(clust)

weight_matrix |>
  mutate(var = colnames(df)) |>
  pivot_longer(
    cols = starts_with("IC"),
    names_to = "IC", values_to = "loading"
  ) |>
  ggplot(aes(x = IC, y = var, fill = loading)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, limits = c(-1, 1)) +
  labs(x="", y="") +
  theme_minimal()
```

### Dominant Variables per Component:

For each Independent Component (IC), we can identify variables with *high absolute* values in each column. These values indicate that the IC captures a strong, independent signal associated with these variables.

-   **IC1**: `soc_freq` and `soc_adq` have strong negative loadings on this component, indicating that this component might represent *social* precariousness.

-   **IC2**: `frd_brk12`, `conf12`, `rel_end12`, `fincri12` and `unemp12` have the most substantial loadings on this component, all with negative signs. This might point to a *recent relational or social stressor* component.

-   **IC3**: `nb_res` and `cul_rec` show notable negative loadings, pointing to a focus on *community resource* precariousness.

-   **IC4**: High loadings for `nb_safe`, `nb_rent`, `nb_res`, `cul_rec`, and `discrim` suggest a theme of *housing and community-based* precariousness, reflecting both safety and social challenges within the neighborhood context.

-   **IC5**: `emp_stat` and `work_sit` both have strong negative loadings on this component, suggesting it captures *employment* precariousness.

## Hierarchical clustering

### Using Euclidean distance

-   Ward.D's method: Minimizes the variance within clusters, producing more compact and spherical clusters.
-   Single linkage: Groups clusters based on the minimum distance between points.
-   Complete linkage: Groups clusters based on the maximum distance between points.
-   Average linkage: Uses the average distance between all pairs of points in the two clusters.

```{r, warning=FALSE, message=FALSE, fig.height=7, echo=FALSE, fig.align='center'}
#| layout-ncol: 2
library(dendextend)  # For coloring the clusters
library(ggdendro)

# Compute the distance matrix between variables
distance_matrix <- dist(t(df))

# wardD
hc_ward <- hclust(distance_matrix, method = "ward.D")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)  # Specify the number of clusters
plot(hc_ward, main = "ward.D")
rect.hclust(hc_ward, k = 5, border = 2:6)

# Avg
hc_avg <- hclust(distance_matrix, method = "average")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)
plot(hc_avg, main = "average")
rect.hclust(hc_avg, k = 5, border = 2:6)

hc_comp <- hclust(distance_matrix, method = "complete")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)
plot(hc_comp, main = "complete")
rect.hclust(hc_comp, k = 5, border = 2:6)

hc_sing <- hclust(distance_matrix, method = "single")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)  # Specify the number of clusters
plot(hc_sing, main = "single")
rect.hclust(hc_sing, k = 5, border = 2:6)
```

#### Consistent Groupings (Across All or Most Methods)

-   `emp_stat` and `work_sit`: This pair consistently clusters together across all linkage methods, suggesting that they are closely related variables, likely capturing a similar aspect of the data (possibly employment status or employment-related information).

-   `nb_safe`, `nb_res`, and `nb_rent`: These variables are often grouped closely in several methods (especially Ward.D, average, and complete linkage). This suggests a similarity or common theme among them, potentially related to neighborhood or housing precariousness.

-   `soc_freq` and `soc_adq`: These two variables frequently cluster together, indicating they likely measure aspects of social frequency and adequacy in similar ways. They appear together in Ward.D, average, and complete linkage.

-   `frd_brk12` and `conf12`: These variables are often clustered closely (though they sometimes join with other variables like rel_end12), suggesting they may capture aspects of relationship or social conflict. This pair appears in close proximity, especially in average and Ward.D.

#### Inconsistent Groupings (Variability Across Methods)

-   `hea_lit`: This variable shows inconsistent clustering across methods. In Ward.D, it joins with `fincri12`, while in other methods, it’s often more isolated or grouped with variables that do not appear similar. This may suggest that `hea_lit` does not strongly correlate with other variables, or it has multidimensional aspects affecting its grouping across methods.

-   `discrim`: This variable also shows variable groupings. In Ward.D, it is grouped with `hea_lit`, while in other methods (e.g., complete and single linkage), it clusters differently, sometimes on its own. This variability may indicate that `discrim` has weaker associations with the main clusters in the data or overlaps partially with multiple clusters.

-   Social and Financial Variables (`inc_dif`, `fincri12`, `unemp12`): These variables appear together in some methods (e.g., Ward.D clusters `fincri12` and `inc_dif`), but in others, they are spread out. This inconsistency suggests that social and financial variables may not have strong or consistent ties across different methods, perhaps due to capturing different aspects of precariousness.

#### Summary

The consistent clusters are likely capturing distinct thematic dimensions of the data (e.g., employment, housing, social contact), while the inconsistent variables may reflect multifaceted or weakly correlated attributes that do not fit neatly into one cluster.

### Using Mutual Information

Using mutual information (MI) as a basis for hierarchical clustering differs from using traditional distance measures (like Euclidean distance) in a few key ways.

```{r echo=FALSE, fig.align='center', fig.height=6, out.width="60%"}
library(infotheo)  # for mutual information functions

# Calculate pairwise mutual information
mi_matrix <- matrix(NA, ncol = ncol(df), nrow = ncol(df))
colnames(mi_matrix) <- rownames(mi_matrix) <- colnames(df)

for (i in 1:ncol(df)) {
  for (j in i:ncol(df)) {
    # Calculate mutual information between pairs of variables
    mi_matrix[i, j] <- mutinformation(discretize(df[, i]), discretize(df[, j]))
    mi_matrix[j, i] <- mi_matrix[i, j]  # symmetric
  }
}

# Normalize mutual information matrix
max_mi <- max(mi_matrix, na.rm = TRUE)
mi_distance <- 1 - (mi_matrix / max_mi)

# Convert to distance object
dist_matrix_mi <- as.dist(mi_distance)

# Hierarchical clustering on MI-based distance
hc_mi <- hclust(dist_matrix_mi, method = "ward.D2")

# Plot the dendrogram
plot(hc_mi, main = "Hierarchical Clustering (Mutual Information Distance)", xlab = "", sub = "")
rect.hclust(hc_mi, k = 5, border = 2:6)

```

#### Comparison to Euclidean Distance Clustering

-   **Housing and Community** Cluster: The variables `nb_safe`, `nb_res`, `nb_rent`, and `cul_rec` cluster together, indicating a strong association among housing-related and community-based factors. This suggests a shared theme of housing or community precariousness. This grouping is also observed in the Euclidean-based clustering, but it appears more tightly connected here, potentially due to the non-linear relationships highlighted by mutual information.

-   **Employment and Social Support** Cluster: `emp_stat` and `work_sit` form a cluster, linking employment status and work situation together as they did in Euclidean-based clustering. These remain closely associated regardless of the distance metric used. `soc_freq` and `soc_adq`, related to social contact frequency and adequacy, cluster nearby, indicating they have a stronger non-linear relationship with employment variables. This is a subtle difference as Euclidean distance might not capture this association as effectively.

-   **Financial Stressor** Cluster: `inc_dif` and `fincri12`, representing income difficulties and recent financial crises, consistently cluster together in both approaches, showing a strong association, likely linear. However, mutual information-based clustering links these financial stressors with social support variables, suggesting that financial challenges may have complex dependencies with social support in this dataset.

-   **Relational Stressor** Cluster: `frd_brk12`, `conf12`, `discrim`, `hea_lit`, `unemp12`, and `rel_end12` form a *looser* cluster focused on social and relational stressors (e.g., friendship breakup, conflicts, and discrimination). Compared to Euclidean clustering, `discrim` and `hea_lit` (health literacy) appear closer to relational stressors here, indicating that non-linear relationships might play a larger role in linking these variables.

#### Summary

In conclusion, mutual information-based clustering provides an alternative perspective that can reveal more intricate associations between variables, especially for those with non-linear relationships. Compared to Euclidean clustering, it shows a similar high-level structure but emphasizes nuanced connections between variables, particularly around social support, employment, and financial stress.

## Conclusions on Precariousness factors

Based on the consistent findings across multiple analyses, we decided to exclude the variables `discrim`, `hea_lit`, `umemp12`, and `rel_end12`, as they do not clearly belong to any specific precariousness factor nor exhibit strong associations with depression (see the correlation table above). Therefore, we propose retaining the following key precariousness factors:

-   Employment Precariousness: `emp_stat`, `work_sit`
-   Social Precariousness: `soc_freq`, `soc_adq`
-   Housing Precariousness: `nb_safe`, `nb_res`, `nb_rent`, `cul_rec`
-   Recent Relational Stressors: `frd_brk12`, `conf12`
-   Recent Financial Stressors: `fincri12`, `inc_diff`

We construct each precariousness factor by calculating the mean value of the combined variables. Below, we present the updated correlation table for the newly composed factors, along with the corresponding distributions of all variables to be used in the causal discovery analysis.

```{r, echo=FALSE, fig.align='center'}
clust_data <- scaled_data |>
  mutate(
    P.emp = rowMeans(cbind(emp_stat, work_sit)),
    P.soc = rowMeans(cbind(soc_freq, soc_adq)),
    P.hou = rowMeans(cbind(nb_safe, nb_res, nb_rent, cul_rec)),
    S.rel = rowMeans(cbind(frd_brk12, conf12)),
    S.fin = rowMeans(cbind(fincri12, inc_dif))
  ) |>
  dplyr::select(17:33) |>
  as.data.frame()

# skimr::skim(clust_data)

# visualize corr
corrplot(cor(clust_data), method="color", col=col(200),
         type="upper", order="hclust",
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", number.cex=0.50, tl.srt=45,tl.cex=0.7,
         diag=FALSE
         )

```

```{r echo=FALSE, fig.align='center', fig.height=4.5, fig.width=8}
# Reshape the data to long format
clust_data_long <- clust_data %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot distributions with density lines
ggplot(clust_data_long, aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)), fill = "lightblue", color = "white", bins = 15) +  # Use density for histogram
  geom_density(adjust=2, linewidth=0.3, color = "gray", alpha = 0.5) +  # Add density line in red
  facet_wrap(~ variable, scales = "free") +
  labs(title = "Distribution of All Variables with Density Overlay", x = "", y = "") +
  theme_minimal()
```

## Results from CCI algorithm {#sec-cci}

```{r}
#| layout-ncol: 2
#| label: fig-sum-cci
#| fig-cap: "Resulting graph of precarity factors and depression sum score using CCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "CCI MAAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/CCI_depsum.png")
knitr::include_graphics("img/depsum_mat_cci.pdf")

```

```{r}
#| label: fig-sym-cci
#| layout-nrow: 2
#| fig-cap: "Resulting graph of precarity factors and individual depression symptoms using CCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "CCI MAAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/symptom_graph_CCI.png")
knitr::include_graphics("img/symptom_mat_cci.pdf")

```

```{r}
#| layout-nrow: 2
#| label: fig-presum-cci
#| fig-cap: "Resulting graph of precarity sum score and individual depression symptoms using CCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "CCI MAAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/presum_graph_CCI.png")
knitr::include_graphics("img/presum_mat_cci.pdf")
```

\clearpage

## Results from PC algorithm

```{r}
#| layout-ncol: 2
#| label: fig-pc_sum
#| fig-cap: "Resulting graphs of precarity factors and depression sum score using PC."
#| fig-subcap:
#|   - "Using both GaussianCI and RCoT"
#|   - "Using only RCoT"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/sum_PC_all.png")
knitr::include_graphics("img/sum_PC_RCoTonly.png")
```

```{r}
#| label: fig-pc_sym
#| fig-cap: "Resulting graphs of precarity factors and individual depression symptoms using PC."
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/PC_symptom.png")
```

```{r}
#| label: fig-pc_presum
#| fig-cap: "Resulting graphs of precarity sum score and individual depression symptoms using PC."
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/PC_presum.png")
```

\clearpage

## Randomized Conditional Independence / Correlation Test (RCIT & RCoT) {#sec-rcot}

RCIT (Randomized Conditional Independence Test) and RCoT (Randomized conditional Correlation Test) are advanced methods for scalable conditional independence (CI) testing, offering computational efficiency while maintaining the accuracy of kernel-based approaches. These methods evaluate conditional independence between two variables $X$ and $Y$ given a third variable $Z$ while addressing computational challenges inherent in kernel-based CI tests.  In this section, we provide a high-level overview of RCIT and RCoT based on [@strobl2019approximate].

### Kernel-Based Conditional Independence Testing

Traditional kernel-based CI tests, such as the Kernel Conditional Independence Test (KCIT), compute dependencies using the Hilbert-Schmidt Independence Criterion (HSIC) in reproducing kernel Hilbert spaces (RKHS) [@zhang2012kernel]. KCIT uses the following hypothesis framework:
$$
H_0: X \perp\!\!\!\perp Y \,|\, Z, \quad H_1: X \not\!\perp\!\!\!\perp Y \,|\, Z.
$$
The core quantity in KCIT is the partial cross-covariance operator:
$$
\Sigma_{XY \cdot Z} = \Sigma_{XY} - \Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY},
$$
where $\Sigma_{XY}$ represents the cross-covariance operator between $X$ and $Y$, and $\Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY}$ removes the dependence mediated by $Z$.

The squared Hilbert-Schmidt (HS) norm of $\Sigma_{XY \cdot Z}$ serves as the test statistic:
$$
\|\Sigma_{XY \cdot Z}\|^2_{HS} = 0 \quad \text{if and only if} \quad X \perp\!\!\!\perp Y \,|\, Z.
$$

KCIT estimates residual dependencies using kernel ridge regression:
$$
f^*(z) = K_Z (K_Z + \lambda I)^{-1} f(x),
$$
where $K_Z$ is the kernel matrix for $Z$, $f(x)$ is the kernel feature map for $X$, and $\lambda$ is the ridge regularization parameter. The residual function for $X$ is:
$$
f_\text{res}(x) = f(x) - f^*(z) = R_Z f(x),
$$
with:
$$
R_Z = I - K_Z (K_Z + \lambda I)^{-1}.
$$
The kernel matrix for residualized $X$ is:
$$
K_{X \cdot Z} = R_Z K_X R_Z,
$$
and similarly for $Y$, $K_{Y \cdot Z} = R_Z K_Y R_Z$.

The test statistic is computed as:
$$
T_{XY \cdot Z} = \frac{1}{n^2} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
$$
which estimates the Hilbert-Schmidt (HS) norm of the partial cross-covariance operator. To ensure convergence, KCIT scales the statistic by $n$:
$$
S_K = n T_{XY \cdot Z}.
$$
The null hypothesis $H_0$ is rejected if $S_K$ exceeds a threshold determined by permutation or moment-matching-based null distribution [@lindsay2000moment].

### Random Fourier Features (RFFs)

Kernel-based methods like KCIT face scalability issues, as they involve operations on $n \times n$ kernel matrices, which scale quadratically with the sample size $n$. RCIT and RCoT overcome this bottleneck using *Random Fourier Features (RFFs)* to approximate kernel operations efficiently.

#### Bochner’s Theorem
Bochner’s theorem provides the foundation for RFFs, stating that any continuous shift-invariant kernel $k(x, y)$ can be expressed as:
$$
k(x, y) = \int_{\mathbb{R}^p} e^{i \omega^\top (x - y)} \, dP_\omega,
$$
where $P_\omega$ is the spectral distribution of the kernel. For the widely used RBF kernel:
$$
k(x, y) = \exp\left(-\frac{\|x - y\|^2}{2\sigma^2}\right),
$$
$P_\omega$ follows a Gaussian distribution: $\omega \sim \mathcal{N}(0, \sigma^2 I)$.

#### RFF Approximation
Using Monte Carlo sampling, the kernel function is approximated as:
$$
k(x, y) \approx \phi(x)^\top \phi(y),
$$
where $\phi(x)$ is the random Fourier feature mapping:
$$
\phi(x) = \sqrt{\frac{2}{D}} \cos(W^\top x + b),
$$
with $W \sim \mathcal{N}(0, \sigma^2 I)$ and $b \sim \text{Uniform}(0, 2\pi)$. Here, $D$ is the number of Fourier features, which balances computational efficiency and approximation accuracy.

### Differences Between RCIT and RCoT

RCIT and RCoT differ in their test statistics, computational efficiency, and practical performance, which makes them suited for different scenarios in causal discovery. RCIT evaluates the Hilbert-Schmidt norm of the full partial cross-covariance operator, providing a general test for conditional independence but at a higher computational cost. RCoT simplifies the process by using the Frobenius norm of a finite-dimensional residualized cross-covariance matrix, significantly reducing complexity and improving scalability.

These distinctions are particularly important for large-scale datasets, where RCoT's computational efficiency makes it a practical choice for high-dimensional causal discovery tasks.

#### RCIT: Randomized Conditional Independence Test

RCIT tests full conditional independence by examining the squared Hilbert-Schmidt (HS) norm of the partial cross-covariance operator $\Sigma_{XY \cdot Z}$:
$$
S_K = n T_{XY \cdot Z} = \frac{1}{n} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
$$
where $T_{XY \cdot Z}$ is an empirical estimate of $\|\Sigma_{XY \cdot Z}\|^2_{HS}$. The null and alternative hypotheses are:
$$
H_0: \|\Sigma_{XY \cdot Z}\|^2_{HS} = 0, \quad H_1: \|\Sigma_{XY \cdot Z}\|^2_{HS} > 0.
$$

RCIT is a general test for conditional independence but becomes computationally demanding as the size of $Z$ increases, due to the high-dimensional kernel operations required.

#### RCoT: Randomized Conditional Correlation Test

RCoT simplifies the testing process by using a finite-dimensional partial cross-covariance matrix, avoiding full HS norm calculations. Instead, it uses the Frobenius norm of the residualized cross-covariance matrix:
$$
S' = n \|C_{AB \cdot C}\|_F^2,
$$
where $C_{AB \cdot C}$ represents the residualized cross-covariance matrix. The hypotheses are:
$$
H_0: \|C_{AB \cdot C}\|_F^2 = 0, \quad H_1: \|C_{AB \cdot C}\|_F^2 > 0.
$$

RCoT is computationally efficient and well-suited for large conditioning sets ($|Z| \geq 4$). Its simplicity enables robust calibration of the null distribution and improved scalability for high-dimensional data.

# Acknowledgements {.unnumbered}

The HELIUS study is conducted by the Amsterdam University Medical Centers, location AMC and the Public Health Service of Amsterdam. Both organisations provided core support for HELIUS. The HELIUS study is also funded by the Dutch Heart Foundation, the Netherlands Organization for Health Research and Development (ZonMw), the European Union (FP-7), and the European Fund for the Integration of non-EU immigrants (EIF). We are most grateful to the participants of the HELIUS study and the management team, research nurses, interviewers, research assistants and other staff who have taken part in gathering the data of this study.

<!-- ### Advantages of RCIT and RCoT -->

<!-- - **Scalability**: Linear complexity ($O(nD)$) makes these methods efficient for large datasets. -->
<!-- - **Robustness**: Provide accurate approximations of kernel-based CI tests while significantly reducing computational costs. -->
<!-- - **Applicability**: Suitable for large-scale causal discovery applications, particularly when integrated with algorithms like PC and FCI. -->

<!-- By combining the flexibility of kernel methods with the computational efficiency of RFFs, RCIT and RCoT represent powerful tools for non-parametric CI testing, enabling their use in high-dimensional and large-scale datasets. -->

<!-- ## Summarized Stable Edges Proportion {#sec-propmatrix} -->

<!-- ```{r} -->
<!-- #| label: fig-sum-mat -->
<!-- #| fig-cap: "Proportions of edge endpoint types for graphs based on depression sum score." -->
<!-- #| fig-align: center -->
<!-- #| out-width: 100% -->
<!-- knitr::include_graphics("img/sumgraph_mat.pdf") -->
<!-- knitr::include_graphics("img/sumgraph_mat.pdf") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-sym-mat -->
<!-- #| fig-cap: "Proportions of edge endpoint types for graphs based on individual depression symptoms." -->
<!-- #| fig-align: center -->
<!-- #| out-width: 100% -->
<!-- knitr::include_graphics("img/symgraph_mat.pdf") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-presum-mat -->
<!-- #| fig-cap: "Proportions of edge endpoint types for graphs based on individual depression symptoms with precarity sum score." -->
<!-- #| fig-align: center -->
<!-- #| out-width: 70% -->
<!-- knitr::include_graphics("img/presumgraph_mat2.pdf") -->
<!-- ``` -->