### 2025-05-16 draft_v1/draft_v1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage[]{palatino}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{amssymb}
\usepackage{booktabs} % Add to your preamble for cleaner table lines
\usepackage{makecell} % Add to your preamble for multi-line cells
\usepackage{tikz}
\usepackage{amsmath}
\usetikzlibrary{arrows.meta, shapes.geometric}

%% Define TikZ arrow styles
\newcommand{\starstar}{% 
\begin{tikzpicture}[baseline=-3pt]
    \draw [{Rays[n=6]}-{Rays[n=6]}] (0,0) -- (0.55,0);
\end{tikzpicture}
}

\newcommand{\circstar}{% 
\begin{tikzpicture}
    \draw [{Circle[open]}-{Rays[n=6]}] (0,0) -- (0.55, 0);
\end{tikzpicture}
}

\newcommand{\starcirc}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-{Circle[open]}] (0,0) -- (0.55, 0);
\end{tikzpicture}
}

\newcommand{\stararrow}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-{Straight Barb[length=2.5pt]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\arrowstar}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-{Rays[n=6]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\circarrow}{% 
\begin{tikzpicture}
    \draw [{Circle[open]}-{Straight Barb[length=2.5pt]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailcirc}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [-{Circle[open]}] (0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\circtail}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [{Circle[open]}-] (0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\circirc}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [{Circle[open]}-{Circle[open]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailstar}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [-{Rays[n=6]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\startail}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailarrow}{% 
\begin{tikzpicture}
    \draw [-{Straight Barb[length=2.5pt]}](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\arrowtail}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\arrowarrow}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-{Straight Barb[length=2.5pt]}](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand\stackedarrows{%
        \mathrel{\vcenter{\mathsurround0pt
                \ialign{##\crcr
                \noalign{\nointerlineskip}$\arrowtail$\crcr
                \noalign{\nointerlineskip}$\tailarrow$\crcr
                }
        }}%
}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={From Causal Discovery to Intervention Simulation: Modeling Precarity and Depression in the HELIUS Study},
  pdfauthor={Kyuri Park; Leonie K. Elsenburg; Mary Nicolao; Karien Stronks; Vítor V. Vasconcelos},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{From Causal Discovery to Intervention Simulation: Modeling
Precarity and Depression in the HELIUS Study}

\author[1]{Kyuri Park}
\author[2]{Leonie K. Elsenburg}
\author[2]{Mary Nicolao}
\author[2]{Karien Stronks}
\author[1, 3]{Vítor V. Vasconcelos}

\affil[1]{\textit{Computational Science Lab, Informatics Institute, University of Amsterdam, PO Box 94323, Amsterdam, 1090GH, the Netherlands}}
\affil[2]{\textit{Department of Public and Occupational Health, Amsterdam Public Health Research Institute, Amsterdam UMC, University of Amsterdam, Amsterdam, the Netherland}}
\affil[3]{\textit{Institute for Advanced Study, University of Amsterdam, Oude Turfmarkt 147, Amsterdam, 1012GC, the Netherland}}

\date{2025-05-14}
\begin{document}
\maketitle
\begin{abstract}
\noindent This study investigates the dynamic relationship between
precarious life conditions and depression by combining constraint-based
causal discovery with computational modeling. Using cross-sectional data
from the HELIUS study, we applied cycle-capable causal discovery
algorithms to uncover directional relationships between financial
stress, domain-specific precariousness indicators, and depressive
symptoms. Our multi-resolution analysis---spanning both aggregate and
symptom-level constructs---identified financial stress as a robust
driver of depression, and highlighted key symptoms, such as sleep
disturbance and depressed mood, as likely initiators or reinforcers of
downstream precariousness. To explore how these structural insights
shape intervention effectiveness, we developed a simplified nonlinear
dynamical model simulating the co-evolution of depression and social
precarity under varying levels of financial stress reduction. Simulation
results revealed that the system's responsiveness depends not only on
the intensity of external intervention but also on internal structural
features such as feedback strength and noise. In particular, some
configurations exhibited bistability, requiring interventions to surpass
a critical threshold to trigger lasting improvement, while others
responded more proportionally. Together, these findings emphasize that
internal feedback dynamics---not just external adversity---can shape the
impact of mental health interventions. More broadly, our study offers a
scalable framework for integrating causal discovery and dynamical
simulation to better understand and inform policy responses to complex
psychosocial phenomena.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}

\section{Introduction}\label{introduction}

Mental health disorders represent a growing global health concern,
especially in urbanized regions (Gruebner et al., 2017; World Health
Organization, 2022). In 2019, one in eight people worldwide were living
with a mental health condition, with the highest disability-adjusted
life years (DALYs) due to mental and addictive disorders concentrated in
high-income countries such as those in Northern Europe, North America,
and Australia (Rehm \& Shield, 2019; World Health Organization, 2022).
Despite growing awareness, governments have struggled to design
effective responses. Mental health outcomes arise from complex,
multi-level dynamics---intertwined with the social, economic, and
spatial structures of urban environments---making both diagnosis and
intervention design deeply challenging (Van Der Wal et al., 2021).

Recent work has shifted focus from individual-level vulnerability to
broader social and structural determinants. While high-income nations
show elevated overall DALYs, within-country disparities reveal that
income inequality is a significant predictor of mental health burden
(Rehm \& Shield, 2019). Moreover, precarious working conditions, housing
instability, and neighborhood disadvantage have all been linked to
increased risk of depression and anxiety (Fone et al., 2014; Pevalin et
al., 2017; Rönnblad et al., 2019; Rugulies et al., 2023). This growing
literature highlights that poor mental health is not merely a personal
issue, but one rooted in sustained socioeconomic stress and structural
uncertainty.

Building on this, recent studies have introduced the concept of
precariousness as a multidimensional condition characterized by
instability and lack of control across several life domains---including
employment, finances, housing, social relations, and cultural belonging
(Elsenburg et al., 2025). This broader view reveals how different forms
of insecurity can co-occur and compound, forming an ecosystem of risk.
However, despite growing evidence of association between precariousness
and mental health, a fundamental question remains unresolved: \emph{How
do these factors causally influence one another?} Most existing studies
rely on cross-sectional correlations, leaving open questions about
directionality and feedback. Without stronger causal insight, it remains
difficult to identify priority targets for intervention or to anticipate
unintended effects.

This study aims to investigate the causal mechanisms linking
precariousness and depression---and how those mechanisms shape system
responses to external intervention. Specifically, we ask: \emph{How does
the internal configuration of the precarity--depression relationship
affect the outcomes of stress-reducing interventions?} To answer this,
we combine two approaches. First, we use cycle-capable causal discovery
algorithms to infer directional relationships among financial stress,
domain-specific precariousness factors, and depressive symptoms. By
analyzing both aggregate scores (e.g., PHQ-9 total, composite precarity
index) and disaggregated variables (individual symptoms and life-domain
indicators), we identify both broad associations and fine-grained causal
pathways. Second, we translate these insights into a computational
dynamical model that simulates how depression and precarity co-evolve
under varying levels of external stress. This model enables us to probe
how system properties---such as feedback strength and stochastic
noise---govern responsiveness to intervention. In some configurations,
the system exhibits bistability, remaining stuck in a high-risk state
unless external support exceeds a critical threshold. In others, it
shows unistable behavior, responding more gradually to incremental
changes. By integrating causal discovery with dynamic simulation, we aim
to demonstrate how the internal structure of the depression--precarity
system shapes its susceptibility to change---and, by extension, how
intervention strategies might be tailored to the architecture of
underlying vulnerability.

\section{Methods}\label{methods}

\subsection{Data}\label{data}

We use data from the HELIUS (HEalthy LIfe in an Urban Setting) study,
which has been described in detail elsewhere (Snijder et al., 2017;
Stronks et al., 2013). It captures the diverse population of the city of
Amsterdam by including the six main ethnic groups and which provides
comprehensive health and lifestyle data, including depressive symptoms
as measured by the PHQ-9 (Galenkamp et al., 2017). To operationalize
indicators of precariousness, we draw on the framework outlined in
previous research (Elsenburg et al., 2025) and select a set of relevant
variables. To ensure a robust representation of precariousness in our
causal discovery models, we conducted various exploratory analyses to
identify consistent and meaningful data-driven factor structures. Based
on these analyses, we identified five precariousness factors, including
two related to recent stressors, each comprising multiple variables as
outlined below. DFurther details on the HELIUS study and factor
extraction are provided in the \hyperref[sec-appendix]{Appendix}.

\begin{itemize}
\tightlist
\item
  Employment precariousness: \texttt{emp\_stat}, \texttt{work\_sit}.
\item
  Social precariousness: \texttt{soc\_freq}, \texttt{soc\_adq}.
\item
  Housing precariousness: \texttt{nb\_safe}, \texttt{nb\_res},
  \texttt{nb\_rent}, \texttt{cul\_rec}.
\item
  Recent relational stressors: \texttt{frd\_brk12}, \texttt{conf12}.
\item
  Recent financial stressors: \texttt{fincri12}, \texttt{inc\_diff}.
\end{itemize}

After preprocessing, the HELIUS dataset comprises 21,628 samples. Along
with the five precariousness factors, we also compute an overall
precarity score as the combined value of these factors. Additionally,
PHQ-9 scores are included to represent depression, both as a total sum
score and as individual symptom scores. In the subsequent causal
discovery analysis, we examine the relationship between depression and
precarity using both aggregated sum scores and their individual-level
representations. Refer to Figure~\ref{fig-dist} for the overall
distributions of the variables used in the analysis.

\begin{figure}

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{draft_v1_files/figure-pdf/fig-dist-1.pdf}

}

\caption{\label{fig-dist}Distributions of variables with density
overlay.\\
\emph{P.emp} = employment precariousness; \emph{P.hou} = housing
precariousness; \emph{P.soc} = social precariousness; \emph{S.fin} =
recent financial stressors; \emph{S.rel} = recent relational stressors;
\emph{Precarity} = overall precarity; \emph{PHQsum} = PHQ-9 sum score;
\emph{anh} = anhedonia; \emph{app} = appetite; \emph{con} =
concentration; \emph{dep} = depressed mood; \emph{ene} = energy;
\emph{glt} = guilty; \emph{mot} = motor; \emph{sui} = suicidal}

\end{figure}%

\subsection{Causal Discovery}\label{sec-analysis}

To uncover directional relationships between precariousness and
depression, we used constraint-based causal discovery algorithms that
allow for feedback loops and latent confounding. Specifically, we
applied the \emph{Fast Causal Inference} (FCI) and \emph{Cyclic Causal
Inference} (CCI) algorithms, both of which can detect cycles and
non-acyclic structures (Mooij \& Claassen, 2020; Strobl, 2019). For
reference, we also included the \emph{PC} algorithm, one of the most
widely used methods for acyclic causal discovery (Spirtes et al., 2001).

A key challenge in applying these algorithms to the HELIUS dataset is
that many variables are non-Gaussian and their relationships are likely
nonlinear. To address this, we supplemented the commonly used Gaussian
conditional independence (CI) test---based on partial
correlations---with a non-parametric alternative: the \emph{Randomized
Conditional Correlation Test} (RCoT). RCoT approximates kernel-based CI
testing using random Fourier features, significantly reducing
computational demands while maintaining sensitivity to nonlinear
dependencies (Strobl et al., 2019; Zhang et al., 2012).

We examined the causal structure using three complementary approaches:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Aggregate-level analysis: relationships between five domain-specific
  precariousness factors and the PHQ-9 sum score.
\item
  Fully disaggregated analysis: relationships between individual
  precariousness items and individual depressive symptoms.
\item
  Mixed analysis: relationships between individual symptoms and a
  composite index of overall precariousness.
\end{enumerate}

The aggregate analysis reduces dimensionality and yields interpretable
summaries of how domains of precarity relate to overall depression
severity. However, it may obscure finer-grained effects. The
disaggregated analysis allows for precise mapping of which symptoms are
influenced by (or influence) specific precariousness conditions, but it
introduces complexity due to the high dimensionality and distributional
properties of the data. The mixed approach, in turn, integrates these
perspectives by examining how individual symptoms respond to cumulative
precarity across domains.

To ensure robust estimation and guard against sensitivity to parameter
settings, we implemented a systematic grid of analysis conditions:

\begin{itemize}
\tightlist
\item
  Significance levels: \(\alpha\) = 0.01 and 0.05
\item
  Bootstrap thresholds: 0.5, 0.6, 0.7, 0.8
\item
  CI tests: Gaussian and RCoT
\item
  Algorithms: FCI, CCI, and PC
\end{itemize}

Each combination was bootstrapped 100 times for sum-score level analyses
and 30 times for symptom-level analyses, yielding a total of 1,600
graphs per algorithm for the full analysis. For symptom networks, we
fixed the skeleton structure using a consensus graph estimated across
PC, FCI, and CCI (\(\alpha\) = 0.01, RCoT). This constraint improves
computational efficiency without limiting the estimation of directional
edges between symptoms and precariousness variables.

To summarize the results, we adopted a two-level aggregation strategy.
First, within each analysis condition, we recorded the most frequently
observed edge-endpoint type (e.g., directed, undirected, uncertain) from
the bootstraps. Then, across all conditions, we tallied the most
frequent endpoint type per edge. In cases of ties, we assigned dashed
edges to reflect uncertainty. For transparency, we accompany each graph
with a table showing the relative frequencies of each edge symbol.

This procedure yields a final graph for each algorithm and analysis
type---offering both robustness and interpretability. See
Figure~\ref{fig-workflow} for an overview of the causal discovery
pipeline.

For detailed explanations of the algorithms, edge interpretations, and
graph types (e.g., PAG, MAAG, CPDAG), as well as algorithmic assumptions
and limitations, see the \hyperref[sec-appendix]{Appendix}.

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{img/simsetup2.png}

}

\caption{\label{fig-workflow}Analysis workflow applied across all three
algorithms.}

\end{figure}%

\subsection{Modeling Dynamics between Precarity and
Depression}\label{modeling-dynamics-between-precarity-and-depression}

To explore how internal system structure shapes responses to external
support, we implement a high-level computational model that formalizes
key mechanisms suggested by the causal discovery analysis. Rather than
attempting to replicate the full estimated causal graph, this model
abstracts the essential structure into a compact, tractable system
focused on dynamic interactions between depression and social
precariousness, both influenced by financial stress.

As illustrated in Figure~\ref{fig-concept}, the model includes three
components: depression (\(D\)), represented by the PHQ-9 sum score,
social precariousness (\(P\)), and external financial stress (\(S\)).
Depression and precariousness are modeled as coupled state variables
that influence each other over time, with \(S\) acting as a shared
input. These dynamics are formalized as a pair of nonlinear stochastic
differential equations (SDEs):

\[
\begin{aligned}
dP &= \lambda_P \left( \tanh(\alpha_1 S + \alpha_2 D) - P \right) dt + \sigma_1 dW_1 \\
dD &= \lambda_D \left( \tanh(\beta_1 S + \beta_2 P) - D \right) dt + \sigma_2 dW_2 
\end{aligned}
\]

Here, \(D\) denotes depression, \(P\) denotes social precariousness, and
\(S\) is the external stressor associated with financial stress. The
parameters \(\lambda_D\) and \(\lambda_P\) control the timescales over
which variable adjusts toward its input-driven target state. The
coefficients \(\alpha_1, \alpha_2\) and \(\beta_1, \beta_2\) represent
the strengths of the causal influences between variables. Stochastic
fluctuations are introduced through the terms \(\sigma_1 dW_1\) and
\(\sigma_2 dW_2\), where \(dW_1\) and \(dW_2\) are independent Wiener
processes. The system uses the hyperbolic tangent function (\(tanh\)) to
bound the influence of inputs, ensuring smooth, saturating dynamics and
preventing runaway feedback loops. Importantly, the nonlinearity
introduced by the \(tanh\) function also allows the system to exhibit
bistability, supporting the existence of two stable states: one
corresponding to low depression and precariousness, and the other to
high depression and precariousness.

\begin{figure}

\centering{

\includegraphics[width=0.35\textwidth,height=\textheight]{img/conceptmodel.png}

}

\caption{\label{fig-concept}Conceptual high-level causal structure of
precariousness and depression system.}

\end{figure}%

\subsubsection{Model Calibration}\label{model-calibration}

To identify realistic parameter settings for the simulation model, we
calibrated six free parameters---\(\alpha_1\), \(\alpha_2\),
\(\beta_1\), \(\beta_2\), and noise amplitudes \(\sigma_1\),
\(\sigma_2\)---using the Non-dominated Sorting Genetic Algorithm II
(NSGA-II) (Deb et al., 2002), a multi-objective evolutionary
optimization method well-suited for noisy and high-dimensional search
spaces. The goal was to match simulated system behavior to empirical
summary statistics derived from the HELIUS dataset, including the means
and variances of depression (\(D\)) and social precariousness (\(P\)),
as well as their partial correlations after adjusting for financial
stress (\(S\)), specifically \(\rho_{DP.S}\), \(\rho_{DS.P}\), and
\(\rho_{PS.D}\). Given the stochastic nature of the model, each
candidate parameter set was evaluated by simulating 1,000 independent
trajectories, each representing an individual with a fixed \(S\) value
sampled from the observed distribution. Final-state values were
extracted and aggregated to compute population-level summary statistics.
Fitness was assessed as the absolute deviation from the empirical
targets, aggregated across all objectives. NSGA-II was implemented via
the nsga2R package (Franz \& Nakamura, 2015), with a population of 300
evolved over 300 generations using standard hyperparameters (crossover
probability = 0.9, mutation probability = 0.25, and tournament size =
2). This process yielded a diverse ensemble of Pareto-efficient
solutions balancing trade-offs between objectives. From the resulting
Pareto front (Figure~\ref{fig-pareto}), we selected the top 30 parameter
sets with the lowest combined error to use in the subsequent simulations
assessing intervention effects.

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{img/paretoplot.png}

}

\caption{\label{fig-pareto}Pareto front of simulated parameter sets.
\emph{Each point represents a solution evaluated by its total variance
error (x-axis) and total partial correlation error (y-axis), with color
indicating total mean error. Red crosses highlight the top 30 solutions
used for simulation.}}

\end{figure}%

\subsubsection{Intervention Simulation}\label{intervention-simulation}

To evaluate the effects of hypothetical support, we simulate scenarios
in which each individual's financial stress level (\(S\)) is
proportionally reduced toward the minimum observed value. This models a
population-wide intervention applied at varying strengths --- from no
reduction in stress (0\%) to complete alignment with the minimum
observed stress level (100\%).

For each of the top 30 calibrated parameter sets, we simulate the
system's evolution across 300 individuals under a continuum of
intervention levels. Each simulation run generates final values for
\(D\) and \(P\), which are then used to assess how internal model
parameters influence the system's behavior under changing external
conditions.

\section{Results}\label{results}

\subsection{Causal Structure Linking Precarity and
Depression}\label{causal-structure-linking-precarity-and-depression}

\subsubsection{Depression as sum score}\label{depression-as-sum-score}

The sum score graphs provide a high-level summary of how precarity
factors collectively influence overall depression severity, focusing on
aggregated relationships. Figure~\ref{fig-sum} illustrates the causal
relationships between precariousness factors (\emph{P.hou},
\emph{P.emp}, \emph{P.soc}, \emph{S.rel}, \emph{S.fin}) and the
depression sum score (\emph{PHQsum}), as identified by the FCI
algorithm. As described in Section~\ref{sec-analysis}, dashed edges
indicate ties, with the exact edge endpoint where the tie occurs is
represented by a bold horizontal bar.

The causal graph in panel (a) highlights key pathways in the
relationships between precarity factors and depression (PHQsum).
Employment precarity (P.emp) and social precarity (P.soc) are not
identified as causes of depression, whereas financial stress (S.fin)
appears to play a causal role. While P.emp and P.soc are generally not
recognized as causes of other precariousness factors, S.fin emerges as a
potential cause, as indicated by its circle edge endpoint. Supporting
this, the proportion matrix plot in panel (b) shows that S.fin has some
probability of causally influencing either P.emp or P.soc.

The matrix plot can be interpreted such that the symbol in
matrix{[}\(i\), \(j\){]} represents the relationship
\(i -~[symbol] \  j\). For example, if matrix{[}\(i\), \(j\){]} is
\textgreater{} and matrix{[}j, i{]} is \(\circ\), then the inferred
relationship is \(i \circarrow j\). In the table, different edge types
are represented by distinct colors: light gray for the absence of an
edge, blue for circles, green for arrowheads, and coral for arrowtails.
The colors are blended based on the proportion of each edge type, with
higher proportions increasing the opacity of the corresponding color,
making dominant symbols more visually prominent.

Relational stress (S.rel) plays a more nuanced role, interacting with
depression and S.fin through a latent confounder. Its relationship with
P.soc remains less certain, as the proportion table suggests a fair
probability that this connection is absent, uncertain, or bidirectional.
Meanwhile, housing precarity (P.hou) is not directly causally related to
depression but is linked to P.emp and S.fin. While P.emp and S.fin are
identified as non-causes of P.hou, it remains unclear whether P.hou
causally influences P.emp or if their relationship is mediated by an
unobserved confounder.

The findings from the CCI algorithm (see Figure~\ref{fig-sum-cci}) are
largely consistent with those from the FCI algorithm, with one
exception: CCI does not identify S.fin as a cause of depression.
Additionally, CCI introduces greater uncertainty in edge directions,
with more ties mainly between circle and arrowhead endpoints, and tends
to favor arrowheads more frequently than FCI. Despite these differences,
the skeleton structure and overall causal directions derived from CCI
align well with the results of the FCI algorithm, supporting the key
pathways.

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/FCI_depsum.png}

}

\subcaption{\label{fig-sum-1}FCI PAG}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/depsum_mat_fci.pdf}

}

\subcaption{\label{fig-sum-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sum}Resulting graph of precarity factors and
depression sum score using FCI and proportion of edge endpoint types.}

\end{figure}%

\subsubsection{Individual depression
symptom}\label{individual-depression-symptom}

Moving from the sum score representation to the disaggregated
symptom-level graph provides a more granular perspective on the causal
relationships between precariousness factors and depression symptoms.
Unlike the sum score graph, which aggregates all symptoms into a single
measure---potentially obscuring nuanced relationships---the
symptom-level graph highlights heterogeneity in how precariousness
factors influence individual depressive symptoms: con (concentration),
slp (sleep), ene (energy), app (appetite), mot (motor), sui (suicidal),
anh (anhedonia), glt (guilt), and dep (depressed mood), and vice versa.

The symptom-level graph in Figure~\ref{fig-sym} panel (a) reveals a more
complex and interconnected structure than the sum score graph,
reflecting the strong interdependence among symptoms and suggesting much
presence of latent confounders, as indicated by numerous bidirectional
edges. As before, dashed edges denote ties, with specific ties marked by
bold horizontal bars. In Figure~\ref{fig-sym} panel (b), the proportion
matrix provides more insight into these directionalities. Here, opaque
green dominates most symptom-to-symptom connections, indicating that
arrowheads are the most frequently inferred edge type. However, blue
regions, which correspond to circle endpoints, are particularly common
for anh, slp, ene, and sui, reflecting uncertainty in the causal
direction among these symptoms. Notably, the only arrowtail connection
appears between dep and anh, suggesting that anhedonia (anh) is a
potential cause of depressed mood (dep).

Looking at the symptom-precariousness connections, one of the key
patterns is that ties are most frequently found in the relationships
between individual symptoms and precariousness factors. This suggests
that these causal links may be less stable across different conditions.
Upon closer examination, most ties arise due to discrepancies between
the Gaussian CI test and the RCoT test---with Gaussian CI favoring
arrowheads and RCoT more often predicting the absence of an edge.

This discrepancy likely arises from a combination of factors. First, as
the analysis shifts from aggregated to disaggregated variables, the
number of conditional independence (CI) tests increases, leading to a
loss of statistical power in high-dimensional causal discovery. With
more variables being conditioned on, the data becomes sparser, making it
harder to detect weaker dependencies. Additionally,
over-conditioning---controlling for too many variables---can
artificially remove statistical associations, leading to false
negatives. Beyond this general statistical power issue, the fundamental
differences in how these CI tests operate contribute to the
inconsistency. RCoT, as a nonparametric test, does not assume linearity
and can capture both linear and nonlinear dependencies. However, it is
more conservative and requires larger sample sizes, making it more prone
to false negatives in high-dimensional settings. In contrast, partial
correlation (Gaussian CI test) assumes linear Gaussian relationships,
making it more permissive and sometimes detecting weak dependencies that
may not be statistically meaningful. As a result, when Gaussian CI
detects an edge while RCoT does not, it could indicate that the
relationship is weak, strictly linear, or requires more data for
reliable nonparametric detection. This explains why many ties in the
proportion matrix show a 50-50 split between symbols---where half of the
conditions with RCoT suggest the absence of an edge, while the other
half using Gaussian CI test favor an arrowhead. This systematic
divergence highlights the challenges of causal discovery in
high-dimensional settings, where varying statistical assumptions can
lead to differing conclusions about the presence and direction of causal
relationships.

Despite these differences, some consistent patterns emerge across both
CI tests, particularly in the case of financial stress (S.fin), which
appears to be connected to nearly all other variables---though many of
these connections are marked by ties, reflecting uncertainty in
directionality. One exception is the stronger evidence suggesting that
S.fin may cause changes in app (appetite), while its relationships with
other symptoms, such as ene, mot, and dep, remain ambiguous. Similarly,
social precarity (P.soc) exhibits numerous connections with depressive
symptoms, with most edges pointing toward P.soc rather than outward from
it. This suggests that depressive symptoms, particularly dep and glt,
may contribute to worsening social precarity rather than the other way
around. Additionally, anh also shows some probability of causally
influencing P.soc, reinforcing the idea that social precarity is more
often a consequence rather than a driver of depressive symptoms.

Other precariousness factors exhibit more uncertain but still notable
relationships. Relational stress (S.rel) shows weak but existing
connections with slp, glt, and app, though directionality remains
unclear in many cases. Employment precarity (P.emp) also connects to
symptoms like slp, mot, and con, but, like S.rel, these relationships
exhibit a 50/50 split in directionality, reflecting uncertainty in the
inferred causal paths. In line with the aggregated graph in
Figure~\ref{fig-sum}, housing precarity (P.hou) does not appear to have
any direct relationship with depression symptoms but consistently shows
associations with employment precarity (P.emp) and financial stress
(S.fin). The causal relationships among precariousness factors remain
largely unchanged from the aggregated analysis, with financial stress
(S.fin) exhibiting the strongest tendency to influence other
precariousness factors.

Among depressive symptoms, dep, glt, and slp appear to be the most
connected to precariousness factors, while P.soc has the most
connections with symptoms, predominantly as a recipient rather than a
driver of influence. Within the symptom network, slp, sui, and anh
emerge as causally influential symptoms, as they exhibit more outgoing
arrows compared to other symptoms. On the other hand, con, mot, dep, and
glt, despite having high connectivity, predominantly receive incoming
arrows, indicating they are more likely effects rather than causes.
Considering both symptom-precariousness connections and symptom-level
dynamics, sleep disturbance (slp) emerges as a central symptom, given
its strong ties to precariousness factors and its influential role
within the symptom network. This suggests that sleep issues could be an
initiating symptom, particularly sensitive to relational stress and
employment precarity. On the other hand, social precarity (P.soc) serves
as a key bridge between the depression and precariousness subsystems, as
depressive symptoms appear to feed back into social precarity,
reinforcing a self-sustaining dynamic between depression and precarious
conditions.

Finally, the results from CCI are largely consistent with those from
FCI, but with some key differences. CCI tends to favor arrowheads more
frequently, resulting in a greater number of bidirectional edges, which
suggests a higher involvement of latent confounders. Additionally, CCI
exhibits more tie situations, though, similar to FCI, most ties occur
between absence and arrowhead edges, reflecting the discrepancy between
the Gaussian CI test and RCoT---where the Gaussian test favors
arrowheads, while RCoT more often suggests the absence of an edge. For a
detailed visualization of the CCI-derived graph and proportion matrix,
see Figure~\ref{fig-sym-cci}.

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{img/symptom_graph_FCI.png}

}

\subcaption{\label{fig-sym-1}FCI PAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{img/symptom_mat_fci.pdf}

}

\subcaption{\label{fig-sym-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sym}Resulting graph of precarity factors and
individual depression symptoms using FCI and proportion of edge endpoint
types.}

\end{figure}%

\subsubsection{Precarity as sum score}\label{precarity-as-sum-score}

Lastly, we examine the relationships between individual symptoms and
overall precarity, an aggregated measure represented by the sum of five
precariousness factors. This analysis provides a complementary
perspective, capturing broader patterns that may not be evident in the
disaggregated symptom-precariousness analysis. By aggregating
precariousness factors into a single score, this approach may capture
distributed relationships across different precarity factors that were
previously overlooked in the disaggregated analysis.

Unlike the disaggregated graph in Figure~\ref{fig-sym}, the aggregated
symptom-precariousness graph (Figure~\ref{fig-presum} panel (a)) shows
more certainty in causal directions, with only a few edges resulting in
ties, all of which involve the overall precarity factor. This is even
more evident in the proportion matrix (Figure~\ref{fig-presum} panel
(b)), where symptom-to-symptom interactions almost fully converge to a
single edge type. Additionally, all symptom interactions are connected
either through bidirectional edges or a combination of an arrowhead and
a circle, suggesting a strong presence of latent confounders.

Examining the symptom-precarity connections, we observe a stronger
overall trend of depressive symptoms influencing precarity, rather than
the other way around. Specifically, anh, dep, and slp show a high
probability of causing precarity, whereas app, glt, and mot exhibit more
uncertainty in directionality, often resulting in circle endpoints. This
pattern suggests that when precariousness factors are aggregated, the
dominant causal flow is from depressive symptoms to precarity, rather
than precarity driving depression. Particularly, dep emerges as the
strongest predictor of precarity, exhibiting the highest proportion of
arrowtails, consistent with findings from the disaggregated analysis.

Comparing this with the CCI-derived graph (Figure~\ref{fig-presum-cci}),
several key differences stand out. The CCI results reveal that most
symptoms are interconnected through bidirectional edges, suggesting a
strong presence of latent confounders. However, dep appears to be a
distinct exception, as it is predominantly caused by nearly all other
symptoms, with one exception sui, which does not contribute to dep.
Similarly, con is influenced by multiple symptoms, albeit with weaker
support compared to dep. Most interestingly, the CCI results highlight a
distinct feedback loop between dep and overall precarity, represented by
a tail-tail edge (\textemdash ). This suggests a possible reinforcing
cycle between depression and overall precariousness, where depression
not only arises from precarious conditions but also contributes to their
persistence, creating a self-sustaining dynamic.

Overall, the aggregated precariousness graph supports the key findings
from both the disaggregated analysis and the depression sum score
analysis, providing greater certainty in several important patterns. The
results highlight the significant role of latent confounders in
symptom-to-symptom interactions. While certain stressor-related
precariousness factors, particularly S.fin and S.rel, appear to causally
influence depressive symptoms, other precariousness factors are more
likely driven by depression rather than acting as primary causes,
suggesting that precariousness may function more as a consequence of
depression than its source. Some symptoms, particularly glt, slp, and
dep, show greater sensitivity to precarious conditions, while also
playing a central role in symptom dynamics. Among them, slp and glt
emerge as influential symptoms within the depression network, whereas
con and dep, despite their high connectivity, behave more as outcomes
rather than primary drivers. Together, these findings suggest that
depression---especially symptoms like sleep disturbances and guilt---may
contribute to sustaining precarious conditions, possibly forming a
reciprocal cycle between depression and precariousness.

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_graph_FCI.png}

}

\subcaption{\label{fig-presum-1}FCI PAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_mat_fci.pdf}

}

\subcaption{\label{fig-presum-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-presum}Resulting graph of precarity sum score and
individual depression symptoms using FCI and proportion of edge endpoint
types.}

\end{figure}%

\subsection{Simulating the Dynamics Between Precarity and Depression
Under
Intervention}\label{simulating-the-dynamics-between-precarity-and-depression-under-intervention}

To evaluate how internal system structure influences the effect of
external support, we simulated outcomes under a range of intervention
strengths across the top 30 parameter sets from the calibrated model
ensemble. The top panel of Figure~\ref{fig-intervention} shows mean
levels of depression (\(D\)) and precarity (\(P\)) after simulation,
with each line representing one parameter set. Lines are color-coded by
combined feedback and input strength, calculated as the product
\(\alpha_1 \times \alpha_2 \times \beta_1 \times \beta_2\).

Parameter sets with stronger interactions exhibit more pronounced
declines in both depression and precarity in response to intervention.
At full intervention---modeled as a complete shift of external stress
(\(S\)) toward its observed minimum---the mean depression level
decreases from -0.113 to -0.963, and precarity from -0.100 to -0.958.
These changes correspond to reductions of approximately 90.8\% relative
to their initial magnitudes, indicating that targeted intervention can
shift the system toward a substantially lower-risk equilibrium.

To better understand these dynamics, we focused on two illustrative
parameter sets: one with strong feedback/input coupling and low noise (a
``bistable'' candidate), and one with weaker coupling and higher noise
(a ``unistable'' candidate). In nonlinear systems, bistability refers to
the existence of two distinct stable equilibria---meaning the system's
long-term outcome depends on its starting point or perturbations.

In our model, bistability arises from the interaction between strong
internal coupling---particularly the product
\(\alpha_2 \times \beta_2\)---and the system's nonlinearity. The
\(\tanh(\cdot)\) activation function saturates at high input values,
meaning sufficient gain is required to push the system into nonlinear
regimes where multiple stable points emerge. If the feedback gain is too
low, or the system is overwhelmed by noise, it behaves more linearly and
settles into a single equilibrium---yielding a unistable
configuration.\footnote{This resembles a classic \emph{pitchfork
  bifurcation} in discrete dynamical systems: when the gain \(k\) in a
  scalar map like \(x_{t+1} = \tanh(k x_t)\) exceeds a critical value,
  the system shifts from a single fixed point to two stable attractors
  and one unstable one. In our model, high values of
  \(\alpha_2 \times \beta_2\) similarly enable the emergence of multiple
  attractor states.}

The bottom panel of Figure~\ref{fig-intervention} illustrates these
contrasting behaviors. In the bistable model, a clear bifurcation
emerges: small interventions have little effect, but once a threshold is
crossed, the system abruptly transitions to a low-risk state. In the
unistable model, the transition is gradual and continuous, with outcomes
improving incrementally as stress is reduced.

These findings underscore that the impact of intervention depends not
only on its intensity, but on the system's internal configuration. In
bistable regimes, small or moderate interventions may be insufficient
unless they push the system past a tipping point. Unistable systems, by
contrast, respond more linearly, offering greater predictability and
flexibility in designing incremental interventions.

This distinction is especially relevant for mental health interventions
in populations exposed to persistent stressors like housing instability,
financial strain, or job insecurity. In such contexts, external
adversity may be reinforced through internal feedback loops between
psychological distress and precarious conditions. Recognizing when a
system is bistable may help identify when strong, early, or multifaceted
interventions are needed to produce lasting change---while unistable
dynamics may justify more gradual or targeted efforts.

Tailoring intervention strategies to the system's underlying structure
may enhance their effectiveness and sustainability, particularly in
addressing the compounding effects of socioeconomic precarity on mental
health.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/intervention_plot2.pdf}

}

\caption{\label{fig-intervention}\textbf{Top}: Final Mean D and P Across
Intervention Sizes. \emph{Mean trajectories of depression (\(D\)) and
precarity (\(P\)) across varying intervention strengths, simulated over
the top 30 parameter sets. Each line represents one parameter set and is
colored by its combined feedback and input strength.} \textbf{Bottom}:
Smoothed Density Heatmap of Final \(D\) and \(P\) for Two Parameter
Sets. \emph{One exhibits unistability (top two panels) and the other
exhibits bistability (bottom two panels). Bistable dynamics reveal a
sharp transition between distinct outcome regimes, while unistable
dynamics show gradual, continuous change across the intervention
gradient.}}

\end{figure}%

\section{Discussion}\label{discussion}

This study combined constraint-based causal discovery and dynamical
modeling to investigate how depression and socioeconomic precarity
interact---and how their internal structure modulates responsiveness to
intervention. Using cross-sectional data from the HELIUS study, we
applied cycle-capable discovery algorithms to estimate plausible
directional relationships among financial stress, multiple domains of
precariousness, and depressive symptoms. By analyzing both aggregate
constructs and individual variables, we aimed to capture patterns that
may be masked in either extreme of abstraction (i.e., purely aggregate
or disaggregated analyses).

The causal analysis consistently identified financial stress
(\emph{S.fin}) as a key plausible driver of depression, appearing across
both aggregated and disaggregated models. At the symptom level, we found
that sleep disturbance (\emph{slp}), depressed mood (\emph{dep}), guilt
(\emph{glt}), and anhedonia (\emph{anh}) were especially sensitive to
external stressors. Among these, \emph{dep} emerged as a central hub:
not only was it heavily influenced by other symptoms, it also exerted
directional influence on social precarity (\emph{P.soc}), suggesting a
feedback loop in which worsening mood contributes to social
disconnection, which in turn may exacerbate depressive symptoms. The CCI
algorithm even suggested a possible cyclic relationship between dep and
overall precarity, hinting at a structurally self-reinforcing mechanism.
\emph{Slp} also played a key initiating role. It showed directed
connections to both social and employment precarity and exhibited many
outgoing edges within the symptom network. This suggests \emph{slp} may
function as an early warning signal or entry point into the
depression--precarity cycle, making it a promising target for preventive
intervention. In contrast, \emph{P.soc} more often appeared as an effect
rather than a cause---receiving arrows from symptoms like \emph{dep} and
\emph{glt}, but rarely sending them. This asymmetry implies that social
precarity may often reflect downstream consequences of worsening mental
health rather than being its root cause. These patterns offer actionable
insights for intervention. Addressing symptoms like \emph{slp} early may
prevent escalation into broader depressive states or worsening
precarity. Intervening on \emph{dep}, as a central node with potential
feedback effects, could also disrupt the self-reinforcing dynamics
between mental health and social isolation.

To examine how such structures shape responsiveness to support, we
developed a simplified two-variable nonlinear dynamical model capturing
the feedback between depression and precarity, modulated by financial
stress. Calibrated against empirical summary statistics from the HELIUS
dataset, the model allowed us to simulate how these dynamics evolve
under varying levels of stress reduction. The simulations revealed that
intervention effectiveness depends not just on the magnitude of external
stressors, but on the internal configuration of the system itself. In
bistable regimes---characterized by strong feedback and low noise---the
system could remain stuck in a high-risk state unless an intervention
crosses a critical threshold, tipping it into a healthier equilibrium.
By contrast, unistable systems---driven by weaker coupling or greater
noise---responded more gradually and proportionally to increasing
support. These dynamics carry practical policy implications. In bistable
systems, even moderate interventions may have little effect unless they
are forceful enough to tip the system toward a lower-risk state. This
underscores the importance of early or intensive support when targeting
self-reinforcing cycles of depression and precarity. Unistable systems,
by contrast, are more responsive to incremental support, suggesting that
sustained, moderate interventions may be sufficient in such contexts.

That said, our study has several limitations. While the FCI and CCI
algorithms are designed to recover causal structure from cross-sectional
data---including feedback and latent confounding---many edge directions
in our results remain ambiguous, as indicated by the prevalence of
circle endpoints and bidirectional arrows, especially in weakly
connected regions of the graph. These uncertainties were most evident in
the links between individual symptoms and precariousness factors, where
edge estimates varied across algorithms and conditional independence
(CI) tests. This likely reflects limited statistical signal in
high-dimensional settings, where large conditioning sets reduce power
and small effect sizes are harder to detect. Although the CCI algorithm
is explicitly designed to detect cycles, we found little clear evidence
of feedback: most candidate loops appeared only as bidirectional edges,
which may reflect latent confounding or statistical imprecision rather
than true cyclic structure. This could point to either an absence of
strong feedback at the symptom level or limitations in the sensitivity
of full-graph discovery approaches applied to densely interconnected
variables. More focused strategies, such as local structure tests or
theory-guided subgraph modeling, may help resolve these ambiguities. In
addition, while our analyses relied solely on observational data, some
structural uncertainties---particularly those involving latent
confounding or bidirectional influence---may ultimately require
interventional or semi-interventional designs. Extensions such as LLC
(Hyttinen et al., 2012), NODAGS-Flow (Sethuraman et al., 2023), or
Bicycle (Rohbeck et al., 2024), which integrate observational and
interventional data, may prove valuable where such data are available.

Second, our findings showed sensitivity to the choice of conditional
independence (CI) test. The Gaussian CI test, which assumes linear
relationships and Gaussian noise, often produced denser graphs---likely
due to its tendency to overestimate weak or noisy linear associations.
In contrast, RCoT, a nonparametric alternative, avoids these assumptions
but relies on a Gaussian radial basis function (RBF) kernel to assess
dependence. While flexible in capturing nonlinear relationships, this
kernel is optimized for smooth, continuous variables and may
underperform with discrete or ordinal data. Its similarity estimates can
be unreliable when applied to variables with discontinuities,
potentially leading to missed dependencies---particularly relevant in
the HELIUS dataset, where some variables are ordinal or mixed-type
(Howlett, 2001). To address this, future work should consider adapting
kernel-based CI tests to better handle mixed data structures---for
example, by employing hybrid kernels that explicitly combine continuous
and discrete similarity measures. This could improve the reliability of
nonparametric causal discovery in settings with heterogeneous variable
types.

Finally, the structure of the dynamical model introduces its own
limitations. While the model was designed to reflect key features of the
discovered causal structure, it represents a highly simplified system:
only two variables---depression and social precarity---modulated by
external financial stress. This abstraction allowed for tractable
analysis and visual intuition, particularly around nonlinear behaviors
such as bistability, but omits other potentially influential constructs
highlighted in the causal graphs, including relational stress,
employment conditions and housing conditions. Moreover, the model's
nonlinearity---introduced via the hyperbolic tangent function---was
chosen for convenience rather than empirical justification. While this
function supports saturation and the emergence of tipping points, the
presence of bistability is a product of simulation, not empirical
verification. It remains possible that simpler, linear dynamics could
account for the observed data just as well. Future work should further
investigate whether bistability is a necessary feature of
depression--precarity dynamics, and explore alternative functional
forms---both linear and nonlinear---using richer data and stronger
empirical constraints.

Our results reveal that the effectiveness of interventions targeting
depression and precarity is shaped not only by the strength of external
stressors but also by the internal structure of the system. The
interplay between feedback strength and noise determines whether a
system responds gradually or only after crossing a tipping
point---highlighting the need for intervention strategies attuned to
these underlying dynamics.

These findings underscore the importance of considering internal
feedback---not just external adversity---when designing interventions.
Mental health systems are not always passively responsive to support;
their structure can constrain how much impact any given policy has. By
diagnosing these structural features---through tools like causal
discovery and computational modeling---researchers and policymakers can
better tailor interventions to the underlying mechanics of
vulnerability.

Beyond these specific findings, this study offers a generalizable
framework for studying complex mental health phenomena. Causal discovery
enables principled hypothesis generation from observational data, while
dynamical modeling helps test how those structures behave under
intervention. Together, these methods move us from descriptive patterns
to mechanistic insight, with practical relevance for real-world mental
health policy.

Future work could build richer dynamical models grounded in
higher-resolution empirical data---potentially incorporating time,
interventions, or more granular psychosocial constructs. Such extensions
would help validate whether mechanisms proposed here, such as
bistability, manifest in real-world trajectories. Ultimately, we hope
this study encourages further integration of causal structure learning
and system-level modeling to improve mental health outcomes under
conditions of socioeconomic precarity.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-deb2002fast}
Deb, K., Pratap, A., Agarwal, S., \& Meyarivan, T. (2002). A fast and
elitist multiobjective genetic algorithm: NSGA-II. \emph{IEEE
Transactions on Evolutionary Computation}, \emph{6}(2), 182--197.

\bibitem[\citeproctext]{ref-dojer2016learning}
Dojer, N. (2016). Learning bayesian networks from datasets joining
continuous and discrete variables. \emph{International Journal of
Approximate Reasoning}, \emph{78}, 116--124.

\bibitem[\citeproctext]{ref-elsenburg2025clustering}
Elsenburg, L. K., Nicolaou, M., Galenkamp, H., Lakerveld, J., \&
Stronks, K. (2025). The clustering of disadvantage in different life
dimensions across ethnic groups: A network analysis of indicators of
precariousness in the HELIUS study. \emph{Social Science \& Medicine},
117970.

\bibitem[\citeproctext]{ref-entner2010causal}
Entner, D., \& Hoyer, P. O. (2010). On causal discovery from time series
data using FCI. \emph{Probabilistic Graphical Models}, \emph{16}.

\bibitem[\citeproctext]{ref-fone2014effect}
Fone, D., White, J., Farewell, D., Kelly, M., John, G., Lloyd, K.,
Williams, G., \& Dunstan, F. (2014). Effect of neighbourhood deprivation
and social cohesion on mental health inequality: A multilevel
population-based longitudinal study. \emph{Psychological Medicine},
\emph{44}(11), 2449--2460.

\bibitem[\citeproctext]{ref-forre2018constraint}
Forré, P., \& Mooij, J. M. (2018). Constraint-based causal discovery for
non-linear structural causal models with cycles and latent confounders.
\emph{arXiv Preprint arXiv:1807.03024}.

\bibitem[\citeproctext]{ref-nsga2R}
Franz, R., \& Nakamura, R. (2015). \emph{nsga2R: Elitist non-dominated
sorting genetic algorithm for multi-objective optimization}.
\url{https://CRAN.R-project.org/package=nsga2R}

\bibitem[\citeproctext]{ref-galenkamp2017measurement}
Galenkamp, H., Stronks, K., Snijder, M. B., \& Derks, E. M. (2017).
Measurement invariance testing of the PHQ-9 in a multi-ethnic population
in europe: The HELIUS study. \emph{BMC Psychiatry}, \emph{17}, 1--14.

\bibitem[\citeproctext]{ref-gruebner2017cities}
Gruebner, O., Rapp, M. A., Adli, M., Kluge, U., Galea, S., \& Heinz, A.
(2017). Cities and mental health. \emph{Deutsches {Ä}rzteblatt
International}, \emph{114}(8), 121.

\bibitem[\citeproctext]{ref-howlett2001radial}
Howlett, R. J. (2001). \emph{Radial basis function networks 1: Recent
developments in theory and applications}.

\bibitem[\citeproctext]{ref-hyttinen2012learning}
Hyttinen, A., Eberhardt, F., \& Hoyer, P. O. (2012). Learning linear
cyclic causal models with latent variables. \emph{The Journal of Machine
Learning Research}, \emph{13}(1), 3387--3439.

\bibitem[\citeproctext]{ref-lindsay2000moment}
Lindsay, B. G., Pilla, R. S., \& Basak, P. (2000). Moment-based
approximations of distributions using mixtures: Theory and applications.
\emph{Annals of the Institute of Statistical Mathematics}, \emph{52},
215--230.

\bibitem[\citeproctext]{ref-mooij2020constraint}
Mooij, J. M., \& Claassen, T. (2020). Constraint-based causal discovery
using partial ancestral graphs in the presence of cycles.
\emph{Conference on Uncertainty in Artificial Intelligence}, 1159--1168.

\bibitem[\citeproctext]{ref-neapolitan2004learning}
Neapolitan, R. E. et al. (2004). \emph{Learning bayesian networks} (Vol.
38). Pearson Prentice Hall Upper Saddle River.

\bibitem[\citeproctext]{ref-park2024discovering}
Park, K., Waldorp, L. J., \& Ryan, O. (2024). Discovering cyclic causal
models in psychological research. \emph{Advances. In/Psychology},
\emph{2}, e72425.

\bibitem[\citeproctext]{ref-park2025individual}
Park, K., Waldorp, L., \& Vasconcelos, V. V. (2025). \emph{The
individual-and population-level mechanistic implications of statistical
networks of symptoms}.

\bibitem[\citeproctext]{ref-pevalin2017impact}
Pevalin, D. J., Reeves, A., Baker, E., \& Bentley, R. (2017). The impact
of persistent poor housing conditions on mental health: A longitudinal
population-based study. \emph{Preventive Medicine}, \emph{105},
304--310.

\bibitem[\citeproctext]{ref-rahimi2007random}
Rahimi, A., \& Recht, B. (2007). Random features for large-scale kernel
machines. \emph{Advances in Neural Information Processing Systems},
\emph{20}.

\bibitem[\citeproctext]{ref-rehm2019global}
Rehm, J., \& Shield, K. D. (2019). Global burden of disease and the
impact of mental and addictive disorders. \emph{Current Psychiatry
Reports}, \emph{21}, 1--7.

\bibitem[\citeproctext]{ref-rohbeck2024}
Rohbeck, M., Clarke, B., Mikulik, K., Pettet, A., Stegle, O., \&
Ueltzhöffer, K. (2024). Bicycle: Intervention-based causal discovery
with cycles. In F. Locatello \& V. Didelez (Eds.), \emph{Proceedings of
the third conference on causal learning and reasoning} (Vol. 236, pp.
209--242). PMLR.

\bibitem[\citeproctext]{ref-ronnblad2019precarious}
Rönnblad, T., Grönholm, E., Jonsson, J., Koranyi, I., Orellana, C.,
Kreshpaj, B., Chen, L., Stockfelt, L., \& Bodin, T. (2019). Precarious
employment and mental health. \emph{Scandinavian Journal of Work,
Environment \& Health}, \emph{45}(5), 429--443.

\bibitem[\citeproctext]{ref-rugulies2023work}
Rugulies, R., Aust, B., Greiner, B. A., Arensman, E., Kawakami, N.,
LaMontagne, A. D., \& Madsen, I. E. (2023). Work-related causes of
mental health conditions and interventions for their improvement in
workplaces. \emph{The Lancet}, \emph{402}(10410), 1368--1381.

\bibitem[\citeproctext]{ref-runge2019detecting}
Runge, J., Nowack, P., Kretschmer, M., Flaxman, S., \& Sejdinovic, D.
(2019). Detecting and quantifying causal associations in large nonlinear
time series datasets. \emph{Science Advances}, \emph{5}(11), eaau4996.

\bibitem[\citeproctext]{ref-sethuraman2023nodags}
Sethuraman, M. G., Lopez, R., Mohan, R., Fekri, F., Biancalani, T., \&
Hütter, J.-C. (2023). NODAGS-flow: Nonlinear cyclic causal structure
learning. \emph{International Conference on Artificial Intelligence and
Statistics}, 6371--6387.

\bibitem[\citeproctext]{ref-snijder2017cohort}
Snijder, M. B., Galenkamp, H., Prins, M., Derks, E. M., Peters, R. J.,
Zwinderman, A. H., \& Stronks, K. (2017). Cohort profile: The healthy
life in an urban setting (HELIUS) study in amsterdam, the netherlands.
\emph{BMJ Open}, \emph{7}(12), e017873.

\bibitem[\citeproctext]{ref-spirtes2001causation}
Spirtes, P., Glymour, C., \& Scheines, R. (2001). \emph{Causation,
prediction, and search}. MIT press.

\bibitem[\citeproctext]{ref-spirtes_causal_1995}
Spirtes, P., Meek, C., \& Richardson, T. (1995). Causal inference in the
presence of latent variables and selection bias. \emph{Proceedings of
the {Eleventh} Conference on {Uncertainty} in Artificial Intelligence},
499--506.

\bibitem[\citeproctext]{ref-storn1997differential}
Storn, R., \& Price, K. (1997). Differential evolution--a simple and
efficient heuristic for global optimization over continuous spaces.
\emph{Journal of Global Optimization}, \emph{11}, 341--359.

\bibitem[\citeproctext]{ref-strobl2019}
Strobl, E. V. (2019). A constraint-based algorithm for causal discovery
with cycles, latent variables and selection bias. \emph{International
Journal of Data Science and Analytics}, \emph{8}(1), 33--56.
\url{https://doi.org/10.1007/s41060-018-0158-2}

\bibitem[\citeproctext]{ref-strobl2019approximate}
Strobl, E. V., Zhang, K., \& Visweswaran, S. (2019). Approximate
kernel-based conditional independence tests for fast non-parametric
causal discovery. \emph{Journal of Causal Inference}, \emph{7}(1),
20180017.

\bibitem[\citeproctext]{ref-stronks2013unravelling}
Stronks, K., Snijder, M. B., Peters, R. J., Prins, M., Schene, A. H., \&
Zwinderman, A. H. (2013). Unravelling the impact of ethnicity on health
in europe: The HELIUS study. \emph{BMC Public Health}, \emph{13}, 1--10.

\bibitem[\citeproctext]{ref-van2021advancing}
Van Der Wal, J. M., Van Borkulo, C. D., Deserno, M. K., Breedvelt, J.
J., Lees, M., Lokman, J. C., Borsboom, D., Denys, D., Holst, R. J. van,
Smidt, M. P., et al. (2021). Advancing urban mental health research:
From complexity science to actionable targets for intervention.
\emph{The Lancet Psychiatry}, \emph{8}(11), 991--1000.

\bibitem[\citeproctext]{ref-WHO_2022}
World Health Organization. (2022). \emph{Mental disorders}. World Health
Organization.
\url{https://www.who.int/news-room/fact-sheets/detail/mental-disorders}

\bibitem[\citeproctext]{ref-zhang2012kernel}
Zhang, K., Peters, J., Janzing, D., \& Schölkopf, B. (2012).
Kernel-based conditional independence test and application in causal
discovery. \emph{arXiv Preprint arXiv:1202.3775}.

\end{CSLReferences}

\section{Appendix}\label{sec-appendix}

\subsection{HELIUS study}\label{helius-study}

The HELIUS (HEalthy LIfe in an Urban Setting) study is a large-scale,
multiethnic cohort study conducted in Amsterdam. Participants were
randomly selected from the municipality register and stratified by
ethnic origin to ensure balanced representation across six major groups.
Invitation letters were sent by mail, followed by a reminder after two
weeks. Non-respondents were contacted via home visits where applicable.

Of those invited, approximately 55\% responded (Dutch: 55\%, Surinamese:
62\%, Ghanaian: 57\%, Turkish: 46\%, Moroccan: 48\%). Among those
contacted, 50\% agreed to participate (Dutch: 60\%, Surinamese: 51\%,
Ghanaian: 61\%, Turkish: 41\%, Moroccan: 43\%), resulting in an overall
participation rate of 28\%.

Participants completed either a digital or paper version of the
questionnaire, with assistance offered when needed, and received a
confirmation letter for a physical examination appointment.

At baseline (2011--2015), 24,780 individuals were enrolled. Of these,
23,936 participants completed the questionnaire that included items
relevant to precariousness. After excluding individuals with missing
data on key variables, the final analytical sample consisted of 21,628
participants.

The HELIUS study received approval from the Medical Ethics Committee of
the Academic Medical Center (AMC), and written informed consent was
obtained from all participants prior to enrollment.

\subsection{Causal Discovery Primer}\label{causal-discovery-primer}

As shown in Table 1, the resulting graphs from FCI and CCI differ
slightly (\emph{PAG}: partial ancestral graph; \emph{MAAG}: maximal
almost ancestral graph) due to their reliance on different underlying
assumptions. Despite these differences, both graphs belong to the class
of \emph{ancestral graphs}, which are designed to encode causal
relationships between variables, where the presence of an edge indicates
causal \emph{ancestry.} In these graphs, directed edges,
\(A \stararrow B\), indicate that \(B\) is not an ancestor of \(A\) in
every graph within the Markov equivalence class, \(Equiv(G)\). The
Markov equivalence class represents a set of graphs that encode the same
conditional independence relationships, ensuring that the same
\emph{d-separation} conditions hold across all graphs in the class
(Spirtes et al., 2001). Conversely, an edge marked as \(A \startail B\)
indicates that \(B\) is an ancestor of \(A\) across all graphs in
\(Equiv(G)\). Circle endpoints, \(A \starcirc B\), represent ambiguity
in the ancestral relationship, meaning \(B\)'s ancestral status relative
to \(A\) varies across graphs in \(Equiv(G)\).\footnote{\(*\) serves as
  a \textit{meta-symbol}, representing one of the three possible
  edge-endpoints. For instance, \(A \tailstar B\) can indicate any of
  the following edges: \(A\) \textemdash~\(B\), \(A\tailarrow B\), or
  \(A \tailcirc B\) (Park et al., 2024).} Finally, when an edge is
represented as \(A \arrowarrow B\), it implies that neither \(A\) nor
\(B\) is an ancestor of the other, suggesting the presence of a latent
confounder influencing both variables.

\begin{table}[ht]
\centering
\small % Make the font size smaller
\caption{Assumptions of causal discovery algorithms}
\begin{tabular}{lccccc}
\toprule
Algorithm & Acyclicity & \makecell{Causal \\ sufficiency} & 
\makecell{Absence of \\ selection bias} & Linearity & Output \\ 
\midrule
PC  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & CPDAG \\ 
FCI & $-^{a}$ & $\checkmark$ & $-^{a}$ & $-^{a}$ & PAG \\ 
CCI & x & x & x & $\checkmark$ & \makecell{\footnotesize(partially oriented) \\ MAAG} \\ 
\bottomrule
\end{tabular}
\caption*{\footnotesize{\textit{Note}. $^{a}$The FCI algorithm, introduced by Spirtes (1995), is a constraint-based causal discovery method for DAGs that accounts for latent confounding and selection bias. Mooij \& Claassen (2020) later showed its applicability to cyclic causal discovery with latent confounding under general faithfulness and Markov conditions, assuming non-linear causal relationships.}}
\end{table}

The algorithms also differ in how they detect cycles and represent
cyclic relationships in their resulting graphs. In CCI's MAAG, \(A\)
\textemdash~\(B\) indicates that \(A\) is an ancestor of \(B\), and
simultaneously \(B\) is an ancestor of \(A\), referring to a cyclic
relationship between \(A \stackedarrows B\). FCI, on the other hand,
identifies potential cycles more subtly. In its PAG, fully-connected
nodes with circle endpoints (\(\circirc\)) may suggest the presence of
cyclic structures. FCI, therefore, provides a sufficient condition to
distinguish variables that are not part of a cycle, offering a more
nuanced approach to handling cyclic relationships (Mooij \& Claassen,
2020).

Table 1 highlights further differences in the assumptions underlying
these algorithms. CCI operates under the assumption of a linear system,
whereas FCI, particularly when used to infer cyclic relationships,
assumes a non-linear system without selection bias and adheres to more
general faithfulness and Markov conditions (i.e.,
\emph{\(\sigma\)-separation} and \emph{\(\sigma\)-faithfulness} setting)
(Forré \& Mooij, 2018). When the respective assumptions of each
algorithm are met, their inferred cyclic relationships align with those
illustrated in Figure~\ref{fig-examplegraphs}.

While CCI's MAAG provides an explicit representation of cyclic
relationships, it has certain theoretical limitations, as the MAAG may
not always fully preserve d-separation relations from the original
graph, \(\mathcal{G}\) (Strobl, 2019). Given these considerations, we
examine and compare results from both algorithms, placing more weight on
the FCI results, which are presented in the main results section, while
discussing CCI results where relevant, with detailed findings provided
in the \hyperref[sec-appendix]{Appendix}. A full discussion of causal
discovery concepts and algorithmic details is beyond the scope of this
paper; however, readers seeking a more in-depth understanding can refer
to Park et al. (2024) for a comprehensive exploration of these methods
and their applications.

\begin{figure}

\begin{minipage}{0.33\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/original_DCG.png}

}

\subcaption{\label{fig-examplegraphs-1}Example graph \(\mathcal{G}\)}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/FCI_graph.png}

}

\subcaption{\label{fig-examplegraphs-2}Corresponding MAAG}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/CCI_graph.png}

}

\subcaption{\label{fig-examplegraphs-3}Corrsponding PAG}

\end{minipage}%

\caption{\label{fig-examplegraphs}Example graph \(\mathcal{G}\)
featuring a cycle and the corresponding MAAG from CCI and PAG from FCI.}

\end{figure}%

The graph produced by the PC algorithm is a \emph{CPDAG} (completed
partially directed acyclic graph), where directed edges (\(A\) → \(B\))
indicate that \(A\) is a direct cause (parent) of \(B\). Unlike FCI and
CCI, the CPDAG does not include circle symbols. Instead, when the PC
algorithm cannot determine the direction of causality, it represents
this uncertainty with bidirectional arrows. While the PC algorithm
serves as a useful reference, its strict assumptions --- acyclicity and
the absence of latent confounders---limit its applicability in more
complex scenarios. For this reason, our primary focus remains on the
results obtained from FCI and CCI, with all PC algorithm results
provided in the \hyperref[sec-appendix]{Appendix} for completeness.

\subsection{Precariousness factors by
Leonie}\label{precariousness-factors-by-leonie}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  EMPLOYMENT PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_Arbeidsparticipatie}: Working status
\item
  \texttt{H1\_WerkSit}: Which work situation most applies to you?
\item
  \texttt{H1\_RecentErv8}: Experiences past 12 months: h. You were
  sacked from your job or became unemployed (\emph{reverse})
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  FINANCIAL PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_InkHhMoeite}: During the past year, did you have problems
  managing your household income?
\item
  \texttt{H1\_RecentErv9}: Experiences past 12 months: i. You had a
  major financial crisis (\emph{reverse})
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  HOUSING PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{veilig\_2012}: Score safety (veiligheid) in 2012
  (\emph{reverse})
\item
  \texttt{vrz\_2012}: Score level of resources (niveau voorzieningen) in
  2012 (\emph{reverse})
\item
  \texttt{P\_HUURWON}: Percentage Huurwoningen
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  CULTURAL PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_Discr\_sumscore}: Perceived discrimination: sum score of 9
  items (range 9-45)
\item
  \texttt{H1\_SBSQ\_meanscore}: Health literacy: SBSQ meanscore (range
  1-5) (\emph{reverse})
\item
  \texttt{A\_BED\_RU}: Aantal bedrijfsvestigingen; cultuur, recreatie,
  overige diensten (\emph{reverse})
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  SOCIAL PRECARIOUSNESS
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{H1\_RecentErv5}: Experiences past 12 months: e. Your steady
  relationship ended (\emph{reverse})
\item
  \texttt{H1\_RecentErv6}: Experiences past 12 months: f.~A long-term
  friendship with a good friend or family member was broken off
  (\emph{reverse})
\item
  \texttt{H1\_RecentErv7}: Experiences past 12 months: g. You had a
  serious problem with a good friend or family member, or neighbour
  (\emph{reverse})
\item
  \texttt{H1\_SSQT}: SSQT (frequency of social contact): sum score of 5
  items (range 5-20) (\emph{reverse})
\item
  \texttt{H1\_SSQSa}: SSQS (adequacy of social contact): sum score of 5
  items, category 3 and 4 not combined (range 5-20) (\emph{reverse})
\end{itemize}

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-12-1.pdf}
\end{center}

\begin{itemize}
\item
  \textbf{High} Correlations: \texttt{emp\_stat} (employment status) and
  \texttt{work\_sit} (work situation) have a strong positive correlation
  of 0.82. This suggests that individuals with higher employment status
  tend to have more secure or favorable work situations.
  \texttt{soc\_freq} (social contact frequency) shows a strong positive
  correlation with \texttt{soc\_adq} (social adequacy) at 0.61. This
  indicates that individuals with more frequent social contact also tend
  to have higher perceived adequacy of social interactions.
\item
  \textbf{Moderate} Correlations: \texttt{nb\_safe} (neighborhood
  safety) and \texttt{nb\_res} (resources) have a moderate positive
  correlation of 0.39, suggesting that areas with higher safety also
  have better resources. \texttt{hea\_lit} (health literacy) has
  moderate correlations with \texttt{emp\_stat} (0.26) and
  \texttt{work\_sit} (0.25), which could mean that higher health
  literacy is associated with better employment situations.
  \texttt{frd\_brk12} (friendship breakups) and \texttt{conf12}
  (conflicts) have a notable correlation of 0.43, indicating a
  relationship between having conflicts and friendship losses.
\item
  \textbf{Low to Moderate} Correlations in Financial Precariousness:
  \texttt{inc\_dif} (income difficulties) has a moderate correlation
  with \texttt{fincri12} (financial crisis) at 0.49. This aligns with
  the expected relationship, where individuals who experience general
  income difficulties are more likely to report financial crises.
\item
  \textbf{Low} Correlations (0.1 - 0.2): Many variables, such as
  \texttt{discrim} (discrimination), \texttt{unemp12} (unemployment
  experience), and \texttt{rel\_end12} (relationship end), have low
  correlations with other variables, suggesting relatively independent
  relationships in the context of this dataset.
\end{itemize}

\subsection{Exploratory Factor Analysis
(EFA)}\label{exploratory-factor-analysis-efa}

\begin{center}
\includegraphics[width=0.55\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-13-1.pdf}
\end{center}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-14-1.pdf}
\end{center}

\subsubsection{Factor Loadings (Pattern
Matrix)}\label{factor-loadings-pattern-matrix}

\begin{itemize}
\tightlist
\item
  \textbf{MR1}: High loadings on \texttt{emp\_stat} and
  \texttt{work\_sit} suggest this factor captures \emph{employment}
  precariousness.
\item
  \textbf{MR2}: Strong loadings on \texttt{soc\_freq} and
  \texttt{soc\_adq} indicate \emph{social} precariousness.
\item
  \textbf{MR3}: Key items like \texttt{frd\_brk12}, \texttt{conf12}, and
  \texttt{fincri12}, suggest recent \emph{stressful events}.
\item
  \textbf{MR4}: High loadings on \texttt{nb\_res} and \texttt{cul\_rec}
  may reflect \emph{community resources} precariousness.
\item
  \textbf{MR5}: Variables \texttt{nb\_safe} and \texttt{nb\_rent} with
  high loadings indicate \emph{housing} precariousness.
\end{itemize}

\subsubsection{Variance Explained}\label{variance-explained}

The factors cumulatively explain \textbf{38\%} of the variance, with MR1
being the most influential factor. Each factor contributes a smaller
proportion to the total variance (MR1 at 12\%, MR2 at 9\%, etc.).

\subsubsection{Factor Intercorrelations}\label{factor-intercorrelations}

Factors are moderately correlated, especially between \emph{MR1 and
MR5}, and \emph{MR2 and MR3}. This indicates that while distinct, these
factors are related---reasonable in a complex socio-economic context.

\subsubsection{Model Fit Statistics}\label{model-fit-statistics}

RMSEA (0.071) suggest an acceptable fit. Tucker Lewis Index (0.802)
suggests moderate reliability for the model.

\subsubsection{Summary}\label{summary}

The 5-factor model appears interpretable and captures distinct
dimensions of precariousness: \emph{employment, social, stressors,
community resources, and housing precariousness}. Although the overall
fit and explained variance could be stronger, these factors offer
insights into the underlying structure of the data, highlighting key
areas of precariousness.

\subsection{PCA}\label{pca}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-15-1.pdf}
\end{center}

\begin{itemize}
\tightlist
\item
  Component Retention: The scree plot shows a clear ``elbow'' after the
  first component. This steep drop suggests that most variance is
  explained by the first component. After Dimension 5, the percentage of
  explained variance decreases slightly more gradually, indicating
  diminishing returns for adding more components. If we need to choose
  multiple components, retaining the first 5 components seems
  reasonable, as they capture most of the variance (cumulatively
  explaining about 54.7\% of the total variance).
\end{itemize}

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-1.pdf}
\end{center}

\subsubsection{Explained variance (contributions) of
variables}\label{explained-variance-contributions-of-variables}

It shows the importance of variables within each component.

\begin{itemize}
\item
  \textbf{Dim1}: High contributions are observed from
  \texttt{emp\_stat}, \texttt{work\_sit}, \texttt{inc\_dif}, and
  \texttt{fincri12}, suggesting that this dimension captures aspects of
  \emph{employment and financial} security.
\item
  \textbf{Dim2}: While \texttt{emp\_stat} and \texttt{work\_sit} overlap
  with Dim1, the strong contributions from \texttt{frd\_brk12} and
  \texttt{rel\_end12} indicate that this dimension captures a focus on
  \emph{recent relationship stressors}.
\item
  \textbf{Dim3}: \texttt{cul\_rec}, \texttt{nb\_res} have the highest
  contributions, indicating this dimension likely represents
  \emph{community and cultural} factors.
\item
  \textbf{Dim4}: \texttt{soc\_freq} and \texttt{soc\_adq} stand out in
  this dimension, suggesting an emphasis on \emph{social}
  precariousness.
\item
  \textbf{Dim5}: \texttt{nb\_safe} and \texttt{nb\_rent} are the top
  contributors, pointing to \emph{housing} security as key themes in
  this component.
\end{itemize}

\subsubsection{Cos² Values}\label{cosuxb2-values}

Cos² (squared cosine) values, or the quality of representation, show how
well each variable is represented by each dimension. where higher cos²
values (closer to 1) indicate better representation of a variable by a
component.

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-17-1.pdf}
\end{center}

\begin{itemize}
\item
  \textbf{Dim.1}: Variables \texttt{emp\_stat}, \texttt{work\_sit},
  \texttt{inc\_dif}, and \texttt{fincri12} show high cos² values,
  meaning that PC1 primarily captures variations in employment and
  financial difficulties. This component could represent
  \emph{employment \& finance} precariousness.
\item
  \textbf{Dim.2}: Variables \texttt{work\_sit}, \texttt{emp\_stat},
  \texttt{frd\_brk12}, and \texttt{conf12} are well-represented in this
  component, suggesting PC2 captures aspects of \emph{recent
  relationship stressors}.
\item
  \textbf{Dim.3}: Variables \texttt{nb\_res} and \texttt{cul\_rec} load
  strongly on PC3. This may represent community or cultural resources,
  indicating that this component is associated with \emph{neighborhood
  resources}.
\item
  \textbf{Dim.4}: This component has high cos² values for
  \texttt{nb\_res}, \texttt{cul\_rec}, \texttt{soc\_freq}, and
  \texttt{soc\_adq.} While \texttt{nb\_res} and \texttt{cul\_rec} are
  also prominent in PC3, PC4 uniquely captures nuanced differentiation
  in \emph{social} precariousness.
\item
  \textbf{Dim.5}: \texttt{nb\_safe} and \texttt{nb\_rent} are
  well-represented by PC5. This component might capture \emph{housing}
  precariousness.
\end{itemize}

\subsection{ICA}\label{ica}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-18-1.pdf}
\end{center}

\subsubsection{Dominant Variables per
Component:}\label{dominant-variables-per-component}

For each Independent Component (IC), we can identify variables with
\emph{high absolute} values in each column. These values indicate that
the IC captures a strong, independent signal associated with these
variables.

\begin{itemize}
\item
  \textbf{IC1}: \texttt{soc\_freq} and \texttt{soc\_adq} have strong
  negative loadings on this component, indicating that this component
  might represent \emph{social} precariousness.
\item
  \textbf{IC2}: \texttt{frd\_brk12}, \texttt{conf12},
  \texttt{rel\_end12}, \texttt{fincri12} and \texttt{unemp12} have the
  most substantial loadings on this component, all with negative signs.
  This might point to a \emph{recent relational or social stressor}
  component.
\item
  \textbf{IC3}: \texttt{nb\_res} and \texttt{cul\_rec} show notable
  negative loadings, pointing to a focus on \emph{community resource}
  precariousness.
\item
  \textbf{IC4}: High loadings for \texttt{nb\_safe}, \texttt{nb\_rent},
  \texttt{nb\_res}, \texttt{cul\_rec}, and \texttt{discrim} suggest a
  theme of \emph{housing and community-based} precariousness, reflecting
  both safety and social challenges within the neighborhood context.
\item
  \textbf{IC5}: \texttt{emp\_stat} and \texttt{work\_sit} both have
  strong negative loadings on this component, suggesting it captures
  \emph{employment} precariousness.
\end{itemize}

\subsection{Hierarchical clustering}\label{hierarchical-clustering}

\subsubsection{Using Euclidean distance}\label{using-euclidean-distance}

\begin{itemize}
\tightlist
\item
  Ward.D's method: Minimizes the variance within clusters, producing
  more compact and spherical clusters.
\item
  Single linkage: Groups clusters based on the minimum distance between
  points.
\item
  Complete linkage: Groups clusters based on the maximum distance
  between points.
\item
  Average linkage: Uses the average distance between all pairs of points
  in the two clusters.
\end{itemize}

\begin{figure}

\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-19-1.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-19-2.pdf}
\end{center}
\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-19-3.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-19-4.pdf}
\end{center}
\end{minipage}%

\end{figure}%

\paragraph{Consistent Groupings (Across All or Most
Methods)}\label{consistent-groupings-across-all-or-most-methods}

\begin{itemize}
\item
  \texttt{emp\_stat} and \texttt{work\_sit}: This pair consistently
  clusters together across all linkage methods, suggesting that they are
  closely related variables, likely capturing a similar aspect of the
  data (possibly employment status or employment-related information).
\item
  \texttt{nb\_safe}, \texttt{nb\_res}, and \texttt{nb\_rent}: These
  variables are often grouped closely in several methods (especially
  Ward.D, average, and complete linkage). This suggests a similarity or
  common theme among them, potentially related to neighborhood or
  housing precariousness.
\item
  \texttt{soc\_freq} and \texttt{soc\_adq}: These two variables
  frequently cluster together, indicating they likely measure aspects of
  social frequency and adequacy in similar ways. They appear together in
  Ward.D, average, and complete linkage.
\item
  \texttt{frd\_brk12} and \texttt{conf12}: These variables are often
  clustered closely (though they sometimes join with other variables
  like rel\_end12), suggesting they may capture aspects of relationship
  or social conflict. This pair appears in close proximity, especially
  in average and Ward.D.
\end{itemize}

\paragraph{Inconsistent Groupings (Variability Across
Methods)}\label{inconsistent-groupings-variability-across-methods}

\begin{itemize}
\item
  \texttt{hea\_lit}: This variable shows inconsistent clustering across
  methods. In Ward.D, it joins with \texttt{fincri12}, while in other
  methods, it's often more isolated or grouped with variables that do
  not appear similar. This may suggest that \texttt{hea\_lit} does not
  strongly correlate with other variables, or it has multidimensional
  aspects affecting its grouping across methods.
\item
  \texttt{discrim}: This variable also shows variable groupings. In
  Ward.D, it is grouped with \texttt{hea\_lit}, while in other methods
  (e.g., complete and single linkage), it clusters differently,
  sometimes on its own. This variability may indicate that
  \texttt{discrim} has weaker associations with the main clusters in the
  data or overlaps partially with multiple clusters.
\item
  Social and Financial Variables (\texttt{inc\_dif}, \texttt{fincri12},
  \texttt{unemp12}): These variables appear together in some methods
  (e.g., Ward.D clusters \texttt{fincri12} and \texttt{inc\_dif}), but
  in others, they are spread out. This inconsistency suggests that
  social and financial variables may not have strong or consistent ties
  across different methods, perhaps due to capturing different aspects
  of precariousness.
\end{itemize}

\paragraph{Summary}\label{summary-1}

The consistent clusters are likely capturing distinct thematic
dimensions of the data (e.g., employment, housing, social contact),
while the inconsistent variables may reflect multifaceted or weakly
correlated attributes that do not fit neatly into one cluster.

\subsubsection{Using Mutual Information}\label{using-mutual-information}

Using mutual information (MI) as a basis for hierarchical clustering
differs from using traditional distance measures (like Euclidean
distance) in a few key ways.

\begin{center}
\includegraphics[width=0.6\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-20-1.pdf}
\end{center}

\paragraph{Comparison to Euclidean Distance
Clustering}\label{comparison-to-euclidean-distance-clustering}

\begin{itemize}
\item
  \textbf{Housing and Community} Cluster: The variables
  \texttt{nb\_safe}, \texttt{nb\_res}, \texttt{nb\_rent}, and
  \texttt{cul\_rec} cluster together, indicating a strong association
  among housing-related and community-based factors. This suggests a
  shared theme of housing or community precariousness. This grouping is
  also observed in the Euclidean-based clustering, but it appears more
  tightly connected here, potentially due to the non-linear
  relationships highlighted by mutual information.
\item
  \textbf{Employment and Social Support} Cluster: \texttt{emp\_stat} and
  \texttt{work\_sit} form a cluster, linking employment status and work
  situation together as they did in Euclidean-based clustering. These
  remain closely associated regardless of the distance metric used.
  \texttt{soc\_freq} and \texttt{soc\_adq}, related to social contact
  frequency and adequacy, cluster nearby, indicating they have a
  stronger non-linear relationship with employment variables. This is a
  subtle difference as Euclidean distance might not capture this
  association as effectively.
\item
  \textbf{Financial Stressor} Cluster: \texttt{inc\_dif} and
  \texttt{fincri12}, representing income difficulties and recent
  financial crises, consistently cluster together in both approaches,
  showing a strong association, likely linear. However, mutual
  information-based clustering links these financial stressors with
  social support variables, suggesting that financial challenges may
  have complex dependencies with social support in this dataset.
\item
  \textbf{Relational Stressor} Cluster: \texttt{frd\_brk12},
  \texttt{conf12}, \texttt{discrim}, \texttt{hea\_lit},
  \texttt{unemp12}, and \texttt{rel\_end12} form a \emph{looser} cluster
  focused on social and relational stressors (e.g., friendship breakup,
  conflicts, and discrimination). Compared to Euclidean clustering,
  \texttt{discrim} and \texttt{hea\_lit} (health literacy) appear closer
  to relational stressors here, indicating that non-linear relationships
  might play a larger role in linking these variables.
\end{itemize}

\paragraph{Summary}\label{summary-2}

In conclusion, mutual information-based clustering provides an
alternative perspective that can reveal more intricate associations
between variables, especially for those with non-linear relationships.
Compared to Euclidean clustering, it shows a similar high-level
structure but emphasizes nuanced connections between variables,
particularly around social support, employment, and financial stress.

\subsection{Conclusions on Precariousness
factors}\label{conclusions-on-precariousness-factors}

Based on the consistent findings across multiple analyses, we decided to
exclude the variables \texttt{discrim}, \texttt{hea\_lit},
\texttt{umemp12}, and \texttt{rel\_end12}, as they do not clearly belong
to any specific precariousness factor nor exhibit strong associations
with depression (see the correlation table above). Therefore, we propose
retaining the following key precariousness factors:

\begin{itemize}
\tightlist
\item
  Employment Precariousness: \texttt{emp\_stat}, \texttt{work\_sit}
\item
  Social Precariousness: \texttt{soc\_freq}, \texttt{soc\_adq}
\item
  Housing Precariousness: \texttt{nb\_safe}, \texttt{nb\_res},
  \texttt{nb\_rent}, \texttt{cul\_rec}
\item
  Recent Relational Stressors: \texttt{frd\_brk12}, \texttt{conf12}
\item
  Recent Financial Stressors: \texttt{fincri12}, \texttt{inc\_diff}
\end{itemize}

We construct each precariousness factor by calculating the mean value of
the combined variables. Below, we present the updated correlation table
for the newly composed factors, along with the corresponding
distributions of all variables to be used in the causal discovery
analysis.

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-21-1.pdf}
\end{center}

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-22-1.pdf}
\end{center}

\subsection{Results from CCI algorithm}\label{sec-cci}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/CCI_depsum.png}

}

\subcaption{\label{fig-sum-cci-1}CCI MAAG}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/depsum_mat_cci.pdf}

}

\subcaption{\label{fig-sum-cci-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sum-cci}Resulting graph of precarity factors and
depression sum score using CCI and proportion of edge endpoint types.}

\end{figure}%

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/symptom_graph_CCI.png}

}

\subcaption{\label{fig-sym-cci-1}CCI MAAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/symptom_mat_cci.pdf}

}

\subcaption{\label{fig-sym-cci-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-sym-cci}Resulting graph of precarity factors and
individual depression symptoms using CCI and proportion of edge endpoint
types.}

\end{figure}%

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_graph_CCI.png}

}

\subcaption{\label{fig-presum-cci-1}CCI MAAG}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/presum_mat_cci.pdf}

}

\subcaption{\label{fig-presum-cci-2}Proportion matrix}

\end{minipage}%

\caption{\label{fig-presum-cci}Resulting graph of precarity sum score
and individual depression symptoms using CCI and proportion of edge
endpoint types.}

\end{figure}%

\clearpage

\subsection{Results from PC algorithm}\label{results-from-pc-algorithm}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/sum_PC_all.png}

}

\subcaption{\label{fig-pc_sum-1}Using both GaussianCI and RCoT}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/sum_PC_RCoTonly.png}

}

\subcaption{\label{fig-pc_sum-2}Using only RCoT}

\end{minipage}%

\caption{\label{fig-pc_sum}Resulting graphs of precarity factors and
depression sum score using PC.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/PC_symptom.png}

}

\caption{\label{fig-pc_sym}Resulting graphs of precarity factors and
individual depression symptoms using PC.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{img/PC_presum.png}

}

\caption{\label{fig-pc_presum}Resulting graphs of precarity sum score
and individual depression symptoms using PC.}

\end{figure}%

\clearpage

\subsection{Randomized Conditional Independence / Correlation Test (RCIT
\& RCoT)}\label{sec-rcot}

RCIT (Randomized Conditional Independence Test) and RCoT (Randomized
conditional Correlation Test) are advanced methods for scalable
conditional independence (CI) testing, offering computational efficiency
while maintaining the accuracy of kernel-based approaches. These methods
evaluate conditional independence between two variables \(X\) and \(Y\)
given a third variable \(Z\) while addressing computational challenges
inherent in kernel-based CI tests. In this section, we provide a
high-level overview of RCIT and RCoT based on (Strobl et al., 2019).

\subsubsection{Kernel-Based Conditional Independence
Testing}\label{kernel-based-conditional-independence-testing}

Traditional kernel-based CI tests, such as the Kernel Conditional
Independence Test (KCIT), compute dependencies using the Hilbert-Schmidt
Independence Criterion (HSIC) in reproducing kernel Hilbert spaces
(RKHS) (Zhang et al., 2012). KCIT uses the following hypothesis
framework: \[
H_0: X \perp\!\!\!\perp Y \,|\, Z, \quad H_1: X \not\!\perp\!\!\!\perp Y \,|\, Z.
\] The core quantity in KCIT is the partial cross-covariance operator:
\[
\Sigma_{XY \cdot Z} = \Sigma_{XY} - \Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY},
\] where \(\Sigma_{XY}\) represents the cross-covariance operator
between \(X\) and \(Y\), and
\(\Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY}\) removes the dependence
mediated by \(Z\).

The squared Hilbert-Schmidt (HS) norm of \(\Sigma_{XY \cdot Z}\) serves
as the test statistic: \[
\|\Sigma_{XY \cdot Z}\|^2_{HS} = 0 \quad \text{if and only if} \quad X \perp\!\!\!\perp Y \,|\, Z.
\]

KCIT estimates residual dependencies using kernel ridge regression: \[
f^*(z) = K_Z (K_Z + \lambda I)^{-1} f(x),
\] where \(K_Z\) is the kernel matrix for \(Z\), \(f(x)\) is the kernel
feature map for \(X\), and \(\lambda\) is the ridge regularization
parameter. The residual function for \(X\) is: \[
f_\text{res}(x) = f(x) - f^*(z) = R_Z f(x),
\] with: \[
R_Z = I - K_Z (K_Z + \lambda I)^{-1}.
\] The kernel matrix for residualized \(X\) is: \[
K_{X \cdot Z} = R_Z K_X R_Z,
\] and similarly for \(Y\), \(K_{Y \cdot Z} = R_Z K_Y R_Z\).

The test statistic is computed as: \[
T_{XY \cdot Z} = \frac{1}{n^2} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
\] which estimates the Hilbert-Schmidt (HS) norm of the partial
cross-covariance operator. To ensure convergence, KCIT scales the
statistic by \(n\): \[
S_K = n T_{XY \cdot Z}.
\] The null hypothesis \(H_0\) is rejected if \(S_K\) exceeds a
threshold determined by permutation or moment-matching-based null
distribution (Lindsay et al., 2000).

\subsubsection{Random Fourier Features
(RFFs)}\label{random-fourier-features-rffs}

Kernel-based methods like KCIT face scalability issues, as they involve
operations on \(n \times n\) kernel matrices, which scale quadratically
with the sample size \(n\). RCIT and RCoT overcome this bottleneck using
\emph{Random Fourier Features (RFFs)} to approximate kernel operations
efficiently.

\paragraph{Bochner's Theorem}\label{bochners-theorem}

Bochner's theorem provides the foundation for RFFs, stating that any
continuous shift-invariant kernel \(k(x, y)\) can be expressed as: \[
k(x, y) = \int_{\mathbb{R}^p} e^{i \omega^\top (x - y)} \, dP_\omega,
\] where \(P_\omega\) is the spectral distribution of the kernel. For
the widely used RBF kernel: \[
k(x, y) = \exp\left(-\frac{\|x - y\|^2}{2\sigma^2}\right),
\] \(P_\omega\) follows a Gaussian distribution:
\(\omega \sim \mathcal{N}(0, \sigma^2 I)\).

\paragraph{RFF Approximation}\label{rff-approximation}

Using Monte Carlo sampling, the kernel function is approximated as: \[
k(x, y) \approx \phi(x)^\top \phi(y),
\] where \(\phi(x)\) is the random Fourier feature mapping: \[
\phi(x) = \sqrt{\frac{2}{D}} \cos(W^\top x + b),
\] with \(W \sim \mathcal{N}(0, \sigma^2 I)\) and
\(b \sim \text{Uniform}(0, 2\pi)\). Here, \(D\) is the number of Fourier
features, which balances computational efficiency and approximation
accuracy.

\subsubsection{Differences Between RCIT and
RCoT}\label{differences-between-rcit-and-rcot}

RCIT and RCoT differ in their test statistics, computational efficiency,
and practical performance, which makes them suited for different
scenarios in causal discovery. RCIT evaluates the Hilbert-Schmidt norm
of the full partial cross-covariance operator, providing a general test
for conditional independence but at a higher computational cost. RCoT
simplifies the process by using the Frobenius norm of a
finite-dimensional residualized cross-covariance matrix, significantly
reducing complexity and improving scalability.

These distinctions are particularly important for large-scale datasets,
where RCoT's computational efficiency makes it a practical choice for
high-dimensional causal discovery tasks.

\paragraph{RCIT: Randomized Conditional Independence
Test}\label{rcit-randomized-conditional-independence-test}

RCIT tests full conditional independence by examining the squared
Hilbert-Schmidt (HS) norm of the partial cross-covariance operator
\(\Sigma_{XY \cdot Z}\): \[
S_K = n T_{XY \cdot Z} = \frac{1}{n} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
\] where \(T_{XY \cdot Z}\) is an empirical estimate of
\(\|\Sigma_{XY \cdot Z}\|^2_{HS}\). The null and alternative hypotheses
are: \[
H_0: \|\Sigma_{XY \cdot Z}\|^2_{HS} = 0, \quad H_1: \|\Sigma_{XY \cdot Z}\|^2_{HS} > 0.
\]

RCIT is a general test for conditional independence but becomes
computationally demanding as the size of \(Z\) increases, due to the
high-dimensional kernel operations required.

\paragraph{RCoT: Randomized Conditional Correlation
Test}\label{rcot-randomized-conditional-correlation-test}

RCoT simplifies the testing process by using a finite-dimensional
partial cross-covariance matrix, avoiding full HS norm calculations.
Instead, it uses the Frobenius norm of the residualized cross-covariance
matrix: \[
S' = n \|C_{AB \cdot C}\|_F^2,
\] where \(C_{AB \cdot C}\) represents the residualized cross-covariance
matrix. The hypotheses are: \[
H_0: \|C_{AB \cdot C}\|_F^2 = 0, \quad H_1: \|C_{AB \cdot C}\|_F^2 > 0.
\]

RCoT is computationally efficient and well-suited for large conditioning
sets (\(|Z| \geq 4\)). Its simplicity enables robust calibration of the
null distribution and improved scalability for high-dimensional data.

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

The HELIUS study is conducted by the Amsterdam University Medical
Centers, location AMC and the Public Health Service of Amsterdam. Both
organisations provided core support for HELIUS. The HELIUS study is also
funded by the Dutch Heart Foundation, the Netherlands Organization for
Health Research and Development (ZonMw), the European Union (FP-7), and
the European Fund for the Integration of non-EU immigrants (EIF). We are
most grateful to the participants of the HELIUS study and the management
team, research nurses, interviewers, research assistants and other staff
who have taken part in gathering the data of this study.

\end{document}