### 2025-08-12 draft_v1/draft_v2.qmd
vitor_feedback_2nd

---
title: "From Causal Discovery to Intervention Simulation: Modeling Precariousness and Depression in the HELIUS Study"
#subtitle: Using the HELIUS dataset
author: 
  - name: Kyuri Park
    affil-id: 1
  - name: Leonie K. Elsenburg
    affil-id: 2
  - name: Mary Nicolao
    affil-id: 2
  - name: Karien Stronks
    affil-id: 2
  - name: Vítor V. Vasconcelos
    affil-id: 1, 3

affiliations:
  - id: 1
    name: \textit{Computational Science Lab, Informatics Institute, University of Amsterdam, PO Box 94323, Amsterdam, 1090GH, the Netherlands}
  - id: 2
    name: \textit{Department of Public and Occupational Health, Amsterdam Public Health Research Institute, Amsterdam UMC, University of Amsterdam, Amsterdam, the Netherland}
  - id: 3
    name: \textit{Institute for Advanced Study, University of Amsterdam, Oude Turfmarkt 147, Amsterdam, 1012GC, the Netherland}   
format: 
  # docx:
  #   number-sections: true
  #   appendix: true
  pdf:
    pdf-engine: pdflatex
    fontfamily: palatino
    toc: true
    number-sections: true
    appendix: true
    colorlinks: true
    keep-tex: true
    template-partials:
      - title.tex
    include-in-header: header.tex
    crossref:
      appendix-title: "App."
      appendix-delim: ":"
date: today
execute:
  echo: false
  warning: false
bibliography: references_draft.bib
nocite: |
  @*
csl: apa.csl
abstract: |
  \noindent This study investigates how the internal structure of depression–precariousness dynamics shapes responsiveness to external support. Using cross-sectional data from the HELIUS study, we first apply cycle-capable causal discovery algorithms to estimate directional dependencies between financial stress, domain-specific precariousness, and depressive symptoms. Our multi-resolution analysis—spanning both aggregate and symptom-level constructs—highlights financial stress as a plausible upstream driver of depression, with symptoms such as sleep disturbance and depressed mood emerging as initiators of broader social vulnerability.
  To translate these structural insights into dynamic predictions, we construct a family of analytically calibrated linear stochastic models. Simulations reveal that variations in internal structure can substantially alter system responses to financial stress reduction. Some models show rapid but transient improvements in depression, while others adjust more slowly yet sustain gains over time. Precariousness tends to respond more immediately across all configurations, though its trajectory also shifts subtly with structural variation.
  Together, these findings demonstrate that intervention outcomes depend not only on the magnitude of external input but also on latent properties of the system. More broadly, this study illustrates how combining causal discovery with dynamical modeling supports principled, simulation-based exploration of intervention scenarios—grounded in data, yet able to probe counterfactual structure. This approach offers a scalable framework for diagnosing hidden system constraints and informing targeted public health strategies in complex psychosocial domains.

  <!-- Understanding the causal mechanisms linking precariousness and depression is critical for developing effective interventions. This study utilizes data from the HELIUS cohort study to explore these relationships using advanced causal discovery methods. By applying algorithms such as FCI, and CCI, and combining traditional Gaussian CI tests with non-parametric approaches like RCoT, we investigate how different indicators of precariousness—including factors related to employment, social relations and relational stress, financial situation, and housing—affect depressed mood, both as a sum score and at the individual symptom level. Our findings reveal that relational stress consistently emerges as a potential causal factor for depressed mood, while symptoms such as sleep disturbances, guilt, and anhedonia are particularly sensitive to external stressors, acting as potential early warning signals or intervention points for prevention. Moreover, the results highlight complexities in the data, including the influence of latent confounders and the challenges of capturing cyclic relationships. Despite some limitations, such as unresolved ambiguities in causal directions and challenges with mixed data distributions, this study demonstrates the utility of causal discovery tools in disentangling the intricate interplay between social and mental health dynamics. By mapping these causal structures into computational models, future research can simulate intervention effects, providing actionable insights to mitigate the impact of precariousness on mental health. This study serves as a foundational effort, offering both methodological advancements and practical implications for addressing depression at a population level. -->
---

```{r, include=FALSE}
library(haven)
library(dplyr)
library(skimr)
#library(bestNormalize)
library(furrr)
library(future)
library(tidyr)
library(GGally)
library(purrr)
library(gt)

library(corrplot)
library(psych)
library(FactoMineR)
library(factoextra)
library(fastICA)
library(umap)

```

```{r, include=FALSE, cache=TRUE}
# Load the dataset 
dat <- read_sav("../data/helius_dat/HELIUS_LEONIE.sav")

# Preprocess and scale the precariousness data
scaled_data <- dat %>%
  # Select precariousness-related variables
                # Employment Precariousness
  dplyr::select(H1_Arbeidsparticipatie, H1_WerkSit, H1_RecentErv8, 
                # Financial Precariousness
                H1_InkHhMoeite, H1_RecentErv9,                      
                # Housing Precariousness
                veilig_2012, vrz_2012, P_HUURWON,                   
                # Cultural Precariousness
                H1_Discr_sumscore, H1_SBSQ_meanscore, A_BED_RU,         
                # Social Precariousness
                H1_RecentErv5, H1_RecentErv6, H1_RecentErv7,       
                H1_SSQT, H1_SSQSa,                                  
                # Depression symptoms / sum scores
                H1_WlbvRecent1, H1_WlbvRecent2, H1_WlbvRecent3, H1_WlbvRecent4, H1_WlbvRecent5, H1_WlbvRecent6, H1_WlbvRecent7, H1_WlbvRecent_89, H1_WlbvRecent10, H1_PHQ9_sumscore,   
                # Ethnicity
                H1_etniciteit,
                # Ags
                H1_lft) %>%                                
  # Replace missing codes with NA 
  mutate(across(everything(), ~ na_if(.x, -9)), 
         across(everything(), ~ na_if(.x, -1))) %>%  
  # Remove NAs 
  na.omit() %>%
  # Convert all columns to numeric
  mutate(across(everything(), as.numeric),
        # Combine categories 5 to 10 into a single category (5-10 as one unemployed)
         H1_WerkSit = case_when(
    H1_WerkSit %in% c(5, 6, 7, 8, 9, 10) ~ 5,  
    .default = H1_WerkSit  # Keep other values as they are 
  ))  %>%
  # Reverse the values of specific variables (reverse transformation)
  mutate(across(c(H1_RecentErv8, H1_RecentErv9, veilig_2012, vrz_2012, 
                  H1_SBSQ_meanscore, A_BED_RU, H1_RecentErv5, H1_RecentErv6, 
                  H1_RecentErv7, H1_SSQT, H1_SSQSa), 
                ~ max(.x) - .x)) %>%
  # scale the data
  mutate(across(!c(H1_etniciteit, H1_lft), scale))
  # Apply Min-Max scaling to all columns
  # mutate(across(everything(), min_max_scaling))

# redefine colnames
colnames(scaled_data) <- c(
  "emp_stat",       # H1_Arbeidsparticipatie
  "work_sit",   # H1_WerkSit
  "unemp12",     # H1_RecentErv8
  "inc_dif",      # H1_InkHhMoeite
  "fincri12",# H1_RecentErv9
  "nb_safe",     # veilig_2012
  "nb_res",   # vrz_2012
  "nb_rent",         # P_HUURWON
  "discrim",      # H1_Discr_sumscore
  "hea_lit",       # H1_SBSQ_meanscore
  "cul_rec",      # A_BED_RU
  "rel_end12",   # H1_RecentErv5
  "frd_brk12", # H1_RecentErv6
  "conf12",         # H1_RecentErv7
  "soc_freq",      # H1_SSQT
  "soc_adq",      # H1_SSQSa
  "anh",       # H1_WlbvRecent1
  "dep",       # H1_WlbvRecent2
  "slp",       # H1_WlbvRecent3
  "ene",       # H1_WlbvRecent4
  "app",       # H1_WlbvRecent5
  "glt",       # H1_WlbvRecent6
  "con",       # H1_WlbvRecent7
  "mot",     # H1_WlbvRecent_89
  "sui",      # H1_WlbvRecent10
  "PHQsum",         # H1_PHQ9_sumscore
  "ethn",         # H1_etniciteit
  "age"          #"H1_lft" 
)
```

# Introduction

Mental health disorders represent a growing global health concern, especially in urbanized regions [@WHO_2022; @gruebner2017cities]. In 2019, one in eight people worldwide were living with a mental health condition, with the highest disability-adjusted life years (DALYs) due to mental and addictive disorders present in high-income countries such as those in Northern Europe, North America, and Australia [@rehm2019global; @WHO_2022]. Despite growing awareness, governments have struggled to design effective responses. Mental health outcomes arise from complex, multi-level dynamics between the social, economic, and spatial structures of urban environments—making both diagnosis and intervention design deeply challenging [@van2021advancing].

Recent work has shifted focus from individual-level vulnerability to broader social and structural determinants. While high-income nations show elevated overall DALYs, within-country disparities reveal that income inequality is a significant predictor of mental health burden [@rehm2019global]. Moreover, precarious working conditions, housing instability, and neighborhood disadvantage have all been linked to increased risk of depression and anxiety [@pevalin2017impact; @fone2014effect; @rugulies2023work; @ronnblad2019precarious]. This growing literature highlights that poor mental health is not merely a personal issue but one rooted in sustained socioeconomic stress and structural uncertainty.

Building on this, recent studies have introduced the concept of precariousness as a multidimensional condition characterized by instability and lack of control across several life domains, including employment, finances, housing, social relations, and cultural belonging [@elsenburg2025clustering; @mckee2017living]. This broader view reveals how different forms of insecurity can co-occur and compound, forming an ecosystem of risk. However, despite growing evidence of association between precariousness and mental health, a fundamental question remains unresolved: *How do these factors causally influence one another?* Most existing studies rely on cross-sectional correlations, leaving open questions about directionality and feedback. Without stronger causal insight, it remains difficult to identify priority targets for intervention or to anticipate and quantify intended and unintended effects.

This study aims to investigate the causal mechanisms linking precariousness and depression and how those mechanisms shape system responses to external intervention. Specifically, we ask: *How does the internal configuration of the precariousness–depression relationship affect the outcomes of stress-reducing interventions?*
To answer this, we combine two approaches. First, we use cycle-capable causal discovery algorithms to infer directional relationships among domain-specific precariousness factors and depressive symptoms.
By examining both aggregated measures (depression sum score and a composite precariousness index) and disaggregated variables (individual depressive symptoms and specific life-domain indicators), we capture broad patterns of relationship as well as fine-grained causal pathways.
Second, we translate these insights into a computational dynamical model that simulates how depression and precariousness co-evolve under varying levels of external stress. 
This model enables us to probe how system properties—such as feedback strength and stochastic noise—govern responsiveness to intervention. 
When the effects of depression symptoms onto the social precarity subsystem are weaker than the reverse path, there is higher sensitivity of depression to intervention, reflecting both a faster reduction post financial intervention but also quicker relapse once financial stresses are reestablished.
By integrating causal discovery with dynamic simulation, we demonstrate that intervention strategies can be tailored to the underlying architecture of vulnerability.

<!-- Mental health problems are rising globally, and populations facing chronic stress and socioeconomic instability bear a disproportionate share of this burden. Despite growing recognition of these challenges, designing effective policies remains difficult—largely because mental health outcomes emerge from complex, interacting factors that do not operate in isolation. To inform better interventions, we must move beyond identifying which factors are associated with mental health problems and instead uncover how those factors influence one another, the underlying causal mechanisms. -->

<!-- Precarious life conditions—such as unstable housing, insecure employment, and financial hardship—are now well established as key correlates of depression and other mental health outcomes. However, the causal dynamics linking these domains remain poorly understood. In particular, it is unclear how different forms of precariousness and depressive symptoms influence one another, and whether their interplay creates self-reinforcing loops that stabilize poor mental health and social adversity over time.  -->

<!-- This study aims to investigate the causal mechanisms linking precariousness and depression, and to examine how those mechanisms shape the system’s response to external intervention. Specifically, we ask: How do different configurations of the precariousness–depression relationship affect the outcomes of interventions aimed at reducing external stress? -->

<!-- To address this, we combine causal discovery and mechanistic modeling. We apply cycle-capable causal discovery algorithms to infer directional relationships among financial stress, precariousness indicators, and depressive symptoms. Our analysis includes both aggregate scores (e.g., PHQ-9 total, precariousness composite index) and disaggregated variables (individual symptoms and domain-specific precariousness), allowing us to capture both broad associations and fine-grained pathways. -->

<!-- Building on the discovered causal patterns, we simulate a set of scenarios using a computational model that captures the co-evolution of depression and precariousness under varying levels of financial stress. This framework allows us to examine how internal system properties—particularly feedback strength and stochastic noise—influence responsiveness to intervention. In some regimes, the system exhibits bistability, resisting change unless pushed past a critical threshold. In others, it transitions gradually, reflecting unistable dynamics. Through this integrative approach, we aim to show how the internal architecture of the depression–precariousness system fundamentally shapes its susceptibility to change—highlighting the importance of structural understanding in designing effective, equitable mental health interventions. -->

# Methods

## Data

We use data from the HELIUS (HEalthy LIfe in an Urban Setting) study [@stronks2013unravelling; @snijder2017cohort]. It  captures the diverse population of the city of Amsterdam by including the six main ethnic groups and provides comprehensive health and lifestyle data, including depressive symptoms as measured by the PHQ-9 [@galenkamp2017measurement]. 
To operationalize indicators of precariousness, we draw on the framework outlined in previous research [@elsenburg2025clustering] and select a set of relevant variables.
To ensure a robust representation of precariousness in our causal discovery models, we conducted exploratory analyses to identify consistent and meaningful data-driven factor structures. These analyses led to the identification of five distinct precariousness factors: three reflecting structural domains—employment, housing, and social relationships—and two capturing recent life stressors—financial and relational. Each factor is composed of multiple observed variables, as outlined below. Additional details on the HELIUS study and the factor extraction process are provided in the [Appendix](#sec-appendix).

-   Employment precariousness: `emp_stat` (current employment status), `work_sit` (nature of work situation).
-   Social precariousness: `soc_freq` (frequency of social contact), `soc_adq` (perceived adequacy of social support).
-   Housing precariousness: `nb_safe` (neighborhood safety), `nb_res` (access to neighborhood resources), `nb_rent` ( proportion of rental housing), `cul_rec` (availability of cultural resources).
-   Recent relational stressors: `frd_brk12` (experience of a friendship breakup), `conf12` (recent interpersonal or relational conflict).
-   Recent financial stressors: `fincri12` (experience of a financial crisis), `inc_diff` ( difficulty managing household income).

Following data cleaning and normalization procedures, the analytical sample comprises 21,628 individuals. In addition to the five identified precariousness factors, we computed an overall precariousness score by summing these five factor scores, thereby capturing a composite measure of accumulated precariousness across domains. Depression is represented using the PHQ-9 [@kroenke2001phq], both as a total sum score and as individual symptom scores. In the subsequent causal discovery analysis, we explore the relationship between depression and precariousness using both aggregated scores and their disaggregated components. See @fig-dist for the overall distributions of the variables used in the analysis.

```{r echo=FALSE, fig.align='center', fig.height=5, fig.width=8, out.width="90%", cache=TRUE}
#| fig-cap: >
#|   Distributions of variables with density overlay.  
#|   *This shows the distributions of the variables used in the causal analysis, plotted as histograms with overlaid density estimates. The x-axis represents the value of each variable after preprocessing. The y-axis shows the corresponding estimated probability density.*
#|   ***P.emp*** = *employment precariousness*;  ***P.hou*** = *housing precariousness*;  ***P.soc*** = *social precariousness*;  ***S.fin*** = *recent financial stressors*;  ***S.rel*** = *recent relational stressors*;  ***precariousness*** = *overall precariousness*;  ***PHQsum*** = *PHQ-9 sum score*;  ***anh*** = *anhedonia*;  ***app*** = *appetite*;  ***con*** = *concentration*;  ***dep*** = *depressed mood*;  ***ene*** = *energy*;  ***glt*** = *guilty*;  ***mot*** = *motor*;  ***slp*** = *sleep*;  ***sui*** = *suicidal*.
#|   
#| label: fig-dist

clust_data <- scaled_data |> 
  mutate(
    P.emp = rowMeans(cbind(emp_stat, work_sit)),
    P.soc = rowMeans(cbind(soc_freq, soc_adq)),
    P.hou = rowMeans(cbind(nb_safe, nb_res, nb_rent, cul_rec)),
    S.rel = rowMeans(cbind(frd_brk12, conf12)),
    S.fin = rowMeans(cbind(fincri12, inc_dif)),
    precariousness = P.emp + P.soc + P.hou + S.rel + S.fin
  ) |>
  dplyr::select(17:34) |>
  as.data.frame()

# Reshape the data to long format
clust_data_long <- clust_data |>
  dplyr::select(-age, -ethn) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot distributions with density lines
ggplot(clust_data_long, aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)), fill = "lightblue", color = "white", bins = 15) +  # Use density for histogram
  geom_density(adjust=2, linewidth=0.3, color = "gray", alpha = 0.5) +  # Add density line in red
  facet_wrap(~ factor(variable, levels =c("P.emp", "P.hou", "P.soc", "S.fin", "S.rel", "precariousness", "anh", "app", "con", "dep", "ene", "glt", "mot", "slp", "sui", "PHQsum")), scales = "free") +
  labs(x = "", y = "") +
  theme_minimal()

```

## Causal Discovery{#sec-analysis}

To uncover directional relationships between precariousness and depression, we use constraint-based causal discovery algorithms that allow for feedback loops and latent confounding. Specifically, we apply the *Fast Causal Inference* (FCI) and *Cyclic Causal Inference* (CCI) algorithms, both of which can detect cycles and non-acyclic structures [@mooij2020constraint; @strobl2019]. 
For reference, we also include the *PC* algorithm, one of the most widely used methods for acyclic causal discovery [@spirtes2001causation].
While our full analysis includes results from three causal discovery algorithms—FCI, CCI, and PC—we focus on FCI in the main text, as it offers a strong balance between flexibility and theoretical soundness. FCI accommodates latent confounding and potential feedback, making it well-suited for the complexity of our data [@mooij2020constraint]. The PC algorithm is included for reference but relies on more restrictive assumptions, such as acyclicity and no unmeasured confounding. CCI, while capable of modeling cycles explicitly, has its own theoretical limitations; for example, its output may not fully preserve the d-separation properties of the true causal graph [@strobl2019]. Results from both PC and CCI are provided in the [Appendix](#sec-appendix) ([@sec-cci; @sec-pc]) for completeness and comparison. For readers interested in detailed explanations of the algorithms, edge interpretations, and graph types (e.g., PAG, MAAG, CPDAG), as well as algorithmic assumptions and limitations, please refer to the [Appendix](#sec-appendix) ([@sec-causalprimer]).

A key challenge in applying these algorithms to the HELIUS dataset is that many variables are non-Gaussian and likely exhibit nonlinear relationships. While kernel-based conditional independence tests (such as KCIT) are well-suited for capturing such complex dependencies, they are computationally intensive—scaling quadratically with sample size due to the need to invert large kernel matrices—making them impractical for large datasets like HELIUS [@rahimi2007random].
To address this, we use the Randomized Conditional Correlation Test (RCoT), a nonparametric alternative that approximates kernel methods using random Fourier features. This reduces computational complexity from quadratic to linear in sample size, substantially lowering runtime while preserving sensitivity to nonlinear relationships [@strobl2019approximate; @zhang2012kernel]. For further details, see @sec-rcot.

We examine the causal structure using three complementary approaches:

1. Aggregate-level analysis: relationships between five domain-specific precariousness factors and the PHQ-9 sum score.
2. Fully disaggregated analysis: relationships between individual precariousness items and individual depressive symptoms.
3. Mixed analysis: relationships between individual symptoms and a composite index of overall precariousness.

The aggregate analysis reduces dimensionality and yields interpretable summaries of how domains of precariousness relate to overall depression severity. However, it may obscure finer-grained effects. The disaggregated analysis allows for precise mapping of which symptoms are influenced by (or influence) specific precariousness conditions, but it introduces complexity due to the high dimensionality and distributional properties of the data. The mixed approach, in turn, integrates these perspectives by examining how individual symptoms respond to cumulative precariousness across domains.

To robustly estimate the causal structure between depression and precariousness variables, we implement a comprehensive bootstrapped causal discovery procedure. This procedure systematically varies key analysis parameters across a grid of configurations, enabling us to assess the stability of inferred relationships under different settings.
Specifically, we vary four main components:

* Significance levels for conditional independence tests: $\alpha$ = 0.01 and 0.05.
* Stability thresholds: 0.5, 0.6, 0.7, and 0.8, which determine how frequently an edge must appear across bootstraps to be retained.
* Conditional independence (CI) tests: Gaussian CI test (partial correlation) and RCoT (a nonparametric test).
* Causal discovery algorithms: FCI, CCI, and PC.

For each unique parameter combination (2 significance levels × 4 thresholds × 2 CI tests = 16 settings), we run 100 bootstrap samples per algorithm for the aggregated analyses (see Step 1 in @fig-workflow). For symptom-level analyses, which are more computationally intensive, we use 30 bootstrap samples per setting.[^5]

[^5]: For symptom-level analyses, we further improve computational efficiency by fixing the skeleton structure—the undirected network of potential connections—using a consensus graph derived from all three algorithms (with $\alpha = 0.01$ and the RCoT test). This allows subsequent analyses to focus exclusively on estimating edge directions, avoiding the repeated and costly computation of skeletons, and thus substantially reducing runtime.

To summarize these results, we apply a two-level aggregation strategy:

1. **Within each condition** (fixed $\alpha$, threshold, CI test, and algorithm), we identify the most frequently occurring edge type (e.g., arrowhead, circle, none) across bootstraps for each pair of variables (Step 2 in @fig-workflow).
2. **Across conditions**, we then determine the most dominant edge type based on its frequency. In cases where no edge type is clearly dominant (i.e., a tie), we mark the edge as uncertain, represented with a dashed line in the final graph (Step 3 in @fig-workflow).

This summarization process is repeated separately for each causal discovery algorithm (FCI, CCI, PC), yielding a single robust graph per algorithm. These final graphs represent only the most stable and consistent relationships across a wide range of plausible parameterizations, enhancing both reliability and interpretability.
To support transparency, each final graph is accompanied by a matrix summarizing the relative frequency of each edge type for every variable pair, enabling readers to assess the degree of uncertainty or agreement underlying each connection.

<!-- To ensure robust estimation and guard against sensitivity to parameter settings, we implemented a systematic grid of analysis conditions: -->

<!-- * Significance levels: $\alpha$ = 0.01 and 0.05 -->
<!-- * Bootstrap thresholds: 0.5, 0.6, 0.7, 0.8 -->
<!-- * CI tests: Gaussian and RCoT -->
<!-- * Algorithms: FCI, CCI, and PC -->

<!-- Each combination was bootstrapped 100 times for sum-score level analyses and 30 times for symptom-level analyses, yielding a total of 1,600 graphs per algorithm for the full analysis. For symptom networks, we fixed the skeleton structure using a consensus graph estimated across PC, FCI, and CCI ($\alpha$ = 0.01, RCoT). This constraint improves computational efficiency without limiting the estimation of directional edges between symptoms and precariousness variables. -->

<!-- To summarize the results, we adopted a two-level aggregation strategy. First, within each analysis condition, we recorded the most frequently observed edge-endpoint type (e.g., directed, undirected, uncertain) from the bootstraps. Then, across all conditions, we tallied the most frequent endpoint type per edge. In cases of ties, we assigned dashed edges to reflect uncertainty. For transparency, we accompany each graph with a table showing the relative frequencies of each edge symbol. -->

<!-- This procedure yields a final graph for each algorithm and analysis type—offering both robustness and interpretability. See @fig-workflow for an overview of the causal discovery pipeline. -->
<!-- For detailed explanations of the algorithms, edge interpretations, and graph types (e.g., PAG, MAAG, CPDAG), as well as algorithmic assumptions and limitations, see the [Appendix](#sec-appendix). -->

<!-- To uncover directional relationships between precariousness and depression, we used constraint-based causal discovery algorithms capable of detecting feedback loops and latent confounding. Specifically, we applied the *Fast Causal Inference* (FCI) and *Cyclic Causal Inference* (CCI) algorithms, both of which support cyclic and non-acyclic structures [@mooij2020constraint; @strobl2019]. Additionally, we include the *PC* algorithm as a reference, given its simplicity and prominence as one of the most widely known causal discovery methods [@spirtes2001causation]. -->

<!-- One practical challenge in applying these algorithms to the HELIUS dataset is that most variables do not follow a Gaussian distribution, and the relationships among variables are unlikely to be strictly linear. To address this, we supplement the commonly used Gaussian conditional independence (CI) test, which relies on partial correlations, with a non-parametric CI test based on reproducing kernel methods [@zhang2012kernel]. -->
<!-- While kernel-based conditional independence tests (KCIT) are theoretically effective in capturing non-linear dependencies and subtle interactions in data with non-Gaussian noise, they are computationally demanding for large datasets like HELIUS. Their quadratic scaling with sample size is primarily due to the need to invert large kernel matrices, making them impractical for large-scale analysis [@rahimi2007random]. -->
<!-- To mitigate this issue, we use the Randomized Conditional Correlation Test (RCoT), which approximates kernel methods using random Fourier features. This approach reduces computational complexity from quadratic to linear scaling with sample size, significantly lowering computational costs while maintaining robustness [@strobl2019approximate]. For a more detailed explanation of RCoT, refer to @sec-rcot. -->

<!-- We analyze the causal structure using three complementary approaches: (1) Examining the relationship between the five individual precariousness factors and the PHQ sum score, which represents overall depression severity. -->
<!-- (2) Investigating the relationship between individual depressive symptoms and individual precariousness factors. (3) Assessing the relationship between individual depressive symptoms and overall precariousness (the sum of all precariousness factors). -->
<!-- Using the PHQ sum score as a measure of depression simplifies the analysis by reducing dimensionality, improving computational efficiency, and providing a broad, interpretable perspective on how precariousness factors relate to overall depression severity. However, this aggregated approach may overlook granular symptom-level relationships. -->

<!-- By zooming in on individual symptoms, we capture a more detailed picture of how specific depressive symptoms interact with different precariousness factors. This approach allows us to map individual causal pathways, identifying which symptoms are particularly sensitive to certain precarious conditions and vice versa. However, this level of detail introduces methodological challenges, including non-standard distributions of symptom variables and the high-dimensional nature of the analysis, which can reduce statistical power and significantly increase computational demands—especially when using non-parametric conditional independence (CI) tests. -->

<!-- Finally, analyzing individual symptoms in relation to overall precariousness provides an integrative perspective, revealing how depression symptoms collectively relate to precarious conditions. This approach complements the fully disaggregated analysis by aggregating precariousness factors, potentially uncovering relationships that may be too weak to detect when examined separately. This is particularly relevant when a symptom is associated with multiple precariousness factors, but these associations are not strong enough to emerge in an individual-level analysis. -->
<!-- By integrating these three approaches — ranging from partially aggregated to fully disaggregated — we aim to achieve a balance between capturing detailed granularity and gaining a broader understanding of the relationship between precariousness and depression.  -->

<!-- Furthermore, to address sensitivity to parameter values and enhance the robustness of our results, we implement multiple settings combined with bootstrapping. For each condition, we generate 100 bootstrap samples, estimate causal graphs, and retain only the edge-endpoints that exceed predefined thresholds. If no edge-endpoint type surpasses the threshold, we assign a circle ($\circ$) to indicate uncertainty in the edge direction. -->

<!-- The analyses are conducted under the following conditions: -->

<!-- -   Significance levels ($\alpha$): 0.01 and 0.05 -->
<!-- -   Thresholds: 0.5, 0.6, 0.7, and 0.8 -->
<!-- -   CI test: Gaussian CI test, RCoT -->
<!-- -   Algorithms: FCI, CCI, and PC -->

<!-- This setup yields 16 combinations (2 significance levels × 4 thresholds × 2 CI tests), applied across three algorithms. Each combination is repeated for 100 bootstrap samples, yielding a total of 1,600 graphs per algorithm.  -->
<!-- For analyses involving individual symptom variables, we simplify the setup by focusing on thresholds of 0.6 and 0.7 and reducing the number of bootstrap samples to 30, easing computational demands. -->
<!-- To further enhance efficiency, we fix the skeleton of the symptom network using a common structure derived from PC, FCI, and CCI algorithms ($\alpha$ = 0.01, RCoT CI test). This step constrains which edges can be present among symptoms but does not restrict the estimation of causal directions within the symptom network. Additionally, the entire causal structure between symptoms and precariousness factors remains fully estimable, ensuring that key relationships are still inferred. -->
<!-- Beyond its computational benefits — particularly in reducing the time required for skeleton estimation — this fixed structure also aligns well with commonly reported symptom network structures in the literature [@park2025individual], ensuring consistency with existing findings. -->

<!-- To obtain a final, stable summary of causal relationships, we identify the most frequently occurring edge-endpoints across different experimental setups. First, within each condition, we determine the most frequent edge-endpoint from all bootstrap samples. Then, across different conditions, we aggregate these selected symbols and determine the overall most common edge-endpoint for each graph entry. -->

<!-- Since we have an even number of experimental setups, ties may occasionally occur. In such cases, we explicitly denote them in the resulting graph using dashed lines. To further help with the interpretation and understanding of the resulting graph, we provide a corresponding table for each graph, showing the proportion of each edge-endpoint type to clarify how dominant a given symbol is in each case. -->

<!-- By double summarizing the results in this way, we generate one final graph per algorithm for each analysis approach: the partially aggregated analysis, which uses either the sum of precariousness factors or the sum of depression factors to capture overall trends, and the fully disaggregated analysis, which examines individual precariousness and depression factors separately. -->
<!-- This approach ensures that only stable and consistent edges — those that persist across different parameter settings — are retained, leading to a robust and reliable understanding of the relationships between precariousness factors and depression. See @fig-workflow for an overview of the analysis workflow. -->

```{r}
#| label: fig-workflow
#| fig-cap: "Causal discovery workflow applied across all three algorithms (FCI, CCI, and PC). *The process involves three main steps: (1) bootstrapping the dataset to create multiple resampled datasets; (2) estimating causal graphs under varying parameter configurations; and (3) summarizing results by identifying the most frequent edge type across bootstraps and settings to construct a final stable graph.*"
#| fig-align: center
#| out-width: 90%

knitr::include_graphics("img/simsetup.pdf")

```

## Modeling Dynamics between precariousness and Depression

To explore how internal system structure shapes responses to external support, we implement a high-level computational model that formalizes key mechanisms suggested by the causal discovery analysis. Rather than attempting to replicate the full estimated causal graph, this model abstracts the essential structure into a compact, tractable system focused on dynamic interactions between depression and social precariousness, both influenced by financial stress.

As illustrated in [@fig-concept], the model includes three components: social precariousness $P_i(t)$, depressive symptoms $D_i(t)$, and a static individual-specific external stressor $S_i$, representing financial stress. Depression and precariousness are modeled as coupled state variables that influence each other over time, with $S_i$ acting as a shared input. These dynamics are formalized as a pair of linear stochastic differential equations (SDEs):

$$
\begin{aligned}
dP_i(t) &= \lambda_P \big( \alpha_{PS} S_i + \alpha_{PD} D_i(t) - P_i(t) \big) \, dt + \sigma_P \, dW_{P,i}(t) \\
dD_i(t) &= \lambda_D \big( \alpha_{DS} S_i + \alpha_{DP} P_i(t) - D_i(t) \big) \, dt + \sigma_D \, dW_{D,i}(t)
\end{aligned}
$$

Here, $D_i(t)$ denotes the depressive state and $P_i(t)$ denotes the level of social precariousness for individual $i$ at time $t$, with $S_i$ representing a static, individual-specific external stressor (financial stress). The parameters $\lambda_D$, $\lambda_P$, as well as the coefficients $\alpha$ and noise terms $\sigma$, are assumed to be constant across individuals and time.
The parameters $\lambda_D$ and $\lambda_P$ govern the timescales over which $D_i(t)$ and $P_i(t)$ adjust toward their respective input-driven target states. The coefficients $\alpha_{DS}$ and $\alpha_{PS}$ quantify the direct effects of the external stressor $S_i$ on depression and precariousness, respectively, while $\alpha_{DP}$ and $\alpha_{PD}$ capture the mutual influence between the two dynamic processes. Stochastic fluctuations are introduced via the independent Wiener processes $dW_{D,i}(t)$ and $dW_{P,i}(t)$, scaled by their respective volatility parameters $\sigma_D$ and $\sigma_P$.
This linear structure offers a transparent and analytically tractable framework for modeling how exogenous variation in financial stress may propagate through individual dynamics of depression and precariousness. Crucially, all parameters can be calibrated directly from empirical covariances and variances observed in the HELIUS cohort. The resulting steady-state distribution of the model aligns with empirical measures: PHQ-9 total score (PHQsum) for depression, and the derived social precariousness factor (P.soc) for social precarity. The full calibration procedure is described in [@sec-calibration].

<!-- Here, $D$ denotes depression, $P$ denotes social precariousness, and $S$ is the external stressor associated with financial stress. The parameters $\lambda_D$ and $\lambda_P$ control the timescales over which each variable adjusts toward its input-driven target state. The coefficients $\alpha_{PS}$ and $\alpha_{DS}$ quantify the direct effects of the stressor on $P$ and $D$, while $\alpha_{DP}$ and $\alpha_{PD}$ represent the cross-influences between depression and precariousness. Stochastic fluctuations are introduced through the noise terms $\sigma_P, dW_P$ and $\sigma_D, dW_D$, with $dW_P$ and $dW_D$ representing independent Wiener processes. -->
<!-- This linear structure offers a transparent and analytically tractable framework for examining how variations in external financial stress influence the population-level dynamics of depression and social precariousness. Crucially, the model's simplicity enables direct calibration from empirical data: all parameters can be derived analytically from observed variances and covariances in the HELIUS cohort. The following  [@sec-calibration] details the procedure used to perform this calibration. -->

```{r}
#| label: fig-concept
#| fig-cap: "High-level schematic of the dynamic model structure used in simulation. Arrows represent influence coefficients $\\alpha$, matching the terms in the system of stochastic differential equations. $D_i(t)$ and $P_i(t)$ denote depressive symptoms and social precariousness for individual $i$, with $S_i$ representing static financial stress."
#| fig-align: center
#| out-width: 40%

knitr::include_graphics("img/conceptmodel.pdf")
```

### Model Calibration {#sec-calibration}

In our linear model, all parameters are derived analytically to reproduce the empirical second-order moments observed in the HELIUS dataset. The system treats depression ($D$) and social precariousness ($P$) as jointly evolving under the influence of a shared external stressor ($S$), forming a coupled system of Ornstein–Uhlenbeck processes. Under the assumption of stationarity, the steady-state covariances between these variables admit closed-form solutions for the system’s structural parameters.

We fix the adjustment rates $\lambda_D = \lambda_P = 1$, which normalizes the time scale given the cross-sectional nature of the HELIUS data.
To resolve the remaining degree of freedom, we treat the feedback from precariousness to depression ($\alpha_{DP}$) as a free parameter, and we solve algebraically for all other coefficients — specifically the direct effects of stress ($\alpha_{DS}, \alpha_{PS}$), the reverse feedback from depression to precariousness ($\alpha_{PD}$), and the noise terms ($\sigma_D, \sigma_P$). These solutions are expressed as closed-form functions of the empirical covariances: $\mathrm{Var}(D)$, $\mathrm{Var}(P)$, $\mathrm{Var}(S)$, $\mathrm{Cov}(D, P)$, $\mathrm{Cov}(D, S)$, and $\mathrm{Cov}(P, S)$.

This structure creates a family of analytically calibrated models, each consistent with the same observed covariances but differing in their internal configuration. We constrain the space of valid solutions by imposing that all structural parameters must be positive, reflecting the assumption that stress and precariousness contribute adversely to depression. This constraint restricts the allowable range of $\alpha_{DP}$ to approximately $[0, 0.698]$ (see Appendix [@sec-analcalibration]).

@fig-tradeoff illustrates how the remaining parameters adjust as $\alpha_{DP}$ increases to maintain consistency with the empirical covariances. As $\alpha_{DP}$ increases, $\alpha_{PD}$ tends to decrease, reflecting a compensatory trade-off between the mutual influences of depression and precariousness. This interplay helps preserve the model’s match with observed covariances, without requiring major shifts in external input strength. Meanwhile, the external coupling terms — $\alpha_{DS}$ and $\alpha_{PS}$ — also shift to absorb variance in a way that maintains alignment with the empirical data. Notably, the noise amplitudes $\sigma_D$ and $\sigma_P$ decline slightly as $\alpha_{DP}$ increases. This indicates that more of the observed variability is captured by deterministic interactions, with less left to be explained by stochastic fluctuations.

Rather than being a limitation, this flexibility enables exploration of a spectrum of equally data-consistent feedback regimes. Each represents a different plausible causal architecture, offering insight into how internal dynamics modulate the system's responsiveness to external stress — even when the observed covariances remain unchanged.

```{r}
#| label: fig-tradeoff
#| fig-cap: >
#|   Trade-off curves showing how calibrated parameters vary with $\alpha_{DP}$.
#|   *Each line shows how a different parameter adjusts in response to changes in $\alpha_{DP}$ to maintain consistency with empirical covariances from the HELIUS data. 
#|   All configurations shown are equally consistent with the data, illustrating how internal dynamics adjust as feedback strength changes.*
#| fig-align: center
#| out-width: 100%
knitr::include_graphics("img/param_tradeoff.pdf")
```

### Intervention Simulation

To evaluate how internal structure affects system responsiveness, we simulate a family of linear models subjected to external intervention. Each model corresponds to a different value of $\alpha_{DP}$, representing the strength of influence from social precariousness to depression. The range spans from 0 (no coupling) to 0.65 (maximal admissible value), with all other parameters derived analytically to preserve the empirical covariances observed in the HELIUS data (see @fig-tradeoff).

In each simulation, financial stress of individual $i$ ($S_i$) is proportionally reduced toward the lowest observed value, modeling a population-wide intervention applied at varying intensities (0–100%). For each model, we simulate 1000 virtual individuals whose baseline stress levels are sampled from the empirical distribution and scaled according to the intervention level.

We assess the system’s behavior using two complementary approaches:

- Snapshot analysis captures the average levels of depression ($D$) and precariousness ($P$) at selected time points ($t = 100$, $300$, $500$) following intervention onset. Since all linear models converge to the same equilibrium, this highlights how internal feedback affects the speed and pathway of adjustment.

- Full trajectory simulation tracks system dynamics over time for two representative models: one with minimal ($\alpha_{DP} = 0$) and one with maximal ($\alpha_{DP} = 0.65$) feedback. Each simulation includes an intervention period followed by its withdrawal, revealing differences in reactivity, delay, and recovery stability.

Together, these simulations demonstrate how internal structure governs not just whether change occurs, but how it unfolds — shaping the pace, persistence, and variability of response to external support.

# Results

## Causal Structure Linking Precariousness and Depression

### Depression as sum score
@fig-sum provides a high-level summary of how precariousness factors collectively influence and are influenced by overall depression severity, focusing on the causal relationships between precariousness factors (*P.hou*, *P.emp*, *P.soc*, *S.rel*, *S.fin*) and the depression sum score (*PHQsum*), as identified by the FCI algorithm.
<!-- As described in [@sec-analysis], dashed edges indicate ties, and the edge endpoint at which the tie occurs is marked by a bold horizontal bar. -->

<!-- The causal graph in panel (a) highlights key pathways in the relationships between precariousness factors and depression (*PHQsum*).  -->
In panel (a) we see that employment precariousness (*P.emp*) and social precariousness (*P.soc*) are not identified as causes of depression, whereas financial stress (*S.fin*) appears to play a causal role. While *P.emp* and *P.soc* are generally not recognized as causes of other precariousness factors, *S.fin* emerges as a potential cause, as indicated by its circle edge endpoint. This pattern is further supported by the edge-type proportion matrix shown in panel (b), which aggregates edge inference outcomes across bootstrapped datasets.

<!-- The matrix plot can be interpreted such that the symbol in matrix[$i$, $j$] represents the relationship $i -~[symbol] \  j$. For example, if matrix[$i$, $j$] is > and matrix[j, i] is $\circ$, then the inferred relationship is $i \circarrow j$.  In the table, different edge types are represented by distinct colors: light gray for the absence of an edge, blue for circles, green for arrowheads, and coral for arrowtails. The colors are blended based on the proportion of each edge type, with higher proportions increasing the opacity of the corresponding color, making dominant symbols more visually prominent.  -->

Relational stress (*S.rel*) plays a more nuanced role, interacting with depression and *S.fin*  through a latent confounder. Its connection to *P.soc* is less definitive—the proportion matrix in @fig-sum-2 suggests that this relationship is frequently absent, ambiguous, or potentially bidirectional across bootstrap samples. Meanwhile, housing precariousness (*P.hou*) is not directly causally related to depression but is linked to *P.emp* and *S.fin*. While *P.emp* and *S.fin* are identified as non-causes of *P.hou*, it remains unclear whether *P.hou* causally influences *P.emp* or if their relationship is mediated by an unobserved confounder.

The findings from the CCI algorithm (see [@fig-sum-cci]) are largely consistent with those from the FCI analysis, with one key distinction: CCI shows greater ambiguity in the directional relationship between *S.fin* and *PHQsum*: CCI assigns an arrowhead from *S.fin* to *PHQsum* in roughly 50% of bootstraps, while the remaining cases display an undirected (circle) endpoint.
Overall, CCI introduces greater uncertainty in edge orientations relative to FCI, yielding more undirected endpoints and more frequent bidirectional (confounded) edges, which in turn contribute to a slightly higher overall count of arrowheads when directions are inferred. 
<!-- Despite this, the skeleton structure and overall causal directions derived from CCI align well with the results of the FCI algorithm. For example,  -->
Both algorithms identify directional pathways from the stressors (*S.rel*, *S.fin*) to *P.soc* and suggest that *P.soc* is not a cause of *PHQsum*, but more likely affected by it. *P.emp* is not inferred to cause any of its neighboring variables (*P.hou*, *PHQsum*, *S.fin*) and instead appears to be affected by them. *P.hou*, in turn, is connected to *P.emp* and *S.fin* with circle endpoints, indicating that it may be a potential source of influence rather than an outcome. 
<!-- These consistent pathways underscore the robustness of the core relational architecture across algorithms amid variation in directional confidence. : " This is discussion point : Vitor-->

<!-- In  graph  (a),  the  key  pathways  suggest  that  employment  precariousness  (P.emp) and social precariousness (P.soc) do not cause depression (PHQsum).P.soc does not cause recent relational stress (S.rel) or financial stress (S.fin), and these stressors are likely related by a latent confounder.   Additionally, P.emp and S.fin are identified as non-causes of housing precariousness (P.hou). Both FCI and CCI detect a dependency between P.emp and S.fin, but they disagree on the direction of the relationship, leaving it unclear whether S.fin causes P.emp or whether a latent confounder mediates their relationship. Similar ambiguities are found in the connections between S.rel and PHQsum and between S.fin and PHQsum -->

<!-- Both FCI and CCI detect a dependency between *P.emp* and *S.fin*, but they disagree on the direction of the relationship, leaving it unclear whether *S.fin* causes *P.emp* or whether a latent confounder mediates their relationship. Similar ambiguities are found in the connections between *S.rel* and *PHQsum* and between *S.fin* and *PHQsum*. -->

<!-- Graph (b), derived solely from the nonparametric RCoT method. Both graphs consistently identify that *P.emp* is not a cause of *P.hou* or *PHQsum*, and that *P.soc* does not cause *PHQsum* or *S.fin*.  -->
<!-- The relationship between stressors and depression remains unresolved between the two algorithms. However, the graph derived from RCoT provides greater confidence that depression is not the cause of *S.fin* and suggests a stronger likelihood that the financial stress may contribute to depression or that their relationship is mediated by a latent confounder. -->
<!-- The role of relational stress has become less pronounced, as the edge between *P.soc* and *S.rel* is omitted, and the relationship between *S.rel* and *S.fin* is now inconsistent between the algorithms. -->

<!-- Overall, both graphs together highlight the potential roles of stressors like *S.fin* and *S.rel*, which are closely tied to depression, either causally or via latent confounders. Employment and social precariousness may also be influenced by depression, either directly or through latent variables. Lastly, housing precariousness does not directly impact depression but remains connected to employment precariousness and financial stress, likely mediated by direct or latent pathways. -->

Causal relationships between five precariousness factors and depression sum score (PHQsum), as estimated using the FCI algorithm across bootstrapped samples.
(a) Aggregated causal graph showing the most frequent edge types across bootstraps. Dashed edges indicate ties, with a bold horizontal bar marking where the tie occurs.
(b) Proportion matrix summarizing edge types across all bootstraps. Each matrix cell $[i, j]$ represents how often an edge from variable $i$ to variable $j$ exhibited a particular edge mark—arrowhead, tail, circle, or no edge—adjacent to $j$. The edge mark in matrix$[i,j]$ reflects the orientation of the edge ending at $j$. For example, if matrix$[i, j]$ shows a green arrowhead ($>$) and matrix$[j, i]$ shows a blue circle ($\circ$), then the inferred relationship is $i \circarrow j$. Edge types are encoded by distinct colors: green for arrowheads, coral for arrowtails, blue for circles, and light gray for absence of an edge. Colors are blended and shaded by frequency, with greater opacity indicating higher confidence in the inferred edge type.

```{r}
#| layout-ncol: 2
#| label: fig-sum
#| fig-cap: "Causal relationships between five precariousness factors and depression sum score (*PHQsum*), as estimated using the FCI algorithm across bootstrapped samples. **(a)** Aggregated causal graph showing the most frequent edge types across bootstraps. Dashed edges indicate ties, with a bold horizontal bar marking where the tie occurs.  **(b)** Proportion matrix summarizing edge types across all bootstraps: each matrix cell $[i, j]$ indicates how often a given edge from variable $i$ to variable $j$ exhibited a particular edge mark (e.g., arrowhead, tail, or circle) adjacent to $j$. Green indicates arrowheads, coral arrowtails, blue circle marks, and gray absence of an edge. Colors are blended and shaded by frequency, with dominant symbols appearing more opaque. For example, matrix[$\\mathit{S.fin}$, $\\mathit{PHQsum}$] shows a green arrowhead ($>$) and matrix[$\\mathit{PHQsum}$, $\\mathit{S.fin}$] shows a coral arrowtail (\\textemdash), together indicating that $\\mathit{S.fin} \\rightarrow \\mathit{PHQsum}$ in the majority of bootstrapped graphs."
#| fig-subcap: 
#|   - "FCI PAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/FCI_depsum.png")
knitr::include_graphics("img/depsum_mat_fci.pdf")

```

<!-- \begin{figure}[htbp] -->
<!--     \centering -->
<!--         \caption{Summary of all sum score graphs} -->
<!--         \includegraphics[width=1\textwidth]{img/sum_all.png} -->
<!-- \end{figure} -->
<!-- \begin{figure}[htbp] -->
<!--     \centering -->
<!--         \caption{Summary of sum score graphs given RCoT} -->
<!--         \includegraphics[width=1\textwidth]{img/sum_rcotonly.png} -->
<!-- \end{figure} -->

### Individual depression symptom

Moving from the sum score representation to the disaggregated symptom-level graph provides a more granular perspective on the causal relationships between precariousness factors and depression symptoms. Unlike the sum score graph, which aggregates all symptoms into a single measure, the symptom-level graph in @fig-sym reveals the heterogeneity in how specific symptoms (*con* (concentration), *slp* (sleep), *ene* (energy), *app* (appetite), *mot* (motor), *sui* (suicidal), *anh* (anhedonia), *glt* (guilt), and *dep* (depressed mood)) are influenced by, and in turn influence, different forms of precariousness.

@fig-sym-1 reveals a strong interdependence among symptoms and suggests much presence of latent confounders, as indicated by numerous bidirectional edges. 
<!-- As before, dashed edges denote ties, with specific ties marked by bold horizontal bars.  -->
In @fig-sym-2, the proportion matrix indicates that arrowheads are the most frequently inferred edge type. However, circle endpoints are particularly common for *anh*, *slp*, *ene*, and *sui*, reflecting uncertainty in the causal direction among these symptoms. Notably, the only arrowtail connection appears between *dep* and *anh*, suggesting that anhedonia is a potential cause of depressed mood.

Looking at the symptom-precariousness connections, one of the key patterns is that ties are most frequently found in the relationships between individual symptoms and precariousness factors. This suggests that these causal links may be less stable across different conditions. Closer examination reveals that most ties arise due to discrepancies between the Gaussian CI test and the RCoT test—with Gaussian CI favoring arrowheads and RCoT more often predicting the absence of an edge.

<!-- This discrepancy likely arises from a combination of factors.  -->
<!-- As the analysis moves from aggregated to disaggregated variables, the number of CI tests increases substantially, leading to a loss of statistical power. With more variables being conditioned on, the effective sample size per test diminishes, making it harder to detect subtle dependencies. Over-conditioning—controlling for too many variables—can also obscure real relationships, resulting in false negatives. -->
<!-- Beyond these statistical power issues, the underlying assumptions of the two CI tests contribute to their divergence. RCoT is a nonparametric test that captures both linear and nonlinear dependencies without assuming a specific distributional form. However, it is more conservative and requires larger sample sizes, especially in high-dimensional settings. In contrast, the Gaussian CI test is based on partial correlations and assumes linear Gaussian relationships, making it more permissive and more likely to detect weak or spurious associations. -->
<!-- As a result, it is common to see edges where Gaussian CI suggests a directional relationship, while RCoT indicates no edge. In the proportion matrix, this appears as a 50–50 tie between absence and arrowhead symbols.  -->
<!-- This systematic divergence illustrates the broader challenge of causal discovery in high-dimensional data: different statistical assumptions can lead to different conclusions about whether—and in which direction—variables are causally related. -->

Despite these differences, some consistent patterns emerge across both CI tests, particularly in the case of *S.fin*, which appears to be connected to nearly all other variables—though many of these connections are marked by ties, reflecting uncertainty in directionality. One exception is the stronger evidence suggesting that *S.fin* may cause changes in *app*, while its relationships with other symptoms, such as *ene*, *mot*, and *dep*, remain ambiguous.
Similarly, *P.soc* exhibits numerous connections with depressive symptoms, with most edges pointing toward *P.soc* rather than outward from it. This suggests that depressive symptoms, particularly *dep* and *glt*, may contribute to worsening social precariousness rather than the other way around. Additionally, *anh* also shows signs of a potential causal influence on *P.soc*, reinforcing the idea that social precariousness is more often a consequence rather than a driver of depressive symptoms.

Other precariousness factors exhibit more uncertainty but still notable relationships. *S.rel* is connected to *slp*, *glt*, and *app*, though directionality remains unclear in many cases. *P.emp* also connects to symptoms like *slp*, *mot*, and *con*, but, like *S.rel*, these relationships exhibit a 50/50 split in directionality, reflecting uncertainty in the inferred causal paths.
In line with the aggregated graph in @fig-sum, *P.hou* does not appear to have any direct relationship with depression symptoms but consistently shows associations with *P.emp* and *S.fin*. The causal relationships among precariousness factors remain largely unchanged from the aggregated analysis, with *S.fin* exhibiting the strongest tendency to influence other precariousness factors.

Among depressive symptoms, *dep*, *glt*, and *slp* appear to be the most connected to precariousness factors, while *P.soc* has the most connections with symptoms, predominantly as a recipient rather than a driver of influence. Within the symptom network, *slp*, *sui*, and *anh* emerge as causally influential symptoms, as they exhibit more outgoing arrows compared to other symptoms. On the other hand, *con*, *mot*, *dep*, and *glt*, despite having high connectivity, predominantly receive incoming arrows, indicating they are more likely effects rather than causes.
Considering both symptom-precariousness connections and symptom-level dynamics, *slp* emerges as a central symptom, given its strong ties to precariousness factors and its influential role within the symptom network.
<!-- This suggests that sleep issues could be an initiating symptom, particularly sensitive to relational stress and employment precariousness.  -->
*P.soc* appears to function as a bridge between the depression and precariousness subsystems, as depressive symptoms appear to feed back into social precariousness, reinforcing a self-sustaining dynamic between depression and precarious conditions.

Finally, the results from CCI are largely consistent with those from FCI, but with some key differences. CCI tends to favor arrowheads more frequently, resulting in a greater number of bidirectional edges, which suggests a higher involvement of latent confounders. Additionally, CCI exhibits more tie situations, though, similar to FCI, most ties occur between absence and arrowhead edges, reflecting the discrepancy between the Gaussian CI test and RCoT—where the Gaussian test favors arrowheads, while RCoT more often suggests the absence of an edge. For a detailed visualization of the CCI-derived graph and proportion matrix, see @fig-sym-cci.

<!-- <!-- Centrality and Connectivity --> 
<!-- Certain nodes in the network emerge as more *causally* central, highlighting their potential importance as intervention points. Among the depressive symptoms, *anh*, *dep*, *slp*, and *glt* emerge as particularly influential, with multiple outgoing edges (o->) to other symptoms,  suggesting their roles as potential key drivers within the symptom network. Especially, symptoms such as *glt*, *slp*, and *anh* are particularly critical, potentially acting as initiator or activator nodes within the network due to their apparent connections with precariousness factors. These symptoms appear to be especially sensitive to external stressors, potentially manifesting early in response to such conditions and subsequently activating other interconnected symptom nodes. -->
<!-- <!-- This sensitivity indicates their potential as early warning signs or key targets for interventions aimed at preventing the progression to severe depression. --> 
<!-- The causal structure involving precariousness factors in the symptom-level graph is largely consistent with the patterns observed in the sum score graphs. As in the sum graphs, relational stress (*S.rel*) emerges as a potential causal factor for depression, while employment (*P.emp*) and social precariousness (*P.soc*) are more likely to be influenced by depressive symptoms.  -->

<!-- The symptom-level analysis, however, provides more specificity by pinpointing the symptoms involved in these relationships. For instance, *slp* is identified as influencing *P.emp* or potentially through a latent confounder (*slp* o-> *P.emp*) , while *glt* appears to affect *P.soc* or may be linked via a confounder (*glt* o-> *P.soc*). -->
<!-- Additional edges are observed in the graph generated using Gaussian CI tests, such as *slp* <-> *S.rel* and *slp* o-> *P.emp*. As seen in the sum graph, Gaussian CI testing tends to produce a denser graph. In this case, causal relationships involving *slp* are particularly prominent. -->

<!-- Some discrepancies, however, exist between the symptom-level and sum graphs. For example, financial stress (*S.fin*) does not have any edges with depressive symptoms in the symptom-level graph, although it maintains associations with other precariousness factors. Additionally, housing precariousness (*P.hou*) becomes entirely disconnected from the rest of the network, appearing as an isolated node.  -->

<!-- However, due to computational constraints that required limiting the sample size, as well as the challenging distributions of certain symptoms, the statistical power is lower compared to the depression sum score graph. -->
```{r}
#| label: fig-sym
#| layout-nrow: 2
#| fig-cap: "Resulting graph of precariousness factors and individual depression symptoms using FCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "FCI PAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 70%

knitr::include_graphics("img/symptom_graph_FCI.png")
knitr::include_graphics("img/symptom_mat_fci.pdf")

```

### Precariousness as sum score

Lastly, we examine the relationships between individual symptoms and overall precariousness, an aggregated measure represented by the sum of five precariousness factors. This analysis provides a complementary perspective, capturing broader patterns that may not be evident in the disaggregated symptom-precariousness analysis. By aggregating precariousness factors into a single score, this approach may capture distributed relationships across different precariousness factors that were previously overlooked in the disaggregated analysis.

Unlike the disaggregated graph in @fig-sym, the aggregated symptom-precariousness graph (@fig-presum-1) shows greater certainty in causal directions, with ties occurring only in edges involving the overall precariousness factor.
This is even more evident in the proportion matrix (@fig-presum-2), where symptom-to-symptom interactions almost fully converge to a single edge type. Additionally, all symptom interactions are connected either through bidirectional edges or a combination of an arrowhead and a circle, suggesting a strong presence of latent confounders.

Examining the symptom-precariousness connections, we observe a stronger overall trend of depressive symptoms influencing precariousness, rather than the other way around. Specifically, *anh*, *dep*, and *slp* frequently emerge as sources of influence on precariousness, whereas *app*, *glt*, and *mot* exhibit more uncertainty in directionality, often resulting in circle endpoints. This pattern suggests that when precariousness factors are aggregated, the dominant causal flow is from depressive symptoms to precariousness, rather than precariousness driving depression. Particularly, *dep* emerges as the strongest predictor of precariousness, exhibiting the highest proportion of arrowtails, consistent with findings from the disaggregated analysis.

Comparing this with the CCI-derived graph (@fig-presum-cci), several key differences stand out. The CCI results reveal that most symptoms are interconnected through bidirectional edges, suggesting a strong presence of latent confounders. However, *dep* appears to be a distinct exception, as it is predominantly caused by nearly all other symptoms, with one exception *sui*, which does not contribute to *dep*. Similarly, *con* is influenced by multiple symptoms, albeit with weaker support compared to *dep*.
Most interestingly, the CCI results highlight a distinct feedback loop between *dep* and overall precariousness, represented by a tail-tail edge (\textemdash \). This suggests a possible reinforcing cycle between depression and overall precariousness, where depression not only arises from precarious conditions but also contributes to their persistence, creating a self-sustaining dynamic.

Across multiple levels of analysis, our results point to a consistent and directional link between precariousness and depression. Financial stress (*S.fin*) emerges as a particularly robust and modifiable upstream driver: in both aggregated and disaggregated models, it shows strong and stable causal influence on depressive symptoms. By contrast, other forms of precariousness—particularly social precarity—more often appear as outcomes rather than antecedents of depression, frequently receiving rather than sending causal arrows. Among symptoms, *dep* and *slp* stand out for their centrality and connections to precariousness factors, with *dep* also implicated in potential feedback loops with social precarity (*P.soc*). These findings motivate the use of a simplified three-node model in the simulation—capturing depression, financial stress, and social precarity—as it preserves the key dynamic motifs observed in the causal graphs while enabling tractable exploration of system-level responses to stress-reducing interventions.

<!-- Overall, the aggregated precariousness graph supports the key findings from both the disaggregated analysis and the depression sum score analysis, providing greater certainty in several important patterns. The results highlight the significant role of latent confounders in symptom-to-symptom interactions. While certain stressor-related precariousness factors, particularly *S.fin* and *S.rel*, appear to causally influence depressive symptoms, other precariousness factors are more likely driven by depression rather than acting as primary causes, suggesting that precariousness may function more as a consequence of depression than its source. Some symptoms, particularly *glt*, *slp*, and *dep*, appear to influence precarious conditions while also occupying central roles within the symptom network. Among them, *slp* and *glt* emerge as influential symptoms within the depression network, whereas *con* and *dep*, despite their high connectivity, behave more as outcomes rather than primary drivers. Together, these findings suggest that depression—especially symptoms like sleep disturbances and guilt—may contribute to sustaining precarious conditions, possibly forming a reciprocal cycle between depression and precariousness. -->

```{r}
#| layout-nrow: 2
#| label: fig-presum
#| fig-cap: "Resulting graph of precariousness sum score and individual depression symptoms using FCI and proportion of edge endpoint types. As before, dashed edges denote ties, with specific ties marked by bold horizontal bars."
#| fig-subcap: 
#|   - "FCI PAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/presum_graph_FCI.png")
knitr::include_graphics("img/presum_mat_fci.pdf")
```

## Simulating the Dynamics Between Precariousness and Depression Under Intervention

To assess how internal structure influences the system’s responsiveness to external support, we simulate a set of linear models that are consistent with the data but vary in the assumed strength of influence from precariousness to depression ($\alpha_{DP}$).
For each model, we generate a virtual population of 1,000 individuals and simulate an external intervention that reduces each individual’s financial stress ($S$) proportionally toward the minimum observed value in the dataset. While individuals begin with different baseline stress levels, the intervention applies a shared proportional reduction across the population. This intervention defines the external input, while variation in $\alpha_{DP}$ reflects different assumptions about internal feedback from precariousness to depression.
We examine system behavior in two complementary ways: (1) by analyzing average levels of depression and precariousness at selected time points following intervention onset, and (2) by simulating full temporal trajectories under sustained intervention followed by recovery, using two representative cases at the extremes of the $\alpha_{DP}$ range.

@fig-intervention summarizes how variation in $\alpha_{DP}$ shapes system-level responses to external support.
<!-- Panel A shows the mean levels of depression ($D$) and precariousness ($P$) at time steps 100, 300, and 500, with the x-axis indicating the strength of external support, defined as the proportional reduction of $S$ toward its minimum observed value. -->
<!-- Each line represents a different scenario with a different $\alpha_{DP}$, color-coded from low (purple) to high (yellow). -->
In Panel A, during the early phase of intervention, depression and precariousness exhibit contrasting patterns. Depression declines more rapidly in scenarios with lower $\alpha_{DP}$, where its dynamics are driven more directly by the external stressor $S$. 
As $\alpha_{DP}$ increases, $D$ becomes more dependent on precariousness ($P$), which itself adjusts more slowly—delaying the depressive response.
Precariousness, by contrast, shows slightly faster early reductions in scenarios with higher $\alpha_{DP}$. This reflects a shift in influence: as $\alpha_{DP}$ increases, $P$ becomes more directly responsive to $S$ (via increasing $\alpha_{PS}$) and less influenced by $D$, allowing it to adjust more promptly to changes in the stressor.

Despite these short-term differences, all scenarios ultimately converge to similar steady-state levels under sustained intervention. AS the systm approaches convergence, some variation remains, particularly in $P$, where gradients between models are still visible. 
<!-- By step 500, trajectories have nearly fully aligned across all configurations, consistent with the shared equilibrium structure of the linear system.  -->
This convergence highlights that the key distinctions lie not in whether intervention is effective, but in how quickly and through which pathways change unfolds.

Panel B shows full time trajectories for two representative cases: one with minimal internal feedback ($\alpha_{DP} = 0$) and one at the upper admissible limit ($\alpha_{DP} = 0.65$). In both, there is a temporary intervention period. When $\alpha_{DP} = 0$, depression responds rapidly to the reduction in stress and rebounds shortly after support is withdrawn. In contrast, when $\alpha_{DP}$ is high, $D$ responds more gradually and recovers more slowly—reflecting its increased dependence on the slower-changing $P$.
Precariousness declines in both scenarios during intervention and returns to baseline afterward, but with a slightly sharper drop in the higher $\alpha_{DP}$ condition. 
This occurs because $P$ becomes more directly responsive to $S$, as consistency with the data requires $\alpha_{PS}$ to increase with $\alpha_{DP}$ (see @fig-tradeoff). These parameter shifts reflect trade-offs needed to preserve the empirical covariances, with further details provided in Appendix @sec-analcalibration.

These asymmetries further highlight how internal structure—not just external inputs—shapes the system’s trajectory. Specifically, changes in coupling parameters reconfigure how stress reduction propagates through the system, altering the speed and pathway of response even when long-run outcomes remain unchanged.

```{r}
#| label: fig-intervention
#| fig-cap: >
#|   System response to external intervention across variations in internal structure.
#|   ***Panel A**: Mean levels of depression ($D$) and precariousness ($P$) at time steps 100, 300, and 500, across varying values of $\alpha_{DP}$ and intervention strengths. The x-axis represents the strength of external support, operationalized as the proportional reduction of financial stress ($S$) toward its minimum observed value. Each line corresponds to a different model defined by a specific $\alpha_{DP}$ value, color-coded from low (purple) to high (yellow). Models with lower $\alpha_{DP}$ exhibit faster early reductions in $D$, while higher $\alpha_{DP}$ leads to more immediate declines in $P$.
#|   
#|   **Panel B**: Full time trajectories for two models—one with no coupling ($\alpha_{DP} = 0$) and one with strong coupling ($\alpha_{DP} = 0.65$)—illustrating system response and recovery following intervention onset (step 300) and withdrawal (step 1500).
#|   Depression responds more gradually and persistently in the high $\alpha_{DP}$ model, whereas precariousness adjusts more quickly.*

#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/intervention_linear.pdf")
```

# Discussion

This study combined constraint-based causal discovery and dynamical modeling to investigate how depression and socioeconomic precariousness interact and how their internal structure modulates responsiveness to intervention.
Consistent with prior work linking financial strain to psychological distress [@lund2010poverty; @ezzy1993unemployment], our causal analysis identified financial stress as a key upstream driver of depression—emerging across both aggregated and disaggregated models. This provides empirical support for longstanding hypotheses that economic stressors can act as modifiable entry points into broader cycles of psychosocial risk.

At the symptom level, our analysis aligns with established theories of psychopathology but offers a more fine-grained map of potential causal relationships, generating concrete, testable hypotheses. It also points to possible limitations of prevailing models, as some observed symptom–symptom links may be shaped by shared confounders rather than direct causal influence. This suggests that, in some contexts, broader mechanisms or clusters of symptoms may provide more effective units of analysis and intervention than individual symptoms alone. 
Within this framework, sleep disturbance emerged as a potentially initiating symptom, with directed connections to both symptom-level and precariousness variables—mirroring longitudinal findings that identify sleep disruption as an early warning signal for depressive episodes and social withdrawal [@zawadzki2013rumination; @alvaro2013systematic].
Depressed mood, another core symptom, occupied a central position in the network. Not only was it influenced by upstream symptoms like anhedonia and guilt, but it also showed directional influence on social precariousness—suggesting that internal psychological states may actively shape external vulnerability. The CCI algorithm further identified a potential feedback loop between depressed mood and overall precariousness, resonating with social drift theories in depression, which posit that declining mental health can lead to social exclusion, job instability, and housing insecurity, reinforcing depressive cycles [@compton2015social; @boschloo2015network].

In contrast, social precariousness itself appeared more often as an effect than a cause, receiving arrows from symptoms like depressed mood and guilt but rarely sending them. This asymmetry supports prior observations that indicators of social vulnerability often reflect downstream consequences of poor mental health, though the relationships are likely bidirectional over time [@lund2010poverty]. 
These patterns, while derived from Amsterdam-specific data, align with mental health research in other high-income city settings, where structural inequalities such as poverty, unemployment, and housing insecurity are persistent drivers of poor psychological outcomes [@fryers2003social; @weich1998poverty].
Although the specific dynamics may vary by context, the parallels with broader literature suggest that similar mechanisms could operate in comparable settings. Nonetheless, replication in other populations will be important for assessing robustness and external validity.

Taken together, these findings yield actionable insights for mental health intervention. Symptoms like sleep disturbance, which demonstrate early and broad influence across both symptom and precariousness networks, may serve as effective targets for prevention—potentially halting the progression into more entrenched depressive states or worsening precariousness. Similarly, intervening on central nodes like depressed mood may offer strategic leverage for breaking mutually reinforcing cycles of mental decline and social isolation.
Importantly, interventions targeting such nodes may have ripple effects across domains. For instance, evidence suggests that treating core depressive symptoms can improve social functioning, employment engagement, and housing stability—particularly when combined with social or financial supports [@mcdaid2019economic; @lund2018social]. These dual benefits underscore the value of integrated, cross-domain strategies that simultaneously address internal symptoms and external stressors.

From a systems perspective, our results also highlight why single-point interventions may fail to produce lasting change. The broader causal structure reveals multiple interlocking feedback loops spanning both psychological symptoms and structural risk factors, which may re-amplify vulnerability even after temporary relief. Supporting this, recent simulation work by @park2025role shows that feedback loops, even when modest in strength, can substantially diminish the long-term impact of isolated interventions unless multiple reinforcing links are simultaneously disrupted.
Our own simulations support this view: while all models converged to similar long-term outcomes under sustained support, differences in internal structure—specifically, the strength of the link from precariousness to depression—produced systematic variation in how quickly and through which pathways systems responded to intervention. Models with stronger internal coupling exhibited slower recovery dynamics and greater persistence of depression following stress reduction, suggesting that internal system architecture, not just external input, shapes the effectiveness of support.
This supports growing calls for system-aware intervention design [@borsboom2017network]. Rather than assuming uniform responses, more durable outcomes may require (1) targeting early-acting symptoms, (2) sustaining support where recovery is slow, and (3) combining structural and psychological interventions. Tailoring strategies to match the system’s architecture can improve effectiveness, particularly in populations with layered vulnerabilities.
In this light, understanding causal architecture is not merely theoretical—it is foundational to designing more effective and equitable interventions. By identifying key modifiable drivers like financial stress and influential symptoms like sleep disturbance, and by modeling how system dynamics unfold under intervention, our study offers a flexible framework for improving mental health outcomes in complex social environments.

<!-- Limitations -->
While the FCI and CCI algorithms are designed to recover causal structure from cross-sectional data—including feedback and latent confounding—many edge directions in our results remain ambiguous, particularly in links between individual symptoms and precariousness factors. These inconsistencies likely reflect limited statistical signal in high-dimensional settings, where large conditioning sets reduce power and small effect sizes are harder to detect. 
Moreover, our findings showed sensitivity to the choice of conditional independence (CI) test. The Gaussian CI test, which assumes linearity and Gaussian noise, often produced denser graphs, likely due to overestimating weak or noisy linear associations. In contrast, RCoT, a nonparametric alternative, avoids these assumptions but relies on a Gaussian radial basis function (RBF) kernel to assess dependence. While flexible in capturing nonlinear relationships, this kernel is optimized for smooth, continuous variables and may underperform with discrete or ordinal data. Its similarity estimates can be unreliable when applied to variables with discontinuities, potentially leading to missed dependencies. This is particularly relevant in the HELIUS dataset, where some variables are ordinal or mixed-type [@howlett2001radial]. To address this, future work should consider adapting kernel-based CI tests to better handle mixed data structures—for example, by employing hybrid kernels that explicitly combine continuous and discrete similarity measures. This could improve the reliability of nonparametric causal discovery in settings with heterogeneous variable types.

Although the CCI algorithm is explicitly designed to detect cycles, aside from the connection between precarity and depression in the aggregate precarity setting, we found little clear evidence of feedback: most candidate loops appeared only as bidirectional edges, which may reflect latent confounding or statistical imprecision. This could point to either an absence of strong feedback at the symptom and the precariousness factor levels or limitations in the sensitivity of full-graph discovery approaches applied to densely interconnected variables. More focused strategies, such as local structure tests or theory-guided subgraph modeling, may help resolve these ambiguities. 
Moreover, our analyses relied solely on observational data from the Amsterdam population. Some structural uncertainties, particularly those involving latent confounding or bidirectional influence, may ultimately require interventional or semi-interventional designs. Extensions such as LLC [@hyttinen2012learning], NODAGS-Flow [@sethuraman2023nodags], or Bicycle [@rohbeck2024], which integrate observational and interventional data, could be especially valuable where such data are available.

Finally, the structure of our dynamical model introduces important limitations. The model focused on a restricted set of variables (depression, social precariousness, and financial stress). This design was intentionally simplified to isolate core mechanisms of interest, but it necessarily omits other relevant domains of precariousness, additional stressors, and potential unintended consequences of intervention. In addition, while the linear formulation offered analytical tractability and transparent calibration, it likely oversimplifies the true dynamics of depression and socioeconomic precariousness. A growing body of evidence suggests that psychosocial systems are fundamentally nonlinear. Empirical and theoretical work points to phenomena such as threshold effects, where improvements only occur after a tipping point is crossed [@van2014critical; @park2025individual]; saturation, where the impact of support diminishes as basic needs are met [@cramer2010comorbidity]; and path dependence, where past adversity shapes sensitivity to future stressors [@wichers2016critical].
These behaviors are fundamentally incompatible with linear dynamics, yet they may be essential for understanding why some individuals remain trapped in high-risk states while others recover more easily. Future modeling efforts should therefore incorporate mechanisms such as saturating input-response functions (e.g., sigmoidal form) and state-dependent feedback (where coupling strength varies by system state).
Incorporating such features would allow models to capture not just average effects but also the critical conditions under which meaningful change becomes possible—clarifying whether some subgroups require forceful, threshold-crossing interventions, while others may benefit from sustained, moderate support.

<!-- Finally, the structure of the dynamical model introduces its own limitations. While the model was designed to reflect key features of the discovered causal structure, it represents a highly simplified system: only two variables—depression and social precariousness—modulated by external financial stress. This abstraction allowed for tractable analysis and visual intuition, particularly around nonlinear behaviors such as bistability, but omits other potentially influential constructs highlighted in the causal graphs, including relational stress, employment conditions and housing conditions. Moreover, the model’s nonlinearity—introduced via the hyperbolic tangent function—was chosen for convenience rather than empirical justification. While this function supports saturation and the emergence of tipping points, the presence of bistability is a product of simulation, not empirical verification. It remains possible that simpler, linear dynamics could account for the observed data just as well. Future work should further investigate whether bistability is a necessary feature of depression–precariousness dynamics, and explore alternative functional forms—both linear and nonlinear—using richer data and stronger empirical constraints. -->

<!-- Implications -->

As public health research has long shown, social and psychological systems do not always respond passively to external inputs; internal constraints can limit the impact of even well-intentioned policies [@wagenaar2013public; @desavigny2009systems]. By using causal discovery and computational modeling, researchers and policymakers can better identify these structural constraints and tailor interventions accordingly.
More broadly, this study offers a generalizable framework for investigating complex mental health dynamics. Causal discovery supports principled hypothesis generation from observational data, while dynamical modeling enables exploration of how hypothesized structures behave under external change. Together, these tools move us beyond static descriptions toward mechanistic insight with direct relevance for real-world mental health policy and intervention design.
Future research should build on this foundation by developing richer, more nuanced dynamical models grounded in longitudinal and interventional data. Incorporating time, multiple domains of precarity, and adaptive feedback processes would help assess whether the mechanisms proposed manifest in actual mental health trajectories. Ultimately, we hope this study encourages further integration of causal structure learning and systems modeling in efforts to understand and  improve mental health outcomes under conditions of socioeconomic precariousness.

# References

::: {#refs}
:::

# Appendix {#sec-appendix}

## HELIUS study {#sec-helius}

The HELIUS (HEalthy LIfe in an Urban Setting) study is a large-scale, multiethnic cohort study conducted in Amsterdam. Participants were randomly selected from the municipality register and stratified by ethnic origin to ensure balanced representation across six major groups. Invitation letters were sent by mail, followed by a reminder after two weeks. Non-respondents were contacted via home visits where applicable.

Of those invited, approximately 55% responded (Dutch: 55%, Surinamese: 62%, Ghanaian: 57%, Turkish: 46%, Moroccan: 48%). Among those contacted, 50% agreed to participate (Dutch: 60%, Surinamese: 51%, Ghanaian: 61%, Turkish: 41%, Moroccan: 43%), resulting in an overall participation rate of 28%.

Participants completed either a digital or paper version of the questionnaire, with assistance offered when needed, and received a confirmation letter for a physical examination appointment.

At baseline (2011–2015), 24,780 individuals were enrolled. Of these, 23,936 participants completed the questionnaire that included items relevant to precariousness. After excluding individuals with missing data on key variables, the final analytical sample consisted of 21,628 participants.

The HELIUS study received approval from the Medical Ethics Committee of the Academic Medical Center (AMC), and written informed consent was obtained from all participants prior to enrollment.

## Causal Discovery Primer {#sec-causalprimer}

As shown in Table 1, the resulting graphs from FCI and CCI differ slightly (*PAG*: partial ancestral graph; *MAAG*: maximal almost ancestral graph) due to their reliance on different underlying assumptions. 
Despite these differences, both graphs belong to the class of *ancestral graphs*, which are designed to encode causal relationships between variables, where the presence of an edge indicates causal *ancestry.*
In these graphs, directed edges, $A \stararrow B$, indicate that $B$ is not an ancestor of $A$ in every graph within the Markov equivalence class, $Equiv(G)$. 
The Markov equivalence class represents a set of graphs that encode the same conditional independence relationships, ensuring that the same *d-separation* conditions hold across all graphs in the class [@spirtes2001causation].
Conversely, an edge marked as $A \startail B$ indicates that $B$ is an ancestor of $A$ across all graphs in $Equiv(G)$. 
Circle endpoints, $A \starcirc B$, represent ambiguity in the ancestral relationship, meaning $B$'s ancestral status relative to $A$ varies across graphs in $Equiv(G)$.[^4] 
Finally, when an edge is represented as $A \arrowarrow B$, it implies that neither $A$ nor $B$ is an ancestor of the other, suggesting the presence of a latent confounder influencing both variables.

\begin{table}[ht]
\centering
\small % Make the font size smaller
\caption{Assumptions of causal discovery algorithms}
\begin{tabular}{lccccc}
\toprule
Algorithm & Acyclicity & \makecell{Causal \\ sufficiency} & 
\makecell{Absence of \\ selection bias} & Linearity & Output \\ 
\midrule
PC  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & CPDAG \\ 
FCI & $-^{a}$ & $\checkmark$ & $-^{a}$ & $-^{a}$ & PAG \\ 
CCI & x & x & x & $\checkmark$ & \makecell{\footnotesize(partially oriented) \\ MAAG} \\ 
\bottomrule
\end{tabular}
\caption*{\footnotesize{\textit{Note}. $^{a}$The FCI algorithm, introduced by Spirtes (1995), is a constraint-based causal discovery method for DAGs that accounts for latent confounding and selection bias. Mooij \& Claassen (2020) later showed its applicability to cyclic causal discovery with latent confounding under general faithfulness and Markov conditions, assuming non-linear causal relationships.}}
\end{table}

<!-- For word docx compiling -->
<!-- **Table 1**: Assumptions of causal discovery algorithms. -->

<!-- | Algorithm | Acyclicity | Causal sufficiency | Absence of selection bias | Linearity | Output | -->
<!-- |-----------|------------|--------------------|---------------------------|-----------|--------| -->
<!-- | PC        | ✓          | ✓                  | ✓                         | ✓         |  CPDAG | -->
<!-- | FCI       | -¹         | ✓                  | -¹                        | -¹        |  PAG   | -->
<!-- | CCI       | x          | x                  | x                         | ✓         |  MAAG  | -->
<!-- ------------------------------------------------------------------------------------------------ -->
<!-- ¹FCI can detect cycles in systems that lack selection bias and exhibit non-linear relationships (Mooij & Claassen, 2020). -->

The algorithms also differ in how they detect cycles and represent cyclic relationships in their resulting graphs.
In CCI's MAAG, $A$ \textemdash\ $B$ indicates that $A$ is an ancestor of $B$, and simultaneously $B$ is an ancestor of $A$, referring to a cyclic relationship between $A \stackedarrows B$. 
FCI, on the other hand, identifies potential cycles more subtly. In its PAG, fully-connected nodes with circle endpoints ($\circirc$) may suggest the presence of cyclic structures.
FCI, therefore, provides a sufficient condition to distinguish variables that are not part of a cycle, offering a more nuanced approach to handling cyclic relationships [@mooij2020constraint].

Table 1 highlights further differences in the assumptions underlying these algorithms. CCI operates under the assumption of a linear system, whereas FCI, particularly when used to infer cyclic relationships, assumes a non-linear system without selection bias and adheres to more general faithfulness and Markov conditions (i.e., *$\sigma$-separation* and *$\sigma$-faithfulness* setting) [@forre2018constraint].
When the respective assumptions of each algorithm are met, their inferred cyclic relationships align with those illustrated in @fig-examplegraphs.

While CCI's MAAG provides an explicit representation of cyclic relationships, it has certain theoretical limitations, as the MAAG may not always fully preserve d-separation relations from the original graph, $\mathcal{G}$ [@strobl2019]. Given these considerations, we examine and compare results from both algorithms, placing more weight on the FCI results, which are presented in the main results section, while discussing CCI results where relevant, with detailed findings provided in @sec-cci.
A full discussion of causal discovery concepts and algorithmic details is beyond the scope of this paper; however, readers seeking a more in-depth understanding can refer to @park2024discovering for a comprehensive exploration of these methods and their applications.

[^4]: $*$ serves as a \textit{meta-symbol}, representing one of the three possible edge-endpoints. For instance, $A \tailstar B$ can indicate any of the following edges: $A$ \textemdash\ $B$, $A\tailarrow B$, or $A \tailcirc B$ [@park2024discovering].

```{r}
#| layout-ncol: 3
#| label: fig-examplegraphs
#| fig-cap: Example graph $\mathcal{G}$ featuring a cycle and the corresponding MAAG from CCI and PAG from FCI.
#| fig-subcap: 
#|   - Example graph $\mathcal{G}$
#|   - Corresponding MAAG 
#|   - Corrsponding PAG
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/original_DCG.png")
knitr::include_graphics("img/FCI_graph.png")
knitr::include_graphics("img/CCI_graph.png")
```

The graph produced by the PC algorithm is a *CPDAG* (completed partially directed acyclic graph), where directed edges ($A$ → $B$) indicate that $A$ is a direct cause (parent) of $B$. Unlike FCI and CCI, the CPDAG does not include circle symbols. Instead, when the PC algorithm cannot determine the direction of causality, it represents this uncertainty with bidirectional arrows. 
While the PC algorithm serves as a useful reference, its strict assumptions — acyclicity and the absence of latent confounders—limit its applicability in more complex scenarios. For this reason, our primary focus remains on the results obtained from FCI and CCI, with all PC algorithm results provided in @sec-pc for completeness.

## Precariousness Domains Based on Elsenburg et al. (2025)

1.  EMPLOYMENT PRECARIOUSNESS

-   `H1_Arbeidsparticipatie`: Working status
-   `H1_WerkSit`: Which work situation most applies to you?
-   `H1_RecentErv8`: Experiences past 12 months: h. You were sacked from your job or became unemployed (*reverse*)

2.  FINANCIAL PRECARIOUSNESS

-   `H1_InkHhMoeite`: During the past year, did you have problems managing your household income?
-   `H1_RecentErv9`: Experiences past 12 months: i. You had a major financial crisis (*reverse*)

3.  HOUSING PRECARIOUSNESS

-   `veilig_2012`: Score safety (veiligheid) in 2012 (*reverse*)
-   `vrz_2012`: Score level of resources (niveau voorzieningen) in 2012 (*reverse*)
-   `P_HUURWON`: Percentage Huurwoningen

4.  CULTURAL PRECARIOUSNESS

-   `H1_Discr_sumscore`: Perceived discrimination: sum score of 9 items (range 9-45)
-   `H1_SBSQ_meanscore`: Health literacy: SBSQ meanscore (range 1-5) (*reverse*)
-   `A_BED_RU`: Aantal bedrijfsvestigingen; cultuur, recreatie, overige diensten (*reverse*)

5.  SOCIAL PRECARIOUSNESS

-   `H1_RecentErv5`: Experiences past 12 months: e. Your steady relationship ended (*reverse*)
-   `H1_RecentErv6`: Experiences past 12 months: f. A long-term friendship with a good friend or family member was broken off (*reverse*)
-   `H1_RecentErv7`: Experiences past 12 months: g. You had a serious problem with a good friend or family member, or neighbour (*reverse*)
-   `H1_SSQT`: SSQT (frequency of social contact): sum score of 5 items (range 5-20) (*reverse*)
-   `H1_SSQSa`: SSQS (adequacy of social contact): sum score of 5 items, category 3 and 4 not combined (range 5-20) (*reverse*)

```{r, echo=FALSE, fig.align='center', fig.width=15}
# check correlation
#cor(df) |> round(2)

# visualize corr
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(cor(scaled_data[,c(1:16, 26, 27)]), method="color", col=col(200),
         type="upper", order="hclust",
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", number.cex=0.50, tl.srt=45,tl.cex=0.7, #Text label color and rotation
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
         )
```

-   **High** Correlations: `emp_stat` (employment status) and `work_sit` (work situation) have a strong positive correlation of 0.82. This suggests that individuals with higher employment status tend to have more secure or favorable work situations. `soc_freq` (social contact frequency) shows a strong positive correlation with `soc_adq` (social adequacy) at 0.61. This indicates that individuals with more frequent social contact also tend to have higher perceived adequacy of social interactions.

-   **Moderate** Correlations: `nb_safe` (neighborhood safety) and `nb_res` (resources) have a moderate positive correlation of 0.39, suggesting that areas with higher safety also have better resources. `hea_lit` (health literacy) has moderate correlations with `emp_stat` (0.26) and `work_sit` (0.25), which could mean that higher health literacy is associated with better employment situations. `frd_brk12` (friendship breakups) and `conf12` (conflicts) have a notable correlation of 0.43, indicating a relationship between having conflicts and friendship losses.

-   **Low to Moderate** Correlations in Financial Precariousness: `inc_dif` (income difficulties) has a moderate correlation with `fincri12` (financial crisis) at 0.49. This aligns with the expected relationship, where individuals who experience general income difficulties are more likely to report financial crises.

-   **Low** Correlations (0.1 - 0.2): Many variables, such as `discrim` (discrimination), `unemp12` (unemployment experience), and `rel_end12` (relationship end), have low correlations with other variables, suggesting relatively independent relationships in the context of this dataset.

### Exploratory Factor Analysis (EFA)

```{r, echo=FALSE, fig.align='center', fig.height=5, out.width="55%", message=FALSE, results='hide'}
# precarious factors
df <- scaled_data[,1:16]

# check suggested number of factors
pl <- fa.parallel(df, fa = "fa", n.iter = 100, show.legend = FALSE)
# run EFA
efa_result <- fa(df, nfactors = 5, rotate = "promax") # oblique rotation
#print(efa_result)
```

```{r, echo=FALSE, fig.align='center', out.width="70%"}
# Create a data frame from the pattern matrix
loadings <- data.frame(
  Variable = c("emp_stat", "work_sit", "unemp12", "inc_dif", "fincri12", "nb_safe", "nb_res", "nb_rent",
               "discrim", "hea_lit", "cul_rec", "rel_end12", "frd_brk12", "conf12", "soc_freq", "soc_adq"),
  MR1 = c(0.98, 0.97, 0.17, 0.21, 0.11, -0.08, 0.02, -0.03, -0.07, 0.27, -0.03, -0.04, -0.09, -0.08, -0.02, -0.03),
  MR2 = c(-0.06, -0.08, -0.03, 0.09, 0.05, -0.08, -0.05, -0.06, 0.17, 0.08, -0.08, -0.04, -0.04, 0.00, 1.00, 0.77),
  MR3 = c(-0.01, -0.06, 0.26, 0.28, 0.47, -0.05, -0.04, -0.02, 0.20, -0.10, -0.05, 0.36, 0.63, 0.56, -0.14, -0.01),
  MR4 = c(-0.02, -0.03, -0.04, 0.06, 0.03, -0.02, 0.71, -0.09, 0.13, 0.04, 0.61, -0.03, -0.01, -0.02, -0.04, -0.12),
  MR5 = c(-0.14, -0.11, -0.03, 0.16, 0.09, 0.61, -0.21, 0.54, 0.10, 0.15, 0.04, -0.01, -0.06, -0.06, -0.10, -0.13)
)

# Convert to long format
loadings_long <- loadings %>%
  pivot_longer(cols = starts_with("MR"), names_to = "Factor", values_to = "Loading")

# Plot heatmap
ggplot(loadings_long, aes(x = Factor, y = Variable, fill = Loading)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, limits = c(-1, 1)) +
  theme_minimal() +
  labs(title = "Factor Loadings Heatmap", x = "", y = "")
```

#### Factor Loadings (Pattern Matrix)

-   **MR1**: High loadings on `emp_stat` and `work_sit` suggest this factor captures *employment* precariousness.
-   **MR2**: Strong loadings on `soc_freq` and `soc_adq` indicate *social* precariousness.
-   **MR3**: Key items like `frd_brk12`, `conf12`, and `fincri12`, suggest recent *stressful events*.
-   **MR4**: High loadings on `nb_res` and `cul_rec` may reflect *community resources* precariousness.
-   **MR5**: Variables `nb_safe` and `nb_rent` with high loadings indicate *housing* precariousness.

#### Variance Explained

The factors cumulatively explain **38%** of the variance, with MR1 being the most influential factor. Each factor contributes a smaller proportion to the total variance (MR1 at 12%, MR2 at 9%, etc.).

#### Factor Intercorrelations

Factors are moderately correlated, especially between *MR1 and MR5*, and *MR2 and MR3*. This indicates that while distinct, these factors are related—reasonable in a complex socio-economic context.

#### Model Fit Statistics

RMSEA (0.071) suggest an acceptable fit. Tucker Lewis Index (0.802) suggests moderate reliability for the model.

#### Summary

The 5-factor model appears interpretable and captures distinct dimensions of precariousness: *employment, social, stressors, community resources, and housing precariousness*. Although the overall fit and explained variance could be stronger, these factors offer insights into the underlying structure of the data, highlighting key areas of precariousness.

### PCA

```{r, echo=FALSE, fig.align='center', out.width="70%"}
# run PCA
res.pca <- prcomp(df, scale = TRUE)
# print(res.pca)
#summary(res.pca)

# Visualize eigenvalues (scree plot)
fviz_eig(res.pca, addlabels = TRUE)
#
# fviz_pca_var(res.pca,
#              col.var = "contrib", # Color by contributions to the PC
#              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#              repel = TRUE     # Avoid text overlapping
#              )
```

-   Component Retention: The scree plot shows a clear "elbow" after the first component. This steep drop suggests that most variance is explained by the first component. After Dimension 5, the percentage of explained variance decreases slightly more gradually, indicating diminishing returns for adding more components. If we need to choose multiple components, retaining the first 5 components seems reasonable, as they capture most of the variance (cumulatively explaining about 54.7% of the total variance).

```{r, echo=FALSE, fig.align='center', fig.height=4.5, fig.width = 8}
# Results for Variables
res.var <- get_pca_var(res.pca)
# res.var$coord |> round(2)         # Coordinates
# res.var$contrib |> round(2)        # Contributions to the PCs
# res.var$cos2           # Quality of representation

# corrplot(res.var$cos2, is.corr=FALSE)

res.var$contrib |> as.data.frame() |> dplyr::select(1:5) |>
  tibble::rownames_to_column(var = "var") |> # adds a column for rownames
    pivot_longer(cols = starts_with("Dim"),
               names_to = "Dimension",
               values_to = "Contribution") |>
# Plot with facet_wrap to show each dimension in a separate facet
ggplot(aes(x = reorder(var, -Contribution), y = Contribution, fill = Contribution)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ Dimension, scales="free_y") +
  coord_flip() +
  labs(title = "Contribution of Variables to 5 Principal Components",
       x = "", y = "Contribution (%)") +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```

#### Explained variance (contributions) of variables

It shows the importance of variables within each component.

-   **Dim1**: High contributions are observed from `emp_stat`, `work_sit`, `inc_dif`, and `fincri12`, suggesting that this dimension captures aspects of *employment and financial* security.

-   **Dim2**: While `emp_stat` and `work_sit` overlap with Dim1, the strong contributions from `frd_brk12` and `rel_end12` indicate that this dimension captures a focus on *recent relationship stressors*.

-   **Dim3**: `cul_rec`, `nb_res` have the highest contributions, indicating this dimension likely represents *community and cultural* factors.

-   **Dim4**: `soc_freq` and `soc_adq` stand out in this dimension, suggesting an emphasis on *social* precariousness.

-   **Dim5**: `nb_safe` and `nb_rent` are the top contributors, pointing to *housing* security as key themes in this component.

#### Cos² Values

Cos² (squared cosine) values, or the quality of representation, show how well each variable is represented by each dimension. where higher cos² values (closer to 1) indicate better representation of a variable by a component.

```{r, echo=FALSE, fig.align='center', out.width="70%"}
res.var$cos2 |> as.data.frame() |> dplyr::select(1:5) %>%
  mutate(var = rownames(.)) |>
  pivot_longer(!var, names_to = "dim", values_to = "cos2") |>
  ggplot(aes(x = dim, y = var, fill = cos2)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue") +
  labs(title = "Cos2 of Variables on Each Component (1-5)", fill = "Cos2", x = "", y = "") +
  theme_minimal()
```

-   **Dim.1**: Variables `emp_stat`, `work_sit`, `inc_dif`, and `fincri12` show high cos² values, meaning that PC1 primarily captures variations in employment and financial difficulties. This component could represent *employment & finance* precariousness.

-   **Dim.2**: Variables `work_sit`, `emp_stat`, `frd_brk12`, and `conf12` are well-represented in this component, suggesting PC2 captures aspects of *recent relationship stressors*.

-   **Dim.3**: Variables `nb_res` and `cul_rec` load strongly on PC3. This may represent community or cultural resources, indicating that this component is associated with *neighborhood resources*.

-   **Dim.4**: This component has high cos² values for `nb_res`, `cul_rec`, `soc_freq`, and `soc_adq.` While `nb_res` and `cul_rec` are also prominent in PC3, PC4 uniquely captures nuanced differentiation in *social* precariousness.

-   **Dim.5**: `nb_safe` and `nb_rent` are well-represented by PC5. This component might capture *housing* precariousness.

### ICA

```{r, echo=FALSE,  fig.align='center', out.width="70%"}
# Load the library
library(fastICA)
# set the seed
set.seed(1)

# Set the number of components you want to extract
n_components <- 5

# Run ICA
ica_model <- fastICA(df,
  n.comp = n_components, alg.typ = "parallel",
  fun = "exp", method = "C", maxit = 5000
)
# glimpse(ica_model)

weight_matrix <- data.frame(t(ica_model$A))
names(weight_matrix) <- stringr::str_glue("IC{seq(weight_matrix)}")

# dist_matrix <- dist(as.matrix(weight_matrix))
# clust <- hclust(dist_matrix)
# plot(clust)

weight_matrix |>
  mutate(var = colnames(df)) |>
  pivot_longer(
    cols = starts_with("IC"),
    names_to = "IC", values_to = "loading"
  ) |>
  ggplot(aes(x = IC, y = var, fill = loading)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, limits = c(-1, 1)) +
  labs(x="", y="") +
  theme_minimal()
```

#### Dominant Variables per Component:

For each Independent Component (IC), we can identify variables with *high absolute* values in each column. These values indicate that the IC captures a strong, independent signal associated with these variables.

-   **IC1**: `soc_freq` and `soc_adq` have strong negative loadings on this component, indicating that this component might represent *social* precariousness.

-   **IC2**: `frd_brk12`, `conf12`, `rel_end12`, `fincri12` and `unemp12` have the most substantial loadings on this component, all with negative signs. This might point to a *recent relational or social stressor* component.

-   **IC3**: `nb_res` and `cul_rec` show notable negative loadings, pointing to a focus on *community resource* precariousness.

-   **IC4**: High loadings for `nb_safe`, `nb_rent`, `nb_res`, `cul_rec`, and `discrim` suggest a theme of *housing and community-based* precariousness, reflecting both safety and social challenges within the neighborhood context.

-   **IC5**: `emp_stat` and `work_sit` both have strong negative loadings on this component, suggesting it captures *employment* precariousness.

### Hierarchical clustering

#### Using Euclidean distance

-   Ward.D's method: Minimizes the variance within clusters, producing more compact and spherical clusters.
-   Single linkage: Groups clusters based on the minimum distance between points.
-   Complete linkage: Groups clusters based on the maximum distance between points.
-   Average linkage: Uses the average distance between all pairs of points in the two clusters.

```{r, warning=FALSE, message=FALSE, fig.height=7, echo=FALSE, fig.align='center'}
#| layout-ncol: 2
library(dendextend)  # For coloring the clusters
library(ggdendro)

# Compute the distance matrix between variables
distance_matrix <- dist(t(df))

# wardD
hc_ward <- hclust(distance_matrix, method = "ward.D")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)  # Specify the number of clusters
plot(hc_ward, main = "ward.D")
rect.hclust(hc_ward, k = 5, border = 2:6)

# Avg
hc_avg <- hclust(distance_matrix, method = "average")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)
plot(hc_avg, main = "average")
rect.hclust(hc_avg, k = 5, border = 2:6)

hc_comp <- hclust(distance_matrix, method = "complete")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)
plot(hc_comp, main = "complete")
rect.hclust(hc_comp, k = 5, border = 2:6)

hc_sing <- hclust(distance_matrix, method = "single")
# dend <- as.dendrogram(hc)
# dend_colored <- dendextend::color_branches(dend, k = 5)  # Specify the number of clusters
plot(hc_sing, main = "single")
rect.hclust(hc_sing, k = 5, border = 2:6)
```

##### Consistent Groupings (Across All or Most Methods)

-   `emp_stat` and `work_sit`: This pair consistently clusters together across all linkage methods, suggesting that they are closely related variables, likely capturing a similar aspect of the data (possibly employment status or employment-related information).

-   `nb_safe`, `nb_res`, and `nb_rent`: These variables are often grouped closely in several methods (especially Ward.D, average, and complete linkage). This suggests a similarity or common theme among them, potentially related to neighborhood or housing precariousness.

-   `soc_freq` and `soc_adq`: These two variables frequently cluster together, indicating they likely measure aspects of social frequency and adequacy in similar ways. They appear together in Ward.D, average, and complete linkage.

-   `frd_brk12` and `conf12`: These variables are often clustered closely (though they sometimes join with other variables like rel_end12), suggesting they may capture aspects of relationship or social conflict. This pair appears in close proximity, especially in average and Ward.D.

##### Inconsistent Groupings (Variability Across Methods)

-   `hea_lit`: This variable shows inconsistent clustering across methods. In Ward.D, it joins with `fincri12`, while in other methods, it’s often more isolated or grouped with variables that do not appear similar. This may suggest that `hea_lit` does not strongly correlate with other variables, or it has multidimensional aspects affecting its grouping across methods.

-   `discrim`: This variable also shows variable groupings. In Ward.D, it is grouped with `hea_lit`, while in other methods (e.g., complete and single linkage), it clusters differently, sometimes on its own. This variability may indicate that `discrim` has weaker associations with the main clusters in the data or overlaps partially with multiple clusters.

-   Social and Financial Variables (`inc_dif`, `fincri12`, `unemp12`): These variables appear together in some methods (e.g., Ward.D clusters `fincri12` and `inc_dif`), but in others, they are spread out. This inconsistency suggests that social and financial variables may not have strong or consistent ties across different methods, perhaps due to capturing different aspects of precariousness.

##### Summary

The consistent clusters are likely capturing distinct thematic dimensions of the data (e.g., employment, housing, social contact), while the inconsistent variables may reflect multifaceted or weakly correlated attributes that do not fit neatly into one cluster.

#### Using Mutual Information

Using mutual information (MI) as a basis for hierarchical clustering differs from using traditional distance measures (like Euclidean distance) in a few key ways.

```{r echo=FALSE, fig.align='center', fig.height=6, out.width="60%"}
library(infotheo)  # for mutual information functions

# Calculate pairwise mutual information
mi_matrix <- matrix(NA, ncol = ncol(df), nrow = ncol(df))
colnames(mi_matrix) <- rownames(mi_matrix) <- colnames(df)

for (i in 1:ncol(df)) {
  for (j in i:ncol(df)) {
    # Calculate mutual information between pairs of variables
    mi_matrix[i, j] <- mutinformation(discretize(df[, i]), discretize(df[, j]))
    mi_matrix[j, i] <- mi_matrix[i, j]  # symmetric
  }
}

# Normalize mutual information matrix
max_mi <- max(mi_matrix, na.rm = TRUE)
mi_distance <- 1 - (mi_matrix / max_mi)

# Convert to distance object
dist_matrix_mi <- as.dist(mi_distance)

# Hierarchical clustering on MI-based distance
hc_mi <- hclust(dist_matrix_mi, method = "ward.D2")

# Plot the dendrogram
plot(hc_mi, main = "Hierarchical Clustering (Mutual Information Distance)", xlab = "", sub = "")
rect.hclust(hc_mi, k = 5, border = 2:6)

```

##### Comparison to Euclidean Distance Clustering

-   **Housing and Community** Cluster: The variables `nb_safe`, `nb_res`, `nb_rent`, and `cul_rec` cluster together, indicating a strong association among housing-related and community-based factors. This suggests a shared theme of housing or community precariousness. This grouping is also observed in the Euclidean-based clustering, but it appears more tightly connected here, potentially due to the non-linear relationships highlighted by mutual information.

-   **Employment and Social Support** Cluster: `emp_stat` and `work_sit` form a cluster, linking employment status and work situation together as they did in Euclidean-based clustering. These remain closely associated regardless of the distance metric used. `soc_freq` and `soc_adq`, related to social contact frequency and adequacy, cluster nearby, indicating they have a stronger non-linear relationship with employment variables. This is a subtle difference as Euclidean distance might not capture this association as effectively.

-   **Financial Stressor** Cluster: `inc_dif` and `fincri12`, representing income difficulties and recent financial crises, consistently cluster together in both approaches, showing a strong association, likely linear. However, mutual information-based clustering links these financial stressors with social support variables, suggesting that financial challenges may have complex dependencies with social support in this dataset.

-   **Relational Stressor** Cluster: `frd_brk12`, `conf12`, `discrim`, `hea_lit`, `unemp12`, and `rel_end12` form a *looser* cluster focused on social and relational stressors (e.g., friendship breakup, conflicts, and discrimination). Compared to Euclidean clustering, `discrim` and `hea_lit` (health literacy) appear closer to relational stressors here, indicating that non-linear relationships might play a larger role in linking these variables.

##### Summary

In conclusion, mutual information-based clustering provides an alternative perspective that can reveal more intricate associations between variables, especially for those with non-linear relationships. Compared to Euclidean clustering, it shows a similar high-level structure but emphasizes nuanced connections between variables, particularly around social support, employment, and financial stress.

### Conclusions on Precariousness factors

Based on the consistent findings across multiple analyses, we decided to exclude the variables `discrim`, `hea_lit`, `umemp12`, and `rel_end12`, as they do not clearly belong to any specific precariousness factor nor exhibit strong associations with depression (see the correlation table above). Therefore, we propose retaining the following key precariousness factors:

-   Employment Precariousness: `emp_stat`, `work_sit`
-   Social Precariousness: `soc_freq`, `soc_adq`
-   Housing Precariousness: `nb_safe`, `nb_res`, `nb_rent`, `cul_rec`
-   Recent Relational Stressors: `frd_brk12`, `conf12`
-   Recent Financial Stressors: `fincri12`, `inc_diff`

We construct each precariousness factor by calculating the mean value of the combined variables. Below, we present the updated correlation table for the newly composed factors, along with the corresponding distributions of all variables to be used in the causal discovery analysis.

```{r, echo=FALSE, fig.align='center'}
clust_data <- scaled_data |>
  mutate(
    P.emp = rowMeans(cbind(emp_stat, work_sit)),
    P.soc = rowMeans(cbind(soc_freq, soc_adq)),
    P.hou = rowMeans(cbind(nb_safe, nb_res, nb_rent, cul_rec)),
    S.rel = rowMeans(cbind(frd_brk12, conf12)),
    S.fin = rowMeans(cbind(fincri12, inc_dif))
  ) |>
  dplyr::select(17:33, -c(age, ethn)) |>
  as.data.frame()

# skimr::skim(clust_data)

# visualize corr
corrplot(cor(clust_data), method="color", col=col(200),
         type="upper", order="hclust",
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", number.cex=0.50, tl.srt=45,tl.cex=0.7,
         diag=FALSE
         )

```

```{r echo=FALSE, fig.align='center', fig.height=4.5, fig.width=8}
# Reshape the data to long format
clust_data_long <- clust_data %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot distributions with density lines
ggplot(clust_data_long, aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)), fill = "lightblue", color = "white", bins = 15) +  # Use density for histogram
  geom_density(adjust=2, linewidth=0.3, color = "gray", alpha = 0.5) +  # Add density line in red
  facet_wrap(~ variable, scales = "free") +
  labs(title = "Distribution of All Variables with Density Overlay", x = "", y = "") +
  theme_minimal()
```

## Results from CCI algorithm {#sec-cci}

```{r}
#| layout-ncol: 2
#| label: fig-sum-cci
#| fig-cap: "Resulting graph of precariousness factors and depression sum score using CCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "CCI MAAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/CCI_depsum.png")
knitr::include_graphics("img/depsum_mat_cci.pdf")

```

```{r}
#| label: fig-sym-cci
#| layout-nrow: 2
#| fig-cap: "Resulting graph of precariousness factors and individual depression symptoms using CCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "CCI MAAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/symptom_graph_CCI.png")
knitr::include_graphics("img/symptom_mat_cci.pdf")

```

```{r}
#| layout-nrow: 2
#| label: fig-presum-cci
#| fig-cap: "Resulting graph of precariousness sum score and individual depression symptoms using CCI and proportion of edge endpoint types."
#| fig-subcap: 
#|   - "CCI MAAG"
#|   - "Proportion matrix"
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/presum_graph_CCI.png")
knitr::include_graphics("img/presum_mat_cci.pdf")
```

\clearpage

## Results from PC algorithm {#sec-pc}

```{r}
#| layout-ncol: 2
#| label: fig-pc_sum
#| fig-cap: "Resulting graphs of precariousness factors and depression sum score using PC."
#| fig-subcap:
#|   - "Using both GaussianCI and RCoT"
#|   - "Using only RCoT"
#| fig-align: center
#| out-width: 100%

knitr::include_graphics("img/sum_PC_all.png")
knitr::include_graphics("img/sum_PC_RCoTonly.png")
```

```{r}
#| label: fig-pc_sym
#| fig-cap: "Resulting graphs of precariousness factors and individual depression symptoms using PC."
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/PC_symptom.png")
```

```{r}
#| label: fig-pc_presum
#| fig-cap: "Resulting graphs of precariousness sum score and individual depression symptoms using PC."
#| fig-align: center
#| out-width: 60%

knitr::include_graphics("img/PC_presum.png")
```

\clearpage

## Randomized Conditional Independence / Correlation Test (RCIT & RCoT) {#sec-rcot}

RCIT (Randomized Conditional Independence Test) and RCoT (Randomized conditional Correlation Test) are advanced methods for scalable conditional independence (CI) testing, offering computational efficiency while maintaining the accuracy of kernel-based approaches. These methods evaluate conditional independence between two variables $X$ and $Y$ given a third variable $Z$ while addressing computational challenges inherent in kernel-based CI tests.  In this section, we provide a high-level overview of RCIT and RCoT based on [@strobl2019approximate].

### Kernel-Based Conditional Independence Testing

Traditional kernel-based CI tests, such as the Kernel Conditional Independence Test (KCIT), compute dependencies using the Hilbert-Schmidt Independence Criterion (HSIC) in reproducing kernel Hilbert spaces (RKHS) [@zhang2012kernel]. KCIT uses the following hypothesis framework:
$$
H_0: X \perp\!\!\!\perp Y \,|\, Z, \quad H_1: X \not\!\perp\!\!\!\perp Y \,|\, Z.
$$
The core quantity in KCIT is the partial cross-covariance operator:
$$
\Sigma_{XY \cdot Z} = \Sigma_{XY} - \Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY},
$$
where $\Sigma_{XY}$ represents the cross-covariance operator between $X$ and $Y$, and $\Sigma_{XZ} \Sigma_{ZZ}^{-1} \Sigma_{ZY}$ removes the dependence mediated by $Z$.

The squared Hilbert-Schmidt (HS) norm of $\Sigma_{XY \cdot Z}$ serves as the test statistic:
$$
\|\Sigma_{XY \cdot Z}\|^2_{HS} = 0 \quad \text{if and only if} \quad X \perp\!\!\!\perp Y \,|\, Z.
$$

KCIT estimates residual dependencies using kernel ridge regression:
$$
f^*(z) = K_Z (K_Z + \lambda I)^{-1} f(x),
$$
where $K_Z$ is the kernel matrix for $Z$, $f(x)$ is the kernel feature map for $X$, and $\lambda$ is the ridge regularization parameter. The residual function for $X$ is:
$$
f_\text{res}(x) = f(x) - f^*(z) = R_Z f(x),
$$
with:
$$
R_Z = I - K_Z (K_Z + \lambda I)^{-1}.
$$
The kernel matrix for residualized $X$ is:
$$
K_{X \cdot Z} = R_Z K_X R_Z,
$$
and similarly for $Y$, $K_{Y \cdot Z} = R_Z K_Y R_Z$.

The test statistic is computed as:
$$
T_{XY \cdot Z} = \frac{1}{n^2} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
$$
which estimates the Hilbert-Schmidt (HS) norm of the partial cross-covariance operator. To ensure convergence, KCIT scales the statistic by $n$:
$$
S_K = n T_{XY \cdot Z}.
$$
The null hypothesis $H_0$ is rejected if $S_K$ exceeds a threshold determined by permutation or moment-matching-based null distribution [@lindsay2000moment].

### Random Fourier Features (RFFs)

Kernel-based methods like KCIT face scalability issues, as they involve operations on $n \times n$ kernel matrices, which scale quadratically with the sample size $n$. RCIT and RCoT overcome this bottleneck using *Random Fourier Features (RFFs)* to approximate kernel operations efficiently.

#### Bochner’s Theorem
Bochner’s theorem provides the foundation for RFFs, stating that any continuous shift-invariant kernel $k(x, y)$ can be expressed as:
$$
k(x, y) = \int_{\mathbb{R}^p} e^{i \omega^\top (x - y)} \, dP_\omega,
$$
where $P_\omega$ is the spectral distribution of the kernel. For the widely used RBF kernel:
$$
k(x, y) = \exp\left(-\frac{\|x - y\|^2}{2\sigma^2}\right),
$$
$P_\omega$ follows a Gaussian distribution: $\omega \sim \mathcal{N}(0, \sigma^2 I)$.

#### RFF Approximation
Using Monte Carlo sampling, the kernel function is approximated as:
$$
k(x, y) \approx \phi(x)^\top \phi(y),
$$
where $\phi(x)$ is the random Fourier feature mapping:
$$
\phi(x) = \sqrt{\frac{2}{D}} \cos(W^\top x + b),
$$
with $W \sim \mathcal{N}(0, \sigma^2 I)$ and $b \sim \text{Uniform}(0, 2\pi)$. Here, $D$ is the number of Fourier features, which balances computational efficiency and approximation accuracy.

### Differences Between RCIT and RCoT

RCIT and RCoT differ in their test statistics, computational efficiency, and practical performance, which makes them suited for different scenarios in causal discovery. RCIT evaluates the Hilbert-Schmidt norm of the full partial cross-covariance operator, providing a general test for conditional independence but at a higher computational cost. RCoT simplifies the process by using the Frobenius norm of a finite-dimensional residualized cross-covariance matrix, significantly reducing complexity and improving scalability.

These distinctions are particularly important for large-scale datasets, where RCoT's computational efficiency makes it a practical choice for high-dimensional causal discovery tasks.

#### RCIT: Randomized Conditional Independence Test

RCIT tests full conditional independence by examining the squared Hilbert-Schmidt (HS) norm of the partial cross-covariance operator $\Sigma_{XY \cdot Z}$:
$$
S_K = n T_{XY \cdot Z} = \frac{1}{n} \text{tr}(K_{X \cdot Z} K_{Y \cdot Z}),
$$
where $T_{XY \cdot Z}$ is an empirical estimate of $\|\Sigma_{XY \cdot Z}\|^2_{HS}$. The null and alternative hypotheses are:
$$
H_0: \|\Sigma_{XY \cdot Z}\|^2_{HS} = 0, \quad H_1: \|\Sigma_{XY \cdot Z}\|^2_{HS} > 0.
$$

RCIT is a general test for conditional independence but becomes computationally demanding as the size of $Z$ increases, due to the high-dimensional kernel operations required.

#### RCoT: Randomized Conditional Correlation Test

RCoT simplifies the testing process by using a finite-dimensional partial cross-covariance matrix, avoiding full HS norm calculations. Instead, it uses the Frobenius norm of the residualized cross-covariance matrix:
$$
S' = n \|C_{AB \cdot C}\|_F^2,
$$
where $C_{AB \cdot C}$ represents the residualized cross-covariance matrix. The hypotheses are:
$$
H_0: \|C_{AB \cdot C}\|_F^2 = 0, \quad H_1: \|C_{AB \cdot C}\|_F^2 > 0.
$$

RCoT is computationally efficient and well-suited for large conditioning sets ($|Z| \geq 4$). Its simplicity enables robust calibration of the null distribution and improved scalability for high-dimensional data.

<!-- ## Analytical Calibration of the Linear Stochastic Model {#sec-analcalibration} -->

<!-- This appendix describes the derivation and calibration of parameters for a linear stochastic model linking depression ($D$), social precariousness ($P$), and an external stressor ($S$). The model is based on coupled Ornstein–Uhlenbeck (OU) processes and is analytically calibrated using empirical covariances from the HELIUS dataset. -->

<!-- ### Model Structure -->

<!-- The model assumes the following system of linear stochastic differential equations: -->

<!-- $$ -->
<!-- dD = \lambda_D \left( \alpha_{DS} S + \alpha_{DP} P - D \right) dt + \sigma_D\, dW_D  -->
<!-- $$ -->
<!-- $$ -->
<!-- dP = \lambda_P \left( \alpha_{PS} S + \alpha_{PD} D - P \right) dt + \sigma_P\, dW_P -->
<!-- $$ -->

<!-- where: -->

<!--  - $D$ = depression -->
<!--  - $P$ = social precariousness -->
<!--  - $S$ = exogenous financial stressor (fixed or sampled from empirical distribution) -->
<!--  - $\lambda_D$, $\lambda_P$ = adjustment rates (set to 1) -->
<!--  - $\alpha_{DS}, \alpha_{DP}, \alpha_{PS}, \alpha_{PD}$ = coupling coefficients -->
<!--  - $\sigma_D, \sigma_P$ = noise amplitudes -->
<!--  - $dW_D$, $dW_P$ = independent Wiener processes -->

<!-- ### Stationary Covariance Solution -->

<!-- At stationarity, the joint distribution of $D$ and $P$ is Gaussian. The system can be written in matrix form: -->

<!-- $$ -->
<!-- \frac{d\mathbf{x}}{dt} = A \mathbf{x} + B \boldsymbol{\xi}(t) -->
<!-- \quad \text{where } \mathbf{x} = \begin{bmatrix} D \\ P \end{bmatrix} -->
<!-- $$ -->

<!-- The drift matrix $A$ and diffusion matrix $BB^\top$ are: -->

<!-- $$ -->
<!-- A = -->
<!-- \begin{bmatrix} -->
<!-- -\lambda_D & \lambda_D \alpha_{DP} \\ -->
<!-- \lambda_P \alpha_{PD} & -\lambda_P -->
<!-- \end{bmatrix}, \quad -->
<!-- BB^\top = -->
<!-- \begin{bmatrix} -->
<!-- \sigma_D^2 & 0 \\ -->
<!-- 0 & \sigma_P^2 -->
<!-- \end{bmatrix} -->
<!-- $$ -->

<!-- Let $\Sigma_{DP}$ be the stationary $2 \times 2$ covariance matrix of $(D, P)$: -->

<!-- $$ -->
<!-- \Sigma_{DP} = -->
<!-- \begin{bmatrix} -->
<!-- v_D & c_{DP} \\ -->
<!-- c_{DP} & v_P -->
<!-- \end{bmatrix} -->
<!-- $$ -->

<!-- Then, the Lyapunov equation is: -->

<!-- $$A \Sigma_{DP} + \Sigma_{DP} A^\top + BB^\top = 0$$ -->

<!-- This provides a system of three equations in three unknowns: $v_D$, $v_P$, and $c_{DP}$. -->

<!-- ### Covariances with the Stressor -->

<!-- To complete the full $3 \times 3$ covariance matrix of $(D, P, S)$, we define: -->

<!-- $$ -->
<!-- \Delta = 1 - \alpha_{PD} \alpha_{DP} \\ -->
<!-- k_D = \frac{\alpha_{DP} \alpha_{PS} + \alpha_{DS}}{\Delta} \\ -->
<!-- k_P = \frac{\alpha_{PD} \alpha_{DS} + \alpha_{PS}}{\Delta} -->
<!-- $$ -->

<!-- Assuming $\mathrm{Var}(S) = v_S$, the unconditional covariances are: -->

<!-- $$ -->
<!-- \mathrm{Cov}(D, S) = k_D v_S  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Cov}(P, S) = k_P v_S  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Cov}(D, P) = k_D k_P v_S + c_{DP}  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Var}(D) = k_D^2 v_S + v_D  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Var}(P) = k_P^2 v_S + v_P -->
<!-- $$ -->

<!-- ### Empirical Calibration Procedure -->

<!-- Empirical estimates from HELIUS provide: -->

<!-- - $\mathrm{Var}(D)$, $\mathrm{Var}(P)$, $\mathrm{Var}(S)$ -->
<!-- - $\mathrm{Cov}(D, S)$, $\mathrm{Cov}(P, S)$, $\mathrm{Cov}(D, P)$ -->

<!-- Fix $\lambda_D = \lambda_P = 1$ and treat $\alpha_{DP}$ as a free parameter. Then solve for the five unknowns: -->
<!-- $\alpha_{DS}, \alpha_{PS}, \alpha_{PD}, \sigma_D^2, \sigma_P^2$ -->
<!-- symbolically using the observed moments. -->

<!-- For example: -->
<!-- $\alpha_{DS} = \frac{\mathrm{Cov}(D, S) - \alpha_{PS} \cdot \mathrm{Cov}(P, S)}{v_S}$ -->

<!-- The solutions are expressed as closed-form functions of empirical covariances and the chosen $\alpha_{DP}$ value. -->

<!-- ### Parameter Functions -->

<!-- The following functional forms are derived in the implementation: -->

<!-- $$ -->
<!-- \sigma_D = f_{\sigma_D}(\mathrm{Cov}(D, S), \mathrm{Cov}(D, P), \mathrm{Cov}(P, S), \alpha_{DP}) -->
<!-- $$ -->
<!-- $$ -->
<!-- \sigma_P = f_{\sigma_P}(\mathrm{Cov}(D, S), \mathrm{Cov}(D, P), \mathrm{Cov}(P, S), \alpha_{DP})  -->
<!-- $$ -->
<!-- $$ -->
<!-- \alpha_{DS} = f_{\alpha_{DS}}(\mathrm{Cov}(D, S), \mathrm{Cov}(D, P), \mathrm{Cov}(P, S), \alpha_{DP})  -->
<!-- $$ -->
<!-- $$ -->
<!-- \alpha_{PS} = f_{\alpha_{PS}}(\mathrm{Cov}(D, S), \mathrm{Cov}(D, P), \mathrm{Cov}(P, S), \alpha_{DP})  -->
<!-- $$ -->
<!-- $$ -->
<!-- \alpha_{PD} = f_{\alpha_{PD}}(\mathrm{Cov}(D, S), \mathrm{Cov}(D, P), \mathrm{Cov}(P, S), \alpha_{DP}) -->
<!-- $$ -->

<!-- The empirical values used in our study are: -->

<!-- $$ -->
<!-- \mathrm{Var}(D) = 1.000  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Var}(P) = 0.807  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Var}(S) = 0.743 -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Cov}(D, P) = 0.374 -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Cov}(D, S) = 0.357  -->
<!-- $$ -->
<!-- $$ -->
<!-- \mathrm{Cov}(P, S) = 0.256 -->
<!-- $$ -->

<!-- ### Summary -->

<!-- This calibration process enables exact, closed-form derivation of all model parameters directly from empirical covariance data. It avoids the need for iterative optimization and allows immediate simulation of the linear system under empirical constraints. -->

## Analytical Calibration of the Linear Feedback Model{#sec-analcalibration}

We consider a continuous-time, linear stochastic differential equation (SDE) model describing the coupled dynamics between depression ($D$) and social precariousness ($P$), each influenced by an external stressor ($S$). The system is specified as:

$$
\begin{aligned}
\frac{dD}{dt} &= \alpha_{DS} S + \alpha_{DP} P - D + \sigma_D \cdot \eta_D(t), \\
\frac{dP}{dt} &= \alpha_{PS} S + \alpha_{PD} D - P + \sigma_P \cdot \eta_P(t),
\end{aligned}
$$

where:

* $S$ is a fixed or exogenous input (external financial stress),
* $\alpha_{DP}$ represents the feedback from precariousness to depression,
* $\alpha_{PD}$ represents the feedback from depression to precariousness,
* $\alpha_{DS}$ and $\alpha_{PS}$ capture the direct effects of $S$,
* $\sigma_D, \sigma_P$ are stochastic noise amplitudes,
* $\eta_D(t), \eta_P(t)$ are independent standard Wiener processes.

We fix the decay rates $\lambda_D = \lambda_P = 1$, normalizing the time scale due to lack of temporal resolution in the HELIUS dataset.

### Stationarity Assumption

We assume the joint distribution of $(D, P)$ reaches stationarity, leading to a multivariate Ornstein–Uhlenbeck process. Under this assumption, the system has a unique stationary distribution characterized by its covariance structure, which we match to the empirical data.

Let:

$$
\Sigma_{DD} = \mathrm{Var}(D), \quad \Sigma_{PP} = \mathrm{Var}(P), \quad \Sigma_{SS} = \mathrm{Var}(S),
$$

$$
\Sigma_{DP} = \mathrm{Cov}(D, P), \quad \Sigma_{DS} = \mathrm{Cov}(D, S), \quad \Sigma_{PS} = \mathrm{Cov}(P, S).
$$

From the HELIUS data, we use the following observed values:

$\Sigma_{DD} = 1.000, \quad \Sigma_{PP} = 0.807, \quad \Sigma_{SS} = 0.743, \quad \Sigma_{DP} = 0.374, \quad \Sigma_{DS} = 0.357, \quad \Sigma_{PS} = 0.256.$

### Analytical Parameter Derivation

We consider a linear stochastic dynamical system in which depression ($D$) and social precariousness ($P$) evolve under mutual influence and are jointly affected by an exogenous stressor ($S$). We assume stationary dynamics governed by a two-dimensional Ornstein–Uhlenbeck process:

$$
\begin{aligned}
dD_t &= \left(-D_t + \alpha_{DP} P_t + \alpha_{DS} S_t\right)dt + \sigma_D\, dW_{1t}, \\
dP_t &= \left(-P_t + \alpha_{PD} D_t + \alpha_{PS} S_t\right)dt + \sigma_P\, dW_{2t}.
\end{aligned}
$$

Here, we fix the autoregressive decay rates to 1 for identifiability and treat $\alpha_{DP}$ — the influence of precariousness on depression — as a free parameter. All other parameters are then derived to ensure the model reproduces the empirical second-order moments: variances $(\Sigma_{DD}, \Sigma_{PP}, \Sigma_{SS})$ and covariances $(\Sigma_{DP}, \Sigma_{DS}, \Sigma_{PS})$.

Under the stationary solution of the linear SDE system, the following algebraic identities hold:

$$
\begin{aligned}
\Sigma_{DS} &= \alpha_{DS} \Sigma_{SS} + \alpha_{DP} \Sigma_{PS}, \\
\Sigma_{DP} &= \alpha_{DS} \Sigma_{PS} + \alpha_{DP} \Sigma_{PP}, \\
\Sigma_{PS} &= \alpha_{PS} \Sigma_{SS} + \alpha_{PD} \Sigma_{DS}.
\end{aligned}
$$

Solving these simultaneously yields the following closed-form expressions as functions of $\alpha_{DP}$:

$$
\alpha_{DS} = \frac{\Sigma_{DS} - \alpha_{DP} \cdot \Sigma_{PS}}{\Sigma_{SS}},
$$

$$
\alpha_{PD} = \frac{2 \Sigma_{DS} \Sigma_{PS} - 2 \Sigma_{DP} \Sigma_{SS} - \alpha_{DP} \Sigma_{PS}^2 + \alpha_{DP} \Sigma_{PP} \Sigma_{SS}}{\Sigma_{DS}^2 - \Sigma_{DD} \Sigma_{SS}},
$$

$$
\alpha_{PS} = \frac{
\Sigma_{DS}^2 \Sigma_{PS}
+ \Sigma_{PS} \Sigma_{DD} \Sigma_{SS}
+ \Sigma_{DS}(-2 \Sigma_{DP} \Sigma_{SS} - \alpha_{DP} \Sigma_{PS}^2 + \alpha_{DP} \Sigma_{PP} \Sigma_{SS})
}{
\Sigma_{SS}(\Sigma_{DS}^2 - \Sigma_{DD} \Sigma_{SS})
}.
$$

The noise amplitudes $\sigma_D^2$ and $\sigma_P^2$ are derived from the Lyapunov equation associated with the stationary covariance matrix. These yield:

$$
\sigma_D^2 = \frac{-2 \left(\Sigma_{DS}^2 - \Sigma_{DD} \Sigma_{SS} - \Sigma_{DS} \Sigma_{PS} \alpha_{DP} + \Sigma_{DP} \Sigma_{SS} \alpha_{DP} \right)}{\Sigma_{SS}},
$$

$$
\sigma_P^2 = \frac{K}{\Sigma_{SS} (\Sigma_{DD} \Sigma_{SS} - \Sigma_{DS}^2)},
$$

where

$$
\begin{aligned}
K &= 8 \Sigma_{DP} \Sigma_{DS} \Sigma_{PS} \Sigma_{SS}
- 4 \Sigma_{DP}^2 \Sigma_{SS}^2
- 2 \Sigma_{DS}^2 (\Sigma_{PS}^2 + \Sigma_{PP} \Sigma_{SS}) \\
&\quad + 2 \Sigma_{DS} \Sigma_{PS} (\Sigma_{PS}^2 - \Sigma_{PP} \Sigma_{SS}) \alpha_{DP}
+ 2 \Sigma_{SS} (\Sigma_{PP} \Sigma_{SS} - \Sigma_{PS}^2)(\Sigma_{DD} + \Sigma_{DP} \alpha_{DP}).
\end{aligned}
$$

These expressions define a one-parameter family of structurally distinct yet observationally equivalent models, each exactly matching the empirical second-order statistics.

### Constraints on $\alpha_{DP}$

Although all expressions are algebraically valid for arbitrary real $\alpha_{DP}$, admissibility requires that the resulting model remains meaningful and stable.

#### Positivity Constraints

To ensure interpretability, we require all effects and noise terms to be strictly positive:

$$
\alpha_{DS} > 0, \quad \alpha_{PD} > 0, \quad \alpha_{PS} > 0, \quad \sigma_D^2 > 0, \quad \sigma_P^2 > 0.
$$

Each of these is a function of $\alpha_{DP}$, and thus these conditions restrict its allowable range. In particular:

* $\alpha_{PD} = 0$ when $\alpha_{DP} \approx 0.698$,
* $\alpha_{DS} = 0$ when $\alpha_{DP} \approx 1.39$,
* $\sigma_D^2 = 0$ when $\alpha_{DP} \approx 3.30$,
* $\sigma_P^2 = 0$ at a value given exactly by:

$$
\alpha_{DP}^{(\sigma_P = 0)} =
\frac{
- \Sigma_{DD} \Sigma_{PP} \Sigma_{SS}^2 + \Sigma_{DD} \Sigma_{PS}^2 \Sigma_{SS}
+ 2 \Sigma_{DP}^2 \Sigma_{SS}^2 - 4 \Sigma_{DP} \Sigma_{DS} \Sigma_{PS} \Sigma_{SS}
+ \Sigma_{DS}^2 \Sigma_{PP} \Sigma_{SS} + \Sigma_{DS}^2 \Sigma_{PS}^2
}{
\Sigma_{DP} \Sigma_{PP} \Sigma_{SS}^2 - \Sigma_{DP} \Sigma_{PS}^2 \Sigma_{SS}
- \Sigma_{DS} \Sigma_{PP} \Sigma_{PS} \Sigma_{SS} + \Sigma_{DS} \Sigma_{PS}^3
}.
$$

This value falls above $\alpha_{DP} = 0.698$, and thus does not tighten the admissible upper bound.

#### Stability Constraint

The system is stable if the deterministic dynamics converge, which requires the drift matrix

$$
A = \begin{bmatrix}
-1 & \alpha_{DP} \\
\alpha_{PD} & -1
\end{bmatrix}
$$

to have all eigenvalues with negative real parts. This is equivalent to:

$$
\text{Tr}(A) < 0 \quad \text{and} \quad \det(A) = 1 - \alpha_{DP} \alpha_{PD} > 0.
$$

The trace condition is always satisfied ($\text{Tr}(A) = -2$), so stability reduces to:

$$
\alpha_{DP} \cdot \alpha_{PD} < 1.
$$

This condition fails exactly when $\alpha_{PD} \to 0^+$, which coincides with the tightest positivity constraint.

### Admissible Range

Numerical evaluation using empirical covariances from the HELIUS dataset shows that:

* $\alpha_{PD} \to 0$ at $\alpha_{DP} \approx 0.698$,
* All other positivity and stability constraints are looser.

Hence, the maximum valid value is:

$$
\boxed{\alpha_{DP}^{\max} = 0.698}
$$

All simulations and model variants are confined to the admissible domain:

$$
\alpha_{DP} \in [0,\, 0.698]
$$

This guarantees that resulting systems are dynamically stable, fully positive, and consistent with empirical second-order moments.

# Acknowledgements {.unnumbered}

The HELIUS study is conducted by the Amsterdam University Medical Centers, location AMC and the Public Health Service of Amsterdam. Both organisations provided core support for HELIUS. The HELIUS study is also funded by the Dutch Heart Foundation, the Netherlands Organization for Health Research and Development (ZonMw), the European Union (FP-7), and the European Fund for the Integration of non-EU immigrants (EIF). We are most grateful to the participants of the HELIUS study and the management team, research nurses, interviewers, research assistants and other staff who have taken part in gathering the data of this study.